### 2023-10-19 CS2324W-OSP-MCS/Reassessing the Efficacy of Code Challenges/paper.tex
Upload - CS23-00W Papers

% Code Challenge Assessment: A Holistic Approach to Evaluating Engineering Talent
% Author: Felipe Alfonso González
% License: MIT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
    a4paper, % Paper size, use either a4paper or letterpaper
    10pt, % Default font size, can also use 11pt or 12pt, although this is not recommended
    unnumberedsections, % Comment to enable section numbering
    twoside, % Two side traditional mode where headers and footers change between odd and even pages, comment this option to make them fixed
]{LTJournalArticle}

\usepackage{amsmath}

\usepackage{hyperref}

\addbibresource{biblio.bib} % BibLaTeX bibliography file

\runninghead{} % Leave this command empty for no running head

\footertext{} % Leave this command empty for no footer text

\setcounter{page}{1} % The page number of the first page, set this to a higher number if the article is to be part of an issue or larger work

\usepackage{amsmath} % Add this line to load the amsmath package

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{Reassessing the Efficacy of Code Challenges: Debunking the Notion of an Engineer as an All-Knowing AI}

% Authors are listed in a comma-separated list with superscript numbers indicating affiliations
% \thanks{} is used for any text that should be placed in a footnote on the first page, such as the corresponding author's email, journal acceptance dates, a copyright/license notice, keywords, etc
\author{%
    Felipe Alfonso González\textsuperscript{1}\thanks{Corresponding author: \href{mailto:f.alfonso@res-ear.ch}{f.alfonso@res-ear.ch}}\\
    \textit{ Computer Science Engineer}\\
    \footnotesize Institute of Arts and Communication Sciences (IACC), Chile\\[-6pt]
    \footnotesize Candidate for Master in Big Data, ENEB / Isabel I University\\[-6pt]
    \footnotesize\href{mailto:f.alfonso@res-ear.ch}{f.alfonso@res-ear.ch} - 
    \href{https://glzengrg.com}{glzengrg.com} - 
    Twitter: \href{https://twitter.com/felipealfonsog}{@felipealfonsog} - 
    LinkedIn: \href{https://linkedin.com/in/felipealfonsog}{felipealfonsog}\\
    \scriptsize This manuscript has been authored using the typesetting system \LaTeX{}. \\[-6pt]
    \scriptsize This manuscript is released under the BSD 3-clause License. \\
}

% Full-width abstract
\renewcommand{\maketitlehookd}{%
    \begin{abstract}
        \noindent Code challenges have become a popular method for assessing technical skills in software engineering hiring processes. However, this paper aims to challenge the efficacy of such challenges by examining their limitations and questioning the misconception that engineers should possess an all-encompassing knowledge akin to artificial intelligence. This paper advocates for a more holistic assessment approach that takes into account problem-solving skills, teamwork, and adaptability. The proposed approach encourages a shift from testing rote memorization to evaluating an engineer's ability to learn, collaborate, and innovate. This paper presents a comprehensive critique of code challenges and proposes a paradigm shift in the evaluation of engineering talent.

    \end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Output the title section

%----------------------------------------------------------------------------------------
%    ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

In today's rapidly evolving software engineering landscape, the assessment of technical skills for prospective engineers has gained considerable importance. Code challenges, a widely employed approach, have become the de facto standard for evaluating a candidate's coding capabilities within a constrained time frame. However, the efficacy of code challenges as a sole indicator of engineering competence warrants closer examination. This paper aims to critically evaluate the limitations of code challenges and advocate for a more comprehensive assessment approach that reflects the multifaceted nature of modern software engineering \cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.

\subsection{Limitations of Code Challenges}

Code challenges, while providing valuable insights into an engineer's problem-solving prowess, suffer from inherent limitations. By design, code challenges prioritize algorithmic efficiency and rapid problem-solving over other crucial attributes such as creativity, adaptability, and teamwork. This singular focus can lead to a skewed evaluation of a candidate's abilities, potentially overlooking the soft skills that are integral to successful software development.

%%%%%%% new stuff to be improved

A pair programming is a really bad idea to try to see the potential in a possible candidate, it doesn't adjust the needs either way, because you are going to interact with someone you don't know, someone with a completely unknown attitude and different abilities to complete a task, therefore as a candidate, there will be chances that you interview will be a complete failure, a complete catastrophe, it's stupid, even if you have been preparing yourself for months something unexpected could happen.

%Insert stats about this a math stat

%——————-

HISTORY OF A LONG STORY TO UNDERSTAND DEVELOPERS

ITS PRETTY UTIPIC TO BELIVE THAT SOMEONE WILL BECOME PRACTIXALLY A AI IN A TECHNICAL INTERVIEW. it's a complete non sense.

IN THE FUTURE WE MIGHT SEE A CHANCE TO SEE AN EVEN NEW BRANCH IN ENGKNEERng, like prompt engineering with different approaches.

PSYCHOLOGICAL ASPECTS OF TECH INTERVIEWS

the importance of having great open source projects : it's a source for different eyes about how someone can handle code and even a whole infrastructure

Junior developers can do a big contribution

——

POSSIBLE OPTIONS TO RESOLVE THIS ISSUES:
to expose candidates to the exploration and research in different topics in engineering and or computer science. testing something but considering the possibility for the candidate to research and explore solutions for a real world problem.

TRUST : a factor in rhe process of finding the right csndisste

the big business of the Bootcamps to ‘teach’ about programming and while toy pay them you they assure you don't go through technical interviews or that kind of procedures

\subsection{High Software Engineer Turnover: The Impact of Code Challenges}

Unfortunately, anyone involved in the field of computer science engineering is well aware of the stress and questionable efficacy associated with code challenges. It seems counterintuitive when one is asked to undergo a testing process where you're required to be in front of a group of people, often utilizing webcams, and using specialized software for pair programming.

The ubiquity of code challenges in the hiring process for software engineers has raised eyebrows and provoked discussions within the industry. The question arises: do these challenges truly assess the skills and qualities necessary for a successful software engineer, or are they a source of unnecessary anxiety and a potential deterrent for talented candidates?

The concept of evaluating technical abilities through problem-solving assessments is not inherently flawed. However, the implementation of such evaluations in the form of code challenges may not always align with the real-world scenarios that software engineers encounter. This discrepancy between the test environment and the actual work environment can lead to a skewed representation of an engineer's capabilities.

The high turnover rate among software engineers is a concern that industry leaders are actively trying to address. Factors contributing to this turnover include burnout, lack of growth opportunities, and misalignment between expectations and reality. The imposition of code challenges, especially those that may not accurately mirror the skills required for the job, could exacerbate these challenges and further contribute to the issue of high turnover.

\subsection{Statistical Insight}

To quantitatively highlight the impact of code challenges on software engineer turnover, we can formulate a statistical equation. Let $\text{Turnover}$ represent the high software engineer turnover, and $\text{Challenges}$ denote the prevalence of code challenges in hiring processes. We can posit the following equation:

\begin{equation}
    \text{Turnover} = k \times \text{Challenges}^2
\end{equation}

Where $k$ is a constant representing the amplification of turnover due to code challenges. This equation showcases how the use of code challenges can lead to a nonlinear increase in the software engineer turnover rate. The undue stress and potential misalignment of assessment with actual job requirements contribute to the rapid growth of turnover rates.

In conclusion, it's imperative to reassess the role of code challenges in the evaluation of software engineers. While the intention behind such assessments may be to gauge technical prowess, their application should be scrutinized to ensure they align with the complexities and demands of real-world software development. It's crucial to create an inclusive and accurate evaluation process that doesn't add unnecessary stress and hinder the retention of skilled engineers in the industry.

\subsection{Limitations of Traditional Technical Interviews}

Traditional technical interviews, while widely used in the software engineering hiring process, present notable limitations that call into question their effectiveness in accurately assessing candidates' expertise. In this section, we examine these limitations and shed light on the discrepancies between the intended purpose of such interviews and their actual outcomes.

\subsubsection*{Narrow Focus and Incomplete Assessment}

One glaring limitation of traditional technical interviews is their narrow focus on specific aspects, such as assessing candidates' knowledge of a particular programming language. This approach fails to provide a comprehensive evaluation of candidates' skills, overlooking critical attributes like problem-solving abilities, collaboration aptitude, and adaptability to evolving technologies. Consequently, the interview's outcomes may not accurately reflect a candidate's overall potential or fit within a dynamic software development environment.

\subsubsection*{Undue Stress and Performance Anxiety}

Technical interviews often create a high-stress environment for candidates, leading to performance anxiety that can hinder their true capabilities. The pressure of being scrutinized and evaluated on the spot can lead to nervousness and suboptimal performance, rendering the interview results less indicative of a candidate's actual skills and expertise. The artificial nature of these interviews fails to replicate real-world scenarios where problem-solving often occurs in a collaborative and less time-constrained environment.

%%%%%% new linea about health - mental health - state of mind

Another thing that's really important is to understand how stressful it is to be updated and hence updated for any sort of interview where developers require e to kind of demonstrate that.. how? usually by projects or work done.. but it is not so simple.

\subsubsection*{Misalignment with Industry Realities}

Another significant limitation is the misalignment between the structure of technical interviews and the complex realities of the software engineering industry. Modern software development emphasizes teamwork, communication, and iterative problem-solving, which traditional technical interviews tend to overlook. Isolating candidates from the collaborative context in which they would typically work fails to provide an accurate representation of their potential contributions to a development team.

%new lines 
%%%%%%%%%%%%%%%%%%%%%%

Frequently encountered online are a plethora of courses vying for attention, accompanied by enticing advertisements that extol the virtues of their offerings. Promises of rapid entry into the IT job market abound, painting an alluring picture that often veers far from reality. The stark truth is that the journey towards becoming proficient in IT is fraught with demanding challenges, far more arduous than many may anticipate. Beneath the veneer of ease, lies a reality that remains hidden from those who readily embrace these courses \cite{dweck2006mindset,duckworth2016grit}.

These promotions, however well-intentioned, frequently overlook the rigorous trials and tribulations that await aspiring learners. The trials, often relentless and formidable, stand as a formidable test of one’s resilience and determination. Enrollees, spurred on by these enticing prospects, seldom fathom the intensity and depth of commitment required to surmount these hurdles. Embarking on the journey without a genuine understanding of the inherent difficulties could lead to disillusionment and a potential crisis of confidence.

It is of paramount importance to underscore the significance of acquiring a solid educational foundation in engineering disciplines. The allure of expedited success and the immediate allure of lucrative job prospects should not eclipse the value of comprehensive learning and development. For those who tread the path without first dedicating themselves to the study of engineering principles, the result could be more than mere disappointment; it could lead to a pervasive feeling of unpreparedness and inadequacy.

In the pursuit of a career in IT, the path is not a sprint but a marathon. True success is built on a sturdy foundation of knowledge, perseverance, and rigorous training. The misperceptions perpetuated by grandiose promises can culminate in a disheartening realization of the true challenges ahead. The prudent course of action is to approach the journey with open eyes and an earnest commitment to a disciplined education. In doing so, one can transcend the superficial allure of swift success and instead forge a solid and lasting trajectory towards a fulfilling career in the realm of IT \cite{tetlock2007illusion}.

\subsubsection*{Bias and Exclusion}

Traditional technical interviews may inadvertently introduce bias into the assessment process. The focus on specific technical skills can favor candidates from certain educational or experiential backgrounds while excluding qualified individuals who possess a broader skill set. This bias can lead to a lack of diversity within the engineering workforce and hinder the discovery of exceptional talents with non-traditional paths.

\subsubsection*{Ethical Concerns}

Ethical considerations surrounding technical interviews also deserve attention. Subjecting candidates to high-stakes evaluations without adequate consideration for their psychological well-being raises ethical questions. The potential negative impact on candidates' self-esteem, regardless of their performance, calls into question the ethical implications of relying solely on traditional technical interviews as a hiring tool.

In conclusion, while traditional technical interviews have been a staple in the software engineering hiring process, their limitations are significant. The narrow focus, stress-inducing nature, misalignment with industry practices, potential bias, and ethical concerns highlight the need for a more comprehensive and thoughtful approach to evaluating candidates' suitability for modern software development roles.

\subsection{Questioning the All-Knowing Engineer Paradigm}

The prevailing notion of an engineer as an all-knowing entity well-versed in every programming language, framework, and technology is unrealistic and counterproductive. In an era of continuous technological advancements, the expectation for engineers to possess an encyclopedic knowledge is untenable. Rather than striving for an impossible ideal, engineers should be encouraged to embrace a growth mindset, focusing on their capacity to learn, adapt, and collaborate effectively.

\subsection{The Need for a Holistic Assessment}

A holistic assessment approach acknowledges that software engineering is not solely about writing flawless code. Instead, it encompasses a spectrum of skills ranging from critical thinking to clear communication and from teamwork to innovation. A comprehensive evaluation process should account for these diverse dimensions, enabling organizations to identify candidates who are equipped to thrive in the dynamic and collaborative environment of modern software development.

%------------------------------------------------

\section{Methodologies}

To discern the effectiveness of code challenges and explore alternative assessment methods, a thorough analysis of existing literature, industry practices, and case studies was conducted. The goal was to elucidate the correlation between success in code challenges and long-term engineering accomplishments.

\subsection{Sample Sites \& Processing}

The study encompassed an extensive array of software engineering platforms, each hosting a variety of code challenges. Analyzing the outcomes of participants, including their subsequent contributions to open-source projects, was integral to evaluating the predictive value of code challenge success.

%------------------------------------------------

\section{The Fallacy of Comprehensive Expertise}

Engineering excellence is not synonymous with being a walking repository of technical minutiae. Rather, it involves cultivating foundational understanding and embracing the aptitude to navigate novel territories. The software engineering landscape is fluid, with languages and tools frequently evolving. Rote memorization pales in comparison to an engineer's capacity to comprehend fundamental principles and apply them innovatively.

\section{Incorporating Real-World Complexity}

Software development extends beyond isolated coding sprints. Collaborative team efforts, iterative design, and communication play pivotal roles. A comprehensive evaluation should emulate real-world scenarios, emphasizing how candidates contribute to a team's dynamic, provide constructive feedback, and navigate through diverse perspectives.

%------------------------------------------------

\section{Results and Implications}

The study revealed that while code challenges serve as an initial filter for technical skills, their predictive value diminishes concerning an engineer's collaborative aptitude, adaptability, and capacity to contribute to complex projects. Candidates excelling in code challenges do not inherently outperform their peers in real-world projects that demand adaptability and collective problem-solving.

\section{Learning from Failures}

The capacity to learn from setbacks is instrumental in software engineering. Rather than penalizing candidates for errors during assessments, it is more instructive to assess their approach to challenges, iteration on solutions, and incorporation of lessons learned. This adaptive resilience is a hallmark of a competent engineer.

%------------------------------------------------

\section{Redefining Engineering Excellence}

Engineering is an amalgamation of cognitive agility, problem-solving acumen, and interpersonal skills. An engineer's worth lies not in the mastery of a predefined set of skills, but in their ability to learn, collaborate, and innovate. By championing these traits, organizations foster a culture of resilience and evolution.

\section{Promoting a Paradigm Shift}

A transformative shift in assessment paradigms is overdue. Code challenges, while valuable, should be complemented with broader evaluations. Organizations stand to gain by emphasizing holistic attributes such as adaptability, critical thinking, and collaboration. The engineering landscape beckons for engineers who can not only code, but also envision, communicate, and innovate.

%------------------------------------------------

\section{Conclusion}

This paper challenges the prevailing narrative that engineers must possess exhaustive knowledge and underscores the limitations of code challenges. It advocates for an inclusive assessment approach that values problem-solving, adaptability, and teamwork. Engineering is an ever-evolving discipline, necessitating engineers who can navigate complexities and contribute meaningfully to the collaborative process \cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.

%----------------------------------------------------------------------------------------
%    REFERENCES
%----------------------------------------------------------------------------------------

\printbibliography % Output the bibliography

%----------------------------------------------------------------------------------------

\end{document}

### 2023-10-19 CS2324W-OSP-MCS/Reevaluando la Eficacia de los Desafíos de Código/paper.tex
Upload - CS23-00W Papers

--- 2023-10-19 CS2324W-OSP-MCS/Reassessing the Efficacy of Code Challenges/paper.tex
@@
-% Code Challenge Assessment: A Holistic Approach to Evaluating Engineering Talent
+% ----------------------------------------------
 % Author: Felipe Alfonso González
@@
 
-\title{Reassessing the Efficacy of Code Challenges: Debunking the Notion of an Engineer as an All-Knowing AI}
+\title{Reevaluando la Eficacia de los Desafíos de Código: Desmitificando la Idea de un Ingeniero como una IA Omnisciente}
 
@@
 PSYCHOLOGICAL ASPECTS OF TECH INTERVIEWS 
-
-the importance of having great open source projects : it's a source for different eyes about how someone can handle code and even a whole infrastructure
-
-Junior developers can do a big contribution
-
 ——
@@
 TRUST : a factor in rhe process of finding the right csndisste
-
-the big business of the Bootcamps to ‘teach’ about programming and while toy pay them you they assure you don't go through technical interviews or that kind of procedures
 

### 2023-10-19 CS2325W-MCS/BigData-paper-ENEB/paper.tex
Upload - CS23-00W Papers

% --------------------------------------------
% Author: Felipe Alfonso González
% License: MIT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
    a4paper, % Paper size, use either a4paper or letterpaper
    10pt, % Default font size, can also use 11pt or 12pt, although this is not recommended
    unnumberedsections, % Comment to enable section numbering
    twoside, % Two side traditional mode where headers and footers change between odd and even pages, comment this option to make them fixed
]{LTJournalArticle}

\usepackage{amsmath}

\usepackage{hyperref}

\addbibresource{biblio.bib} % BibLaTeX bibliography file

\runninghead{} % Leave this command empty for no running head

\footertext{} % Leave this command empty for no footer text

\setcounter{page}{1} % The page number of the first page, set this to a higher number if the article is to be part of an issue or larger work

\usepackage{amsmath} % Add this line to load the amsmath package

\usepackage{graphicx}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{Big Data en Marketing}

% Authors are listed in a comma-separated list with superscript numbers indicating affiliations
% \thanks{} is used for any text that should be placed in a footnote on the first page, such as the corresponding author's email, journal acceptance dates, a copyright/license notice, keywords, etc
\author{%
    Felipe Alfonso González\textsuperscript{1}\thanks{Corresponding author: \href{mailto:f.alfonso@res-ear.ch}{f.alfonso@res-ear.ch}}\\
    \textit{ Computer Science Engineer}\\
    \footnotesize Institute of Arts and Communication Sciences (IACC), Chile\\[-6pt]
    \footnotesize Candidate for Master in Big Data, ENEB / Isabel I University\\[-6pt]
    \footnotesize\href{mailto:f.alfonso@res-ear.ch}{f.alfonso@res-ear.ch} - 
    \href{https://glzengrg.com}{glzengrg.com} - 
    Twitter: \href{https://twitter.com/felipealfonsog}{@felipealfonsog} - 
    LinkedIn: \href{https://linkedin.com/in/felipealfonsog}{felipealfonsog}\\
    \scriptsize This manuscript has been authored using the typesetting system \LaTeX{}. \\[-6pt]
    \scriptsize This manuscript is released under the BSD 3-clause License. \\
}

% Full-width abstract
\renewcommand{\maketitlehookd}{%
    \begin{abstract}
        \noindent En el entorno altamente competitivo actual, la retención y satisfacción de clientes son imperativos fundamentales. A lo largo de la historia del marketing, se han implementado diversas estrategias para impulsar las ventas y fomentar la lealtad, pero en la actualidad, la personalización se ha vuelto esencial en este contexto de feroz competencia. 
        Este paper explora el marketing desde la perspectiva del big data, destacando cómo las empresas pueden aprovechar los datos disponibles sobre clientes y entorno para diseñar estrategias altamente personalizadas. Ya no es suficiente con promociones genéricas; ahora, la clave radica en ofertas específicas que se alineen con las preferencias individuales de cada cliente. 
        Al reconocer que los clientes son la piedra angular de una empresa, se enfatiza la importancia de planificar acciones y estrategias personalizadas. Estos enfoques no solo aseguran la supervivencia de la empresa en un mercado dinámico, sino que también generan lealtad y fidelidad del cliente, aportando un valor significativo al negocio.

    \end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Output the title section

%----------------------------------------------------------------------------------------
%    ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

En el panorama empresarial actual, marcado por una competencia feroz, la retención de clientes y su satisfacción plena se presentan como objetivos cruciales para la supervivencia y el éxito de cualquier empresa. A lo largo de la historia, el marketing ha evolucionado desde estrategias generales hasta reconocer la necesidad de personalización en un mundo donde cada cliente es único. Este paper aborda el desafío de implementar estrategias de marketing desde la perspectiva del big data, explorando cómo la información sobre clientes y su entorno puede ser aprovechada para diseñar ofertas altamente específicas. En un contexto donde los clientes son la esencia misma de la empresa, este enfoque personalizado no solo asegura la permanencia en el mercado, sino que también cultiva la lealtad y la conexión duradera con la clientela, generando un valor distintivo para el negocio. \cite{ENEB2023}.

\subsection{El Marketing: necesidades, deseos y demandas}

%%% Necesidades, deseos, demandas 
Todas las personas pueden tener diferentes tipos de necesidades, que son parte de la esencia humana y pueden ser físicas (como alimentarse), seguridad, o sociales o individuales; pero sus deseos pueden estar alineados con sus necesidades, que pueden ser físicas, como alimentarse o sentirse seguro, o algún tipo de realización personal, entre otras. Pueden variedad según incluso por cuan importantes deseas sentirse. En relación a las demandas, son los deseos, y estas se convierten en demanda cuando existe esa capacidad adquisitiva.

%%% Productos o Servicios
El ser humano, satisfaces sus necesidades con productos y servicios. Pueden ser estos intangibles como los servicios en algún área, o tangibles. Pueden ser susceptibles a algo que puede ser ofrecido para satisfacer necesidades y deseos.

%%% Valor coste y satisfacción
 Todos los consumidores toman decisiones basadas en expectativas que les plantean las diferentes ofertas. Las expectativas de valor que les plantean en las diferentes ofertas.

Los consumidores eligen productos según las expectativas netas de valor, la diferencia entre aspectos positivos y negativos esperados. Este valor impacta en la satisfacción del cliente y su comportamiento futuro, al comparar expectativas con la experiencia real de compra.

El marketing surge del intercambio, un proceso donde las personas satisfacen necesidades a través de ofrecer algo a cambio de dinero, bienes o servicios. Para que el intercambio ocurra, deben existir al menos dos partes con algo de valor para la otra, capacidad de comunicación, libertad de aceptar o rechazar la oferta, y la creencia de que es apropiado tratar con la otra parte. El intercambio es un proceso continuo, y cuando se llega a un acuerdo, se denomina transacción. El marketing de relaciones busca construir relaciones duraderas basadas en confianza mediante la entrega consistente de productos de calidad, buen servicio y precios razonables.

Un mercado es el conjunto de consumidores que comparten una necesidad o deseo y están dispuestos a intercambiar elementos de valor. Su tamaño depende del número de personas con la necesidad específica, recursos atractivos para otros, y la disposición de intercambiar.

El marketing se centra en actividades relacionadas con los mercados para facilitar intercambios potenciales que satisfacen necesidades y deseos. En este proceso, participan dos partes: el buscador de intercambios, usualmente una empresa compitiendo en un mercado, y el receptor. La eficacia del buscador se ve afectada por factores del entorno como demografía, economía, tecnología y aspectos socio-culturales.

%%  Inserting image 
\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\linewidth]{./images/graph-mkt.jpg}
  \caption{Partes que confirman el marketing}
  \label{fig:etiqueta}
\end{figure}
%% End insert image

\subsection{Gestión del marketing}

La gestión de marketing implica planificar y ejecutar estrategias para productos, precios, promoción y distribución. Se centra en la gestión de la demanda, ajustando la oferta a las cambiantes condiciones del mercado. La retención de clientes es esencial, dada la alta inversión en la adquisición de nuevos clientes y el riesgo de repercusiones negativas en redes sociales. Las orientaciones empresariales, como producción, producto, ventas, marketing y marketing social, reflejan filosofías distintas que impactan las estrategias de marketing. En resumen, la gestión de marketing es un proceso dinámico que busca no solo intercambios, sino también relaciones a largo plazo con los clientes.

%% Enfoque de producción - pp. 8

%------------------------------------------------

\section{Methodologies}

To discern the effectiveness of code challenges and explore alternative assessment methods, a thorough analysis of existing literature, industry practices, and case studies was conducted. The goal was to elucidate the correlation between success in code \section{Redefining Engineering Excellence}

Engineering is an amalgamation of cognitive agility, problem-solving acumen, and interpersonal skills. An engineer's worth lies not in the mastery of a predefined set of skills, but in their ability to learn, collaborate, and innovate. By championing these traits, organizations foster a culture of resilience and evolution.

\section{Promoting a Paradigm Shift}

A transformative shift in assessment paradigms is overdue. Code challenges, while valuable, should be complemented with broader evaluations. Organizations stand to gain by emphasizing holistic attributes such as adaptability, critical thinking, and collaboration. The engineering landscape beckons for engineers who can not only code, but also envision, communicate, and innovate.

%------------------------------------------------

\section{Conclusion}

This paper challenges the prevailing narrative that engineers must possess exhaustive knowledge and underscores the limitations of code challenges. It advocates for an inclusive assessment approach that values problem-solving, adaptability, and teamwork. Engineering is an ever-evolving discipline, necessitating engineers who can navigate complexities and contribute meaningfully to the collaborative process \cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.

%----------------------------------------------------------------------------------------
%    REFERENCES
%----------------------------------------------------------------------------------------

\printbibliography % Output the bibliography

%----------------------------------------------------------------------------------------

\end{document}

### 2023-10-19 CS2325W-MCS/BigData-paper-ENEB/paper.tex
Updates

@@
 % Author: Felipe Alfonso González
-% License: MIT
+% https://github.com/felipealfonsog
+% f.alfonso@res-ear.ch / felipe.alfonso.glz@gmail.com
+% License: MIT / BSD 3-clause License
 %
@@
 
+\usepackage{amsmath} % amsmath package
+
+\usepackage{graphicx}
+
+\usepackage{comment}
+
 \addbibresource{biblio.bib} % BibLaTeX bibliography file
@@
 \setcounter{page}{1} % The page number of the first page, set this to a higher number if the article is to be part of an issue or larger work
-
-\usepackage{amsmath} % Add this line to load the amsmath package
-
-\usepackage{graphicx}
 
@@
 
+%----------------------------------------------------------------------------------------
+
 % Full-width abstract
@@
 %----------------------------------------------------------------------------------------
-%    ARTICLE CONTENTS
+%    PAPER CONTENTS
 %----------------------------------------------------------------------------------------
@@
 
-%%% Necesidades, deseos, demandas 
 Todas las personas pueden tener diferentes tipos de necesidades, que son parte de la esencia humana y pueden ser físicas (como alimentarse), seguridad, o sociales o individuales; pero sus deseos pueden estar alineados con sus necesidades, que pueden ser físicas, como alimentarse o sentirse seguro, o algún tipo de realización personal, entre otras.
@@
 
-%%% Productos o Servicios
 El ser humano, satisfaces sus necesidades con productos y servicios.
@@
 
-%%% Valor coste y satisfacción
  Todos los consumidores toman decisiones basadas en expectativas que les plantean las diferentes ofertas. Las expectativas de valor que les plantean en las diferentes ofertas.
@@
 
-La gestión de marketing implica planificar y ejecutar estrategias para productos, precios, promoción y distribución.
-Se centra en la gestión de la demanda, ajustando la oferta a las cambiantes condiciones del mercado.
-La retención de clientes es esencial, dada la alta inversión en la adquisición de nuevos clientes y el riesgo de repercusiones negativas en redes sociales.
-Las orientaciones empresariales, como producción, producto, ventas, marketing y marketing social, reflejan filosofías distintas que impactan las estrategias de marketing.
-En resumen, la gestión de marketing es un proceso dinámico que busca no solo intercambios, sino también relaciones a largo plazo con los clientes.
+La gestión de marketing se ocupa de planificar y ejecutar estrategias para productos, precios, promoción y distribución, con el objetivo de crear intercambios que satisfagan tanto objetivos individuales como organizativos.
+Se enfoca en la gestión de la demanda, influenciando su nivel, momento y composición para alcanzar metas organizativas.
+En un entorno empresarial, la gestión de la demanda busca ajustarse a períodos variables, controlando la oferta para mantenerla alineada con las expectativas del mercado.
 
-%% Enfoque de producción - pp. 8
+Esta gestión no solo trata con la demanda sino también con los clientes actuales, nuevos y potenciales.
+A lo largo del tiempo, ha evolucionado hacia la creación de relaciones a largo plazo con los clientes, especialmente en la era del big data marketing.
+La retención y fidelización de clientes son fundamentales en la actual competencia empresarial, lo que destaca la importancia del trato personalizado y la prevención de la pérdida de clientes.
+
+La adquisición de nuevos clientes se considera costosa, siendo aproximadamente cinco veces más costoso que mantener a un cliente existente satisfecho.
+Además, la pérdida de un cliente habitual puede tener repercusiones más amplias, ya que la insatisfacción puede propagarse a través de redes sociales, afectando la imagen de la empresa.
+
+En relación con las orientaciones empresariales, existen cinco enfoques: producción, producto, ventas, marketing y marketing social.
+Cada enfoque refleja la filosofía y los intereses que guían a las organizaciones en su desarrollo empresarial.
+Estos aspectos son fundamentales en la planificación estratégica y el proceso de marketing.
+
+\subsection{Enfoque de producción}
+
+El enfoque de producción masiva es el más antiguo en la dirección de intercambios, priorizando la disponibilidad y el bajo costo de los productos.
+Se centra en alcanzar economías de escala y una amplia distribución, con énfasis en la producción en lugar de la creación de clientes.
+Predomina en situaciones donde la demanda supera la oferta, como en países del tercer mundo, o cuando se busca reducir costos para ampliar el mercado.
+Un ejemplo destacado es el enfoque de producción masiva y precios bajos de Henry Ford en 1900, que perfeccionó la producción de automóviles para hacerlos más accesibles a la población.
+En resumen, implica producir en gran cantidad para lograr costos bajos y precios de venta asequibles.
+
+\subsection{Enfoque de producto}
+
+El segundo enfoque destaca la importancia de ofrecer productos de alta calidad o con resultados superiores para atraer a los consumidores.
+Implica esfuerzos constantes en la fabricación de productos de excelencia y su mejora continua.
+Los directores de empresas que siguen este enfoque creen que los compradores valoran la calidad del producto, centrándose en las ventajas y beneficios.
+Sin embargo, puede haber descuidos en la atención al mercado y la comprensión de la demanda, llevando a la miopía de marketing.
+Esto podría resultar en productos que no cumplen con las expectativas y carecen de utilidad.
+Equilibrar la calidad del producto con una comprensión profunda del mercado es crucial para el éxito sostenible de la empresa.
+
+\subsection{Enfoque de ventas}
+
+El enfoque de ventas se centra exclusivamente en impulsar las ventas mediante estrategias agresivas de venta y promoción.
+Su objetivo principal es vender la mayor cantidad posible de productos, priorizando la venta de lo que se produce en lugar de producir según la demanda.
+Este enfoque se utiliza principalmente para bienes no buscados, aquellos que los consumidores no tienen la intención de adquirir habitualmente.
+Sin embargo, también se aplica a productos buscados o de áreas no lucrativas.
+
+Este enfoque también es comúnmente empleado por las empresas cuando tienen capacidad productiva excedente y buscan liquidar el stock disponible.
+
+\subsection{Enfoque de marketing}
+
+El enfoque de marketing se diferencia de los anteriores al poner énfasis en identificar las necesidades y deseos de los clientes para satisfacerlos efectivamente. En este enfoque:
+
+- Se centra en las necesidades del comprador.
+- Busca satisfacer las necesidades del cliente a través del producto y los beneficios asociados con su entrega y consumo.
+
+Este enfoque se fundamenta en cuatro pilares esenciales: definición del mercado, orientación al cliente, coordinación de marketing y rentabilidad.
+
+\section{Definición del mercado}
+
+Cada empresa debe cuidadosamente definir su público objetivo, ya que es imposible operar en todos los mercados y satisfacer todas las necesidades.
+La elección del público objetivo influye en las acciones que la empresa emprenderá para satisfacer las necesidades de ese grupo específico.
+Por ejemplo, una empresa de viajes organizados debe determinar con precisión el tipo de cliente al que se dirige, ya que esto condicionará el producto y la presentación del mismo.
+La planificación de viajes de lujo para recién casados difiere significativamente de organizar viajes de fin de curso para universitarios o viajes del IMSERSO.
+
+\section{Orientación al cliente}
+
+Una vez definido el público objetivo, es esencial comprender sus necesidades desde su perspectiva, no desde la de la empresa. Esto implica identificar cinco tipos de necesidades:
+
+Necesidades declaradas: El cliente desea un coche asequible.
+Necesidades reales: El cliente busca un coche con bajos costos operativos, no solo un precio de compra bajo.
+Necesidades no declaradas: El cliente desea recibir un buen trato y servicio de la empresa vendedora.
+Necesidades de deleite: El cliente recibe regalos o servicios adicionales con la compra del coche.
+Necesidades secretas: El cliente quiere ser percibido como alguien que obtiene un buen valor por su dinero.
+Para satisfacer estas necesidades, el departamento de marketing debe investigar los aspectos más valorados por el público objetivo.
+El objetivo principal es vender a través de la satisfacción del cliente, ya que un cliente satisfecho:
+
+Compra repetidamente y es leal.
+Adquiere nuevos productos lanzados por la empresa.
+Habla positivamente sobre la empresa y sus productos.
+Presta menos atención a las marcas y publicidad de competidores.
+Es más rentable de servir debido al conocimiento detallado sobre sus preferencias.
+Contribuye con ideas para nuevos productos y servicios.
+La satisfacción del cliente es crucial, ya que un cliente insatisfecho puede dañar la reputación de la empresa.
+Medir regularmente el grado de satisfacción, incluyendo quejas y reclamaciones, es vital para la mejora continua.
+
+\section{Coordinación de marketing}
+
+El éxito de una empresa depende crucialmente de la coordinación. En el ámbito de la coordinación de marketing, dos aspectos fundamentales deben ser considerados:
+
+Coordinación de funciones de marketing: Es esencial que las diversas funciones de marketing, como ventas, publicidad, gestión de productos, investigación, entre otras, trabajen de manera conjunta y coordinada desde la perspectiva del consumidor.
+Este enfoque busca lograr la plena satisfacción de los clientes.
+
+Coordinación del departamento de marketing con el resto de la empresa: El departamento de marketing no puede operar de manera aislada; debe estar coordinado con otros departamentos de la empresa.
+La falta de coordinación interna dificulta la plena satisfacción de los clientes.
+
+La empresa debe llevar a cabo tanto acciones de marketing externo como interno.
+El marketing interno se refiere a la contratación, capacitación y motivación del personal para que brinde un servicio adecuado a los clientes.
+Es crucial reconocer que el marketing interno precede al marketing externo; prometer un excelente servicio a los clientes carece de significado si internamente la empresa no está preparada para cumplir con esas promesas.
+
+\section{Rentabilidad}
+
+La rentabilidad es un objetivo clave del enfoque de marketing, buscando contribuir a que las organizaciones alcancen sus metas, generalmente centradas en maximizar beneficios.
+Por lo tanto, el departamento de marketing no solo trabaja para satisfacer a los clientes, sino también para lograr rentabilidad.
+
+\subsection{Enfoque de marketing social}
+
+El enfoque de marketing social va más allá de la satisfacción del cliente y la rentabilidad, considerando también el impacto en la sociedad y el interés público en la toma de decisiones.
+La Responsabilidad Social Corporativa (RSC) es esencial en este enfoque, gestionando los impactos de la actividad empresarial en clientes, empleados, accionistas, medio ambiente y la sociedad en general.
+
+La RSC equilibra el crecimiento económico, el bienestar social y la sostenibilidad ambiental.
+Esto no solo beneficia a la sociedad, sino que también aporta ventajas para la empresa, como mayor productividad, lealtad del cliente, acceso a mercados y credibilidad.
+La transparencia y la adaptabilidad a las expectativas de la sociedad son clave para lograr la responsabilidad social y el éxito a largo plazo.
+
+El marketing social no solo busca beneficios para la empresa y los clientes, sino también para la sociedad y el entorno.
+Se destaca la importancia de la adaptación al mercado y a las necesidades de los diferentes grupos de interés.
+En este contexto, el big data marketing surge como una evolución del marketing tradicional, proporcionando herramientas avanzadas para comprender y abordar las complejidades del mercado actual.
+
+\section{Big Data Marketing}
+
+El big data marketing se centra en aprovechar la gran cantidad de datos disponibles sobre clientes para personalizar estrategias y promociones de manera individualizada, evitando campañas genéricas.
+A diferencia de la segmentación tradicional, el enfoque del big data marketing se dirige exclusivamente a un cliente concreto.
+
+Los elementos clave para trabajar con éxito el big data marketing son:
+
+Tener una estrategia: Es esencial definir qué tipo de cliente se desea abordar y cómo llevar a cabo la interacción.
+Esto implica establecer una estrategia de interacción y definir la estrategia analítica, determinando qué datos son relevantes para el negocio.
+
+Personalización, relevancia y recompensa: Implementar un sistema de big data conlleva un alto costo, por lo que es crucial utilizar el conocimiento del cliente para personalizar ofertas y comunicaciones, identificar clientes clave y recompensarlos según el valor que aportan.
+
+En el big data marketing, el cliente se convierte en el protagonista principal, y se debe adaptar el marketing al ciclo de vida del cliente.
+El customer analytics, herramienta utilizada para analizar datos de clientes, se vuelve crítico en este entorno de alto conocimiento del cliente.
+
+El ciclo de vida del cliente consta de seis fases:
+
+Descubrir: Conocer la marca.
+Explorar: Evaluar si gusta.
+Comprar
+Usar
+Preguntar
+Ser un fan: Convertirse en un cliente de alto valor.
+La gestión efectiva de este ciclo de vida implica que un cliente que descubre y explora la marca llegue a comprar y convertirse en un fan.
+Cada fase tiene objetivos de negocio específicos y requiere métodos analíticos particulares para su gestión óptima.
+
+%%  Inserting image 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\linewidth]{./images/bgmkt.jpg}
+  \caption{Fases, métodos y objetivos del Big Data Marketing}
+  \label{fig:etiqueta}
+\end{figure}
+%% End insert image
+
+En el contexto de la gestión de clientes mediante el big data marketing, es fundamental considerar el concepto de "customer journey" o viaje del cliente, que se analiza en el módulo de Customer Analytics.
+Este enfoque implica mapear las etapas, interacciones, canales y elementos por los cuales se desplaza el cliente dentro de la empresa.
+El customer journey no se limita a la etapa de cliente hasta su finalización; puede centrarse en un punto específico, como el proceso de compra o la atención al cliente para resolver un problema.
+
+Al construir estos mapas, se deben tener en cuenta varios aspectos:
+
+Identificación del cliente: Es esencial conocer y entender al cliente desde su perspectiva.
+
+Comprensión de las etapas del vínculo: Observar cómo se siente el cliente en cada interacción con la empresa, servicio o producto.
+
+Registro de indecisiones y motivaciones: Conocer lo que motiva a los consumidores y sus inquietudes.
+
+Mapeo de interacciones: Identificar los puntos de contacto repetitivos en las relaciones con la empresa.
+
+Análisis de instancias clave: Detectar momentos cruciales en los que el cliente se siente feliz, enojado o perdido, ya que impactan en la decisión de avanzar o abandonar.
+
+Procesos internos de la empresa: Opcionalmente, agregar procesos internos de la empresa relacionados con cada interacción del cliente para entender cómo contribuyen a la experiencia.
+
+Oportunidades y sentimientos: Comprender las sensaciones y sentimientos del cliente en cada etapa o interacción, especialmente identificar lo que le molesta o incomoda.
+
+Trazar este viaje ayuda a comprender mejor las motivaciones y la satisfacción del cliente, facilitando una gestión personalizada que guía al cliente a través de todas las fases del ciclo de vida.
+Esto solo se logra mediante un análisis detallado del cliente y la implementación de acciones de marketing personalizadas basadas en los resultados del análisis.
+
+El big data marketing representa un cambio significativo en las relaciones con los clientes, pasando de una comunicación masiva a una personalizada que se adapta a las necesidades individuales de cada cliente.
+Aunque la segmentación sigue siendo importante, la personalización adquiere una relevancia aún mayor.
+
+Estas consideraciones preparan el terreno para explorar las herramientas esenciales en el ámbito del big data marketing.
+
+\subsection{Herramientas y prácticas de big data marketing}
+
+Para implementar con éxito el big data marketing, es crucial emplear diversas herramientas además del análisis de clientes. Estas herramientas incluyen:
+
+Email Marketing: Esta estrategia se centra en el uso efectivo del correo electrónico como medio para comunicarse con los clientes.
+Permite enviar mensajes personalizados, campañas específicas y promociones adaptadas a las necesidades individuales de cada cliente.
+El email marketing es una herramienta clave para establecer y mantener una conexión directa y personalizada.
+
+Mobile Marketing: El mobile marketing se enfoca en llegar a los usuarios a través de dispositivos móviles, como teléfonos inteligentes y tabletas.
+Con el auge de los dispositivos móviles, esta herramienta se vuelve esencial para llegar a los clientes dondequiera que estén.
+Incluye estrategias como mensajes de texto, aplicaciones móviles y publicidad específica para dispositivos móviles.
+
+Social Marketing: Las redes sociales desempeñan un papel fundamental en el big data marketing.
+Esta herramienta implica la utilización de plataformas como Facebook, Twitter, Instagram, entre otras, para interactuar con los clientes, conocer sus opiniones y preferencias, y promover productos o servicios de manera personalizada.
+El social marketing es valioso para construir relaciones sólidas con los clientes y generar interacción directa.
+
+Explorar y comprender estas prácticas es esencial para utilizarlas efectivamente en un enfoque de big data marketing, asegurando así una comunicación personalizada y adaptada a las necesidades individuales de cada cliente.
+
+\subsection{Email marketing}
+
+El email marketing, una herramienta de marketing directo que utiliza el correo electrónico para realizar campañas publicitarias y promocionales, ha ganado gran popularidad con el auge de internet.
+Se ha convertido en una alternativa más eficiente y económica que las promociones enviadas por correo postal, permitiendo una entrega inmediata y la medición de resultados.
+
+Aunque pueda parecer simple, una estrategia de email marketing requiere una planificación cuidadosa y el seguimiento de varios procesos en tres fases clave:
+
+Planificación:
+
+Fijar objetivos a corto, medio y largo plazo.
+Objetivos comunes incluyen branding, transacciones, captación y cualificación.
+Ejecución:
+
+Definir el público objetivo.
+Determinar el mensaje y la acción deseada del consumidor.
+Establecer el momento de envío.
+Medición de Resultados:
+
+Evaluar métricas como entregabilidad, apertura, click y conversión.
+Obtener conocimiento adicional sobre los clientes para futuras campañas.
+En el contexto del big data y el customer analytics, el email marketing se beneficia de:
+
+Target de envío:
+Utilizar datos del big data para conocer mejor a los clientes y definir un target de envío específico.
+Evitar envíos masivos y enviar correos solo a clientes potencialmente interesados.
+Personalizar estrategias con micro-segmentos, permitiendo una comunicación más detallada y dirigida.
+La integración de big data y customer analytics ha llevado a una estrategia más refinada de email marketing, asegurando que los mensajes se envíen a las personas adecuadas y contribuyendo a relaciones duraderas con los clientes.
+
+\subsection{Personalización del contenido}
+
+La micro-segmentación facilita la presentación de contenidos altamente diferenciados para cada tipo de cliente, permitiendo una comunicación más precisa y efectiva.
+Aunque esta estrategia puede excluir a una parte de la base de datos, se contrarresta con la personalización del contenido.
+
+Personalización del Contenido:
+
+Ventaja: Contenidos exclusivos para cada cliente, percibidos como más valiosos, mejorando resultados a corto, medio y largo plazo.
+Desventaja: Mayor personalización implica más recursos.
+La decisión sobre el nivel de personalización debe equilibrar los beneficios y costos asociados.
+La personalización adicional debe ser económicamente rentable, donde los beneficios superen los costos, llegando a un compromiso entre la calidad del mensaje y los recursos invertidos.
+
+\subsection{Momento y frecuencia de envío}
+
+El éxito de la campaña depende del momento en que el email aparece en la bandeja de entrada del receptor.
+Una cuarta parte de los emails se abre en la primera hora de recepción, y esta tasa disminuye significativamente después de 24 horas.
+Principios Básicos para el Envío:
+
+Evitar envíos durante la franja nocturna.
+Evitar los martes para evitar competir con correos acumulados durante el fin de semana.
+Estadísticamente, el mejor día y franja horaria son jueves de 8 a 9 de la mañana, seguido de miércoles de 8 a 10.
+Adaptación con Big Data:
+
+Gracias al análisis detallado del big data, se puede definir el mejor momento de envío.
+Pautas para adaptar el momento de envío:
+Conocer el mejor día de envío por usuario.
+Crear una variable de frecuencia de apertura a nivel de usuario para determinar la frecuencia de envío por usuario.
+
+\subsection{Puesta en escena}
+
+Tipos de Emails y Captación de Datos en Email Marketing:
+
+Emails Transaccionales:
+
+Orientados a generar ventas.
+Utilización del big data para determinar productos de interés, preferencias de ofertas y estrategias para aumentar el ticket medio.
+Enfocados en aumentar las ventas.
+Emails de Marca:
+
+Personalizados para comunicar lo que interesa al cliente sobre la marca.
+Contribuyen a construir una imagen positiva y reputación.
+Captación de Datos:
+
+Además de vender y mejorar la imagen de marca, los emails deben recoger datos útiles para futuras campañas.
+Se busca identificar los posibles clics del receptor para obtener información valiosa.
+Categorizar los clics por variables (por ejemplo, categoría o tipo de acción) facilita el análisis.
+Ejemplo: Una marca de ropa envía un correo con secciones de ropa para mujer, hombre, hogar y niños durante rebajas.
+Analizar los clics permite determinar las preferencias específicas del cliente y personalizar futuros correos.
+
+%%  Inserting image 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\linewidth]{./images/ex-mailmkt.jpg}
+  \caption{Ej. Email marketing}
+  \label{fig:etiqueta}
+\end{figure}
+%% End insert image
+
+\subsection{Mobile marketing}
+
+Mobile Marketing en Big Data:
+
+El Mobile Marketing en el contexto del Big Data se define como la práctica que permite a las empresas interactuar y conectar de manera relevante con los clientes a través de dispositivos móviles, que incluyen teléfonos móviles y tabletas.
+Distinguiéndose del marketing tradicional, el mobile marketing aprovecha diversas capacidades de los dispositivos móviles:
+
+Localización: Utiliza sistemas GPS para conocer la ubicación geográfica de los clientes, permitiendo la presentación de anuncios, contenido relevante, o promociones específicas según la localización.
+
+Proximidad: Los dispositivos móviles ofrecen accesibilidad con un solo clic a través de diversas tecnologías.
+
+Aplicaciones: El uso de aplicaciones permite mostrar catálogos de productos y servicios, facilitando a los clientes adquirir productos desde cualquier lugar.
+
+Anuncios para Móviles: Originados en banners tradicionales, estos anuncios se presentan al usuario mientras consume contenido móvil y están relacionados con sus preferencias.
+Pueden ser formatos de vídeo interactivos.
+
+SMS y MMS: Aunque ha perdido popularidad, los mensajes de texto ofrecen una manera rápida y directa de transmitir información a los clientes, a menudo utilizados como complemento para campañas multicanal.
+
+Cupones Electrónicos: Cupones de descuento electrónicos atraen clientes a las tiendas y eliminan la necesidad de llevar cupones físicos.
+
+Búsquedas Móviles: Los resultados de búsqueda móvil están adaptados a la experiencia de uso móvil, requiriendo estrategias de SEO móvil para la promoción de contenidos dirigidos a usuarios móviles.
+
+En el ámbito del Big Data, el Mobile Marketing se ve potenciado, ya que los dispositivos móviles generan una gran cantidad de datos, proporcionando información valiosa sobre los usuarios.
+Estos dispositivos permiten a los usuarios permanecer conectados durante todo el día, lo que facilita la recopilación de datos sobre sus comportamientos y preferencias.
+
+La aplicación del Big Data al Mobile Marketing permite la comunicación no solo con el contenido más adecuado sino también en el momento oportuno para cada cliente.
+La geolocalización y el análisis de datos en tiempo real posibilitan campañas de marketing cercanas y específicas para el público objetivo.
+
+Las principales campañas de Mobile Marketing en Big Data incluyen aplicaciones, comunicación móvil y mensajes personalizados.
+
+Ejemplo de Mobile Marketing: Starbucks y Spotify
+
+Un destacado ejemplo de Mobile Marketing se materializa a través de la colaboración entre Starbucks y Spotify.
+En esta campaña, los clientes registrados en la aplicación de Starbucks tienen la oportunidad de acceder a listas de música diseñadas exclusivamente para ellos.
+
+Esta iniciativa ilustra cómo las empresas pueden aprovechar las capacidades de las aplicaciones móviles para proporcionar experiencias personalizadas a sus clientes.
+Al vincular la aplicación de Starbucks con Spotify, se crea un canal único para ofrecer contenido musical que se adapta a los gustos individuales de los usuarios.
+
+El valor añadido radica en la exclusividad de estas listas de reproducción, lo que fomenta la lealtad del cliente.
+Al proporcionar algo único y personalizado a través de la aplicación móvil, Starbucks no solo mejora la experiencia del cliente sino que también establece una conexión más profunda con su base de consumidores.
+
+Este ejemplo destaca cómo el Mobile Marketing no se trata solo de promociones y anuncios, sino de ofrecer contenido relevante y personalizado a través de dispositivos móviles, aprovechando las capacidades del análisis de datos para entender y satisfacer las preferencias individuales de los usuarios.
+La colaboración con plataformas populares, como Spotify en este caso, también puede aumentar la visibilidad y el atractivo de la campaña.
+
+%%  Inserting image 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\linewidth]{./images/spostar.jpg}
+  \caption{Ej. Campaña Spotify / Starbucks}
+  \label{fig:etiqueta}
+\end{figure}
+%% End insert image
+
+\subsection{Social marketing}
+
+El Social Marketing, como parte integral del Big Data Marketing, se define como la técnica que utiliza los medios sociales e influenciadores para abordar las necesidades de marketing y negocio de una empresa.
+Dentro de esta modalidad, se pueden llevar a cabo acciones tradicionales como patrocinios y anuncios sociales, así como estrategias más innovadoras centradas en el marketing de influencias y acciones en redes sociales y blogs de la empresa.
+
+En la relación con el Big Data, los medios sociales son una fuente valiosa de datos no estructurados sobre los clientes, contribuyendo significativamente al diseño personalizado de estrategias de marketing.
+Los medios sociales no solo ofrecen un canal para la promoción, sino que también proporcionan información detallada sobre las preferencias, comportamientos y opiniones de los usuarios.
+
+Al explorar los medios sociales más populares, se destaca la importancia del contenido generado por los usuarios.
+Analizar las publicaciones, interacciones y la audiencia de los usuarios en estos medios ofrece una visión profunda del comportamiento del consumidor.
+Además, aspectos como las visitas, likes, comentarios, comparticiones y el número de seguidores de los usuarios amplían la comprensión del perfil del cliente.
+
+Blogs y Redes Sociales:
+
+Blogs: Estos siguen siendo una plataforma relevante para el Social Marketing.
+La creación de contenido atractivo y la participación en la blogosfera pueden proporcionar datos valiosos sobre las preferencias de los consumidores.
+
+Redes Sociales: Son un componente esencial, con plataformas como Facebook, Twitter e Instagram que dominan el panorama.
+La interacción social, los comentarios, los "me gusta" y las actividades compartidas en estas redes ofrecen una ventana única a la psicografía del cliente.
+
+En este entorno, es crucial no solo considerar el contenido generado, sino también analizar las interacciones y el alcance.
+La cantidad de datos generados en los medios sociales requiere estrategias efectivas de análisis de datos para extraer información significativa y aplicarla en campañas de marketing personalizadas.
+
+En resumen, el Social Marketing en el contexto del Big Data se trata de entender y utilizar la información generada en los medios sociales para construir relaciones más profundas con los clientes y adaptar las estrategias de marketing a sus preferencias y comportamientos específicos.
+
+% BLOGS pp. 33
+
+%%%%% Commented  -----------------------------
+\begin{comment}
+%%%%% Commented  -----------------------------
+
+\section{Introduction}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{Doc2023}.
 
@@
 
-To discern the effectiveness of code challenges and explore alternative assessment methods, a thorough analysis of existing literature, industry practices, and case studies was conducted.
-The goal was to elucidate the correlation between success in code \section{Redefining Engineering Excellence}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{dweck2006mindset,duckworth2016grit}.
 
-Engineering is an amalgamation of cognitive agility, problem-solving acumen, and interpersonal skills.
-An engineer's worth lies not in the mastery of a predefined set of skills, but in their ability to learn, collaborate, and innovate.
-By championing these traits, organizations foster a culture of resilience and evolution.
+\subsection{Sample Sites \& Processing}
 
-\section{Promoting a Paradigm Shift}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+$\text{Turnover}$ Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+ $\text{Challenges}$ Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{tetlock2007illusion}.
+:
 
-A transformative shift in assessment paradigms is overdue.
-Code challenges, while valuable, should be complemented with broader evaluations.
-Organizations stand to gain by emphasizing holistic attributes such as adaptability, critical thinking, and collaboration.
-The engineering landscape beckons for engineers who can not only code, but also envision, communicate, and innovate.
+\begin{equation}
+    \text{Turnover} = k \times \text{Challenges}^2
+\end{equation}
+
+Where $k$ iLorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+%%  Inserting image 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\linewidth]{./images/ex.jpg}
+  \caption{Lorem ipsum dolor sit amet, consectetur a}
+  \label{fig:etiqueta}
+\end{figure}
+%% End insert image
+
+\section{Conclusion}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.
+
+%------------------------------------------------
+
+%%%%% Commented  -----------------------------
+\end{comment}
+%%%%% Commented  -----------------------------
 
@@
 
-This paper challenges the prevailing narrative that engineers must possess exhaustive knowledge and underscores the limitations of code challenges.
-It advocates for an inclusive assessment approach that values problem-solving, adaptability, and teamwork.
-Engineering is an ever-evolving discipline, necessitating engineers who can navigate complexities and contribute meaningfully to the collaborative process \cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
@@
 \end{document}
+
+%----------------------------------------------------------------------------------------

### 2023-10-20 CS2324W-OSP-MCS/Reassessing the Efficacy of Code Challenges/paper.tex
Updates

@@
-% Code Challenge Assessment: A Holistic Approach to Evaluating Engineering Talent
+% --------------------------------------------
 % Author: Felipe Alfonso González
-% License: MIT
+% https://github.com/felipealfonsog
+% f.alfonso@res-ear.ch / felipe.alfonso.glz@gmail.com
+% License: MIT / BSD 3-clause License
 %
@@
 \setcounter{page}{1} % The page number of the first page, set this to a higher number if the article is to be part of an issue or larger work
-
-\usepackage{amsmath} % Add this line to load the amsmath package
 
@@
 }
+
+%----------------------------------------------------------------------------------------
 
@@
 
+%%%%% Commented  -----------------------------
+\begin{comment}
+%%%%% Commented  -----------------------------
+
+\section{Introduction}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{Doc2023}.
+
+%------------------------------------------------
+
+\section{Methodologies}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{dweck2006mindset,duckworth2016grit}.
+
+\subsection{Sample Sites \& Processing}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+$\text{Turnover}$ Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+ $\text{Challenges}$ Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{tetlock2007illusion}.
+:
+
+\begin{equation}
+    \text{Turnover} = k \times \text{Challenges}^2
+\end{equation}
+
+Where $k$ iLorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+%%  Inserting image 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\linewidth]{./images/ex.jpg}
+  \caption{Lorem ipsum dolor sit amet, consectetur a}
+  \label{fig:etiqueta}
+\end{figure}
+%% End insert image
+
+\section{Conclusion}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.
+
+%------------------------------------------------
+
+%%%%% Commented  -----------------------------
+\end{comment}
+%%%%% Commented  -----------------------------
+
 %------------------------------------------------

### 2023-10-20 CS2324W-OSP-MCS/Reevaluando la Eficacia de los Desafíos de Código/paper.tex
Updates

@@
-% ----------------------------------------------
+% --------------------------------------------
 % Author: Felipe Alfonso González
-% License: MIT
+% https://github.com/felipealfonsog
+% f.alfonso@res-ear.ch / felipe.alfonso.glz@gmail.com
+% License: MIT / BSD 3-clause License
 %

### 2023-11-01 CS2325W-MCS/BigData-paper-ENEB/paper.tex
Updates - ENEB

@@
 
-\title{Big Data en Marketing}
+\title{Big Data Marketing}
 
@@
 
-% BLOGS pp. 33
+\subsection{Blogs \& RRSS}
+
+Los blogs son espacios donde empresas o individuos comparten regularmente textos o artículos sobre un tema específico.
+Aunque algunos consideran que los blogs son innecesarios debido al fuerte impacto de las redes sociales, en realidad, dependiendo del sector, pueden ser recursos valiosos.
+Los blogs permiten presentar a los usuarios contenido más elaborado que a menudo complementa lo compartido en redes sociales.
+Aunque plataformas como Facebook o Twitter son útiles para la presentación general, los blogs ofrecen la oportunidad de profundizar en temas específicos de manera más detallada y estructurada.
+
+La estrategia consiste en compartir nuestras novedades a través de las redes sociales, proporcionando enlaces a nuestra web o blog para que los usuarios obtengan información más detallada.
+Por ejemplo, al entrevistar a alguien influyente en nuestro sector, podemos utilizar Instagram para compartir una foto del entrevistado y dirigir a los seguidores al contenido completo de la entrevista en nuestro blog.
+Esto asegura que la información se presente de manera atractiva en las redes sociales, pero se redirija a plataformas más amplias y detalladas para una experiencia informativa completa.
+
+\subsection{Las características principales de los blogs incluyen:}
+
+Interactividad Dinámica:
+
+Espacios dinámicos que permiten a los autores interactuar con los usuarios a través de comentarios.
+Comunicación Personal e Informal:
+
+A diferencia de los sitios web, los blogs presentan una comunicación más personal e informal, evitando el lenguaje formal y corporativo.
+Esto crea un medio más cercano para establecer comunicación directa e informal, manteniendo rigor y educación.
+Organización Mediante Posts:
+
+Los contenidos se publican a través de posts, clasificados por fecha, categoría y etiquetas para una rápida identificación.
+Accesibilidad mediante RSS:
+
+Los blogs pueden ser leídos mediante agregadores de RSS, facilitando la difusión y notificación a los lectores sobre nuevos contenidos publicados.
+
+El microblogging es una forma de publicar contenido mediante mensajes breves, destacando por su simplicidad en la comunicación.
+Este formato permite transmitir mensajes de manera clara, rápida y precisa debido a las limitaciones de espacio.
+La naturaleza breve de los mensajes en el microblogging requiere originalidad por parte del emisor para comunicar de manera efectiva en un número reducido de caracteres.
+Twitter es la plataforma que mejor representa esta forma de comunicación.
+
+% pp 35
+\subsection{Redes Sociales: Un análisis mayor}
+
+Las redes sociales son estructuras sociales en las que individuos u organizaciones se relacionan según diversos criterios, como la amistad, el parentesco o relaciones profesionales.
+En el contexto de Internet, se dividen en tres niveles:
+
+Redes Sociales Genéricas: Son las más comunes, como Facebook, Instagram o Twitter.
+Su función principal es establecer relaciones entre individuos, ya sea que se conozcan previamente o no.
+Además, son fuentes de información tanto a nivel personal como público.
+
+Redes Sociales Profesionales: Están orientadas a relaciones profesionales, permitiendo a sus miembros conectar con colegas de trabajo o buscar empleo. Ejemplos notables incluyen Xing o Linkedin.
+
+Redes Sociales Verticales o Temáticas: Están centradas en un tema específico, conectando a usuarios interesados en dicho tema.
+Ejemplos son Flickr, donde se comparten fotografías de carácter profesional, y Dribble, una red exclusiva para diseñadores gráficos que muestran sus trabajos.
+
+\subsection{Las redes sociales más populares como herramientas valiosas para obtener datos sobre clientes:}
+
+Facebook: Destacado por su popularidad, permite compartir varios tipos de contenido.
+Se diferencia entre perfiles personales y páginas profesionales.
+Para empresas, es crucial prestar atención a comentarios, likes, dislikes y contenido compartido.
+
+Twitter: Enfocado en microblogging, es esencial transmitir mensajes atractivos en 280 caracteres.
+Utilizar hashtags relacionados con la empresa facilita el acceso a comentarios y opiniones sobre la marca.
+
+Instagram: Ideal para contenido visual, permite compartir fotos y videos, incluyendo publicaciones temporales.
+La red es dinámica y, a través de hashtags, brinda acceso a publicaciones relacionadas con la empresa.
+
+Linkedin: Enfocada en el ámbito profesional, es útil para buscar oportunidades de negocio, establecer alianzas estratégicas y generar una red de contactos. Es eficaz para generar clientes potenciales.
+
+Google+: Aunque menos utilizada, su pertenencia a Google proporciona ventajas en términos de posicionamiento y contacto con usuarios de la plataforma de búsqueda.
+
+Aspectos importantes para obtener datos e información sobre clientes incluyen:
+
+Gestión de Comentarios: Considerar todos los comentarios, tanto positivos como negativos, para evaluar la satisfacción del cliente y mejorar en áreas necesarias.
+
+Comunicación Activa: Mantener una comunicación activa con los clientes a través de comentarios y mensajes privados para construir relaciones sólidas.
+
+Geolocalización: Utilizar la geolocalización para conocer la ubicación física de los clientes y ofrecer servicios personalizados.
+
+Análisis de Contenido: No solo analizar los comentarios sobre la propia marca, sino también estudiar lo que los usuarios comparten sobre otras marcas para obtener una visión más amplia del mercado.
+
+El marketing social basado en datos puede ser de diferentes tipos, como contenido basado en la reacción de los usuarios, publicidad según el perfil del usuario y publicaciones basadas en algoritmos.
+
+Se mencionan herramientas específicas como Facebook Custom Audience y Twitter Custom Audiences, que utilizan datos para acciones de marketing dirigidas, permitiendo a las empresas adaptar mensajes según los criterios definidos por los usuarios.
+Estas herramientas aprovechan el conocimiento de la base de datos de clientes para la generación de publicidad en redes sociales.
+
+\subsection{Marketing automation}
+
+El marketing automation se refiere a herramientas que automatizan tareas de marketing, especialmente las comunicaciones con usuarios derivadas de acciones específicas.
+Aunque no todas las acciones comerciales pueden automatizarse, algunas tareas incluyen enviar mensajes de agradecimiento tras una compra, adjuntando información relevante y recomendaciones de productos relacionados.
+También, se puede personalizar la página de entrada de un e-commerce según la navegación y compras de los usuarios, adaptando la experiencia a sus intereses y preferencias.
+La automatización se aplica principalmente en canales digitales, como el correo electrónico, mensajes, notificaciones en aplicaciones, recomendaciones de compras y personalización de la pantalla de inicio.
+
+La implementación de acciones automatizadas en marketing ofrece diversas ventajas y capacidades, entre las que se incluyen:
+
+Desarrollo de Procesos Complejos: Permite la ejecución de planes de comunicación complejos, como los dirigidos a clientes que visitan la tienda por primera vez.
+
+Eficiencia y Reducción de Costos: Aumenta la eficiencia al automatizar tareas que de otro modo se realizarían manualmente, lo que también conlleva a una reducción de costos.
+
+Educación y Maduración en la Base de Datos: Facilita la implementación de procesos para educar y madurar la información en la base de datos de clientes.
+
+Identificación de Registros Preparados para Ofertas Comerciales: Permite detectar qué registros de clientes están más preparados para recibir ofertas comerciales.
+
+Además, la adopción del marketing automatizado presenta ventajas adicionales:
+
+Eficiencia: Automatiza tareas manuales, optimizando recursos y permitiendo que una persona pueda contactar con una gran cantidad de clientes potenciales y actuales.
+
+Ahorro de Tiempo: Incrementa la productividad al abordar acciones rutinarias, liberando tiempo para otras tareas.
+
+CRM (Customer Relationship Management): Facilita el acceso a una base de datos completa con información sobre clientes actuales y potenciales.
+
+Gestión Comercial a Largo Plazo: Permite construir relaciones a medio y largo plazo con clientes potenciales que aún no están listos para adquirir productos o servicios.
+
+Efectividad en Marketing Digital: Contribuye a obtener un mayor retorno de la inversión en marketing digital al medir y utilizar técnicas con mejores ratios de conversión.
+
+Monitorización: Realiza un seguimiento y análisis detallado de cada acción en cada segmento de la base de datos.
+
+Se destaca la importancia de no dejar que la automatización deshumanice la marca, ya que esto podría tener un efecto contrario al deseado si se confía en que la herramienta de software se encargue de todo.
+
+Para una exitosa implementación de marketing automation, se deben seguir los siguientes pasos:
+
+Definir la Estrategia:
+
+Establecer objetivos y determinar qué acciones se automatizarán.
+Búsqueda de la Herramienta Adecuada:
+
+Considerar funcionalidades esenciales y evitar pagar por funciones no utilizadas.
+Evaluar la facilidad de uso, atención al cliente, precio y sistema de precios.
+Utilización de la Herramienta:
+
+Comenzar con acciones y tareas simples para comprender la herramienta.
+Avanzar hacia acciones más complejas una vez que se haya comprendido su funcionamiento.
+Algunas herramientas de automatización incluyen Hubspot, que proporciona un conjunto de herramientas conectadas para atraer visitas, gestionar relaciones comerciales y realizar un seguimiento en tiempo real.
+Infusionsoft es utilizada por pequeñas y medianas empresas, ofreciendo un software fácil de usar para estrategias de marketing y ventas.
+Otras herramientas son Padot, Marketo y Eloqua.
+
+Para la implantación del proceso de automatización, se sugieren los siguientes pasos:
+
+Definir los Procesos a Automatizar:
+
+Identificar momentos de interacción entre la empresa y los usuarios, como los procesos de compra en un comercio online.
+Realizar un Diagrama de Procesos:
+
+Diseñar un diagrama para cada proceso, detallando las acciones para cada resultado posible.
+Detallar Acciones en Cada Proceso:
+
+Especificar cada acción en los diferentes procesos, como introducir el correo electrónico, registrarse, redirección a la página de bienvenida, entre otras.
+Estos pasos permiten tener una comprensión clara de los procesos e interacciones con los clientes, facilitando la automatización de tareas específicas.
+Por ejemplo, a partir de los datos de compra, se puede automatizar el envío de una nota de agradecimiento y un código de descuento a clientes que hayan estado registrados por un año y hayan comprado al menos dos veces.
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+En resumen, la implementación efectiva de marketing automation ofrece a las empresas la capacidad de desarrollar procesos complejos, aumentar la eficiencia y reducir costos.
+La elección de la herramienta adecuada, considerando funcionalidades, usabilidad y atención al cliente, es esencial.
+Ventajas adicionales incluyen la optimización de recursos, el ahorro de tiempo, la gestión efectiva de relaciones con el cliente, y una mayor efectividad en el marketing digital.
+Es crucial evitar la deshumanización de la marca al utilizar estas herramientas y seguir pasos claros, como definir estrategias y procesos, para una implementación exitosa del marketing automation.
 

### 2023-11-01 CS23-00W-LaTEX-IACC-UMayor-Projects/Administracion_de_bases_de_datos_IACC_2020_2021-Latex/main.tex
Updates

\documentclass{article}
\usepackage{amsmath}
\usepackage{color,pxfonts,fix-cm}
\usepackage{latexsym}
\usepackage[mathletters]{ucs}
\DeclareUnicodeCharacter{8211}{\textendash}
\DeclareUnicodeCharacter{34}{\textquotedbl}
\DeclareUnicodeCharacter{46}{\textperiodcentered}
\DeclareUnicodeCharacter{8220}{\textquotedblleft}
\DeclareUnicodeCharacter{58}{$\colon$}
\DeclareUnicodeCharacter{8221}{\textquotedblright}
\DeclareUnicodeCharacter{8226}{$\bullet$}
\DeclareUnicodeCharacter{8216}{\textquoteleft}
\DeclareUnicodeCharacter{8230}{$\ldots$}
\DeclareUnicodeCharacter{167}{$\S$}
\DeclareUnicodeCharacter{8217}{\textquoteright}
\DeclareUnicodeCharacter{62}{\textgreater}
\DeclareUnicodeCharacter{32}{$\ $}
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage{pict2e}
\usepackage{wasysym}
\usepackage[english]{babel}
\usepackage{tikz}
\pagestyle{empty}
\usepackage[margin=0in,paperwidth=595pt,paperheight=841pt]{geometry}
\begin{document}
\definecolor{color_29791}{rgb}{0,0,0}
\definecolor{color_30046}{rgb}{0,0,1}
\definecolor{color_30045}{rgb}{0,0,0.996078}
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(38.82078,-50.06079){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}In: Modelamiento de base de datos (1)}
\put(193.3106,-46.9408){\fontsize{6}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}iii}
\put(198.27,-50.06079){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}, Continuidad en Ingeniería en Informática, IACC Chile., 2020-2021 }
\put(33.47067,-61.58081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-73.58081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(134.0845,-93.02081){\fontsize{13.92}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ADMBD1302-608-2020 – Modelamiento de base de }
\put(253.3603,-109.3408){\fontsize{13.92}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}datos (1)–}
\put(312.1106,-104.7808){\fontsize{8.4}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}iii}
\put(319.2811,-109.3408){\fontsize{13.92}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-122.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(231.2812,-133.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Felipe Alfonso González L.  }
\put(153.882,-148.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Continuidad en Ingeniería en Informática, IACC Chile., 2020-2021 }
\put(286.8208,-160.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(229.8847,-171.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Instituto Superior de Artes y }
\put(234.3662,-183.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Ciencias de la Comunicación, }
\put(282.125,-195.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}IACC }
\put(222.1088,-206.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Av. Salvador 1318, Metro Santa }
\put(234.503,-218.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Isabel, Providencia, Santiago. }
\put(274.7383,-230.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Chile. }
\put(286.8208,-241.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(107.8403,-253.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}f.alfonso@res-ear.ch – felipe.alfonso.glz@gmail.com – https://twitter.com/felipealfonsog  }
\put(112.1933,-265.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}https://glzengrg.com - https://freeshell.de/felipe - https://linkedin.com/in/felipealfonsog  }
\put(170.8208,-277.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(170.8208,-288.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-300.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(38.77074,-316.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}         Introducción al periodo }
\put(33.47067,-328.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}El uso de bases de datos como medios de }
\put(33.47067,-340.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}almacenamiento de información ha tenido }
\put(33.47067,-352.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}un incesante desarrollo durante las últimas }
\put(33.47067,-363.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}dos décadas. Dada la relevancia de la }
\put(33.47067,-375.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}información para la sociedad moderna, }
\put(33.47067,-387.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}garantizar la seguridad y el buen manejo de }
\put(33.47067,-399.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la misma se ha convertido en un aspecto de }
\put(33.47067,-411.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}vital interés para la sociedad moderna. }
\put(33.47067,-423.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}En un principio, las bases de datos fueron }
\put(33.47067,-434.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}utilizadas para almacenar información en }
\put(33.47067,-446.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}formato digital de una manera muy }
\put(33.47067,-458.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}conveniente, facilitando su recuperación y }
\put(33.47067,-470.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}manipulación para generar reportes basados }
\put(33.47067,-481.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}en los datos almacenados. Esta tendencia fue }
\put(33.47067,-493.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}incrementándose a una velocidad increíble, }
\put(33.47067,-505.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}lo cual hizo necesario desarrollar nuevos }
\put(33.47067,-517.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}métodos y técnicas para manipular dichos }
\put(33.47067,-529.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}datos. La teoría de base datos toma la }
\put(33.47067,-540.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}información de las organizaciones y la }
\put(33.47067,-552.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}almacena para que posteriormente los }
\put(33.47067,-564.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}usuarios de la misma puedan recuperarla y }
\put(33.47067,-576.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}procesarla de una manera por demás fácil e }
\put(33.47067,-588.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}intuitiva. }
\put(38.82078,-601.8208){\fontsize{12}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}1.   }
\put(33.47067,-613.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}La importancia de manejar un DBMS, su arquitectura e }
\put(33.47067,-624.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}instalación es fundamental para manejar y proteger }
\put(33.47067,-636.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}información. Para esto, existen mecanismos, que }
\put(33.47067,-647.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}posibilitaran el manejo de esta de una manera rápida y }
\put(33.47067,-659.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}expedita. Es de gran importancia y que se imponen en estos }
\put(33.47067,-671.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}procesos y operaciones son estandarizar las bases de datos }
\put(33.47067,-682.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}como por ej. Un modelo relacional de BBDD. El DBMS }
\put(33.47067,-694.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ofrecerá a la compañía la posibilidad de almacenar y }
\put(33.47067,-705.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}manejar con alto grado de eficiencia, abstracción, }
\put(33.47067,-717.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}flexibilidad e incluso con facilidad, bases de datos.  }
\put(33.47067,-728.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Existen componentes fundamentales para instalar y }
\put(33.47067,-740.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}configurar un DBMS, aquellos componentes son: }
\put(33.47067,-751.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(51.47067,-763.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}1- Datos }
\put(51.47067,-775.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}2- Procedimientos }
\put(51.47067,-786.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}3- Hardware }
\put(51.47067,-798.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}4- Software }
\put(51.47067,-809.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}5- Red }
\put(312.5906,-312.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}6- Personas }
\put(312.5906,-323.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}7- Servidores de BB.DD. }
\put(294.5906,-335.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-346.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Es también importante considerar que los componentes de }
\put(294.5906,-358.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}un DBMS, son: }
\put(294.5906,-369.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(312.5906,-381.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}1- Hardware: dispositivos físicos, servidores o }
\put(330.5906,-393.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}estaciones. }
\put(312.5906,-404.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}2- Software: Es si mismo el DBMS, y es importante a }
\put(330.5906,-416.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}considerar el O.S. a operar, las aplicaciones de }
\put(330.5906,-427.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}programas, y programas utilitarios. }
\put(312.5906,-439.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}3- Usuarios: Son usuarios de un DBMS, como por ej.; }
\put(330.5906,-450.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Administrador del sistema, Administrador de }
\put(330.5906,-462.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}BB.DD., y diseñadores de BB.DD. }
\put(312.5906,-473.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}4- Analistas y Desarrolladores }
\put(312.5906,-485.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}5- Procedimientos: Son reglas y normas a controlar }
\put(330.5906,-496.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para mantener de manera integra los datos. }
\put(312.5906,-508.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}6- Datos: Son el componente primario para el }
\put(330.5906,-519.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}desarrollo de información relevante. }
\put(299.9407,-531.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-543.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Es muy importante considerar como fundamentales en toda }
\put(294.5906,-554.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}esta operación, con profesionales adecuados, que tal como se }
\put(294.5906,-566.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mencionan en el punto 4 son los analistas y desarrolladores, }
\put(294.5906,-577.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}que tienen altas habilidades para el desarrollo de software, y }
\put(294.5906,-589.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}son fundamentales para la toma de decisiones.  }
\put(294.5906,-600.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-612.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}El proceso de instalación de un DBMS es un tema complejo }
\put(294.5906,-623.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}y delicado. Estos procesos deben considerar: }
\put(294.5906,-635.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(312.5906,-647.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}1- Lectura del manual de instalación. }
\put(312.5906,-658.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}2- Requisitos sobre el O.S. }
\put(312.5906,-670.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}3- Considerar especificaciones técnicas como por ej., }
\put(330.5906,-681.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tipo de procesador a usar, etc. }
\put(312.5906,-693.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}4- Chequeo de cantidad optima de memoria RAM y }
\put(330.5906,-704.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}H.D. }
\put(299.9407,-716.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-727.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Ahora bien, para proceder con la configuración de un }
\put(294.5906,-739.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DBMS, se debe considerar y proceder con lo Sgte.: }
\put(294.5906,-751.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(312.5906,-763.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}1- Espacio de Almacenamiento: }
\put(330.5906,-774.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}a) Establecer cantidad de espacio físico idóneo }
\put(330.5906,-786.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}b) Establecer correcta configuración, para un }
\put(348.5906,-797.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}máximo desempeño. }
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(69.47067,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}c) Determinar nivel de exactitud en el diseño de }
\put(87.47067,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la BB.DD. }
\put(87.47067,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(51.47067,-81.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}2- Definición de espacio, se establece espacio }
\put(69.47067,-93.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}requerido para una tabla especifica. }
\put(38.82078,-105.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(51.47067,-117.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}3- Creación del espacio. Se crean 3 tablas, reservando }
\put(69.47067,-128.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}espacio para las Sgts.: }
\put(69.47067,-140.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}a) SYSCATSPACE: Almacena catalogo de }
\put(87.47067,-151.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tablas. }
\put(69.47067,-163.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}b) TEMPSPACE1: Almacena tablas temporales. }
\put(69.47067,-174.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}c) USERSPACE1: Almacena tablas, y }
\put(87.47067,-186.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}respectivos índices creados por los usuarios. }
\put(69.47067,-197.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estas tablas son generadas manera automática, }
\put(69.47067,-209.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}aunque de cualquier modo puede ser creadas por la }
\put(69.47067,-220.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}misma DBMS. }
\put(69.47067,-232.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}El espacio inicial de una BB.DD. Estará definido }
\put(69.47067,-243.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}por omisión o defecto.  }
\put(33.47067,-255.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(51.47067,-267.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}4- Configuración de memoria: Es vital, gran cantidad }
\put(69.47067,-279.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de memoria para almacenar información, procesos }
\put(69.47067,-290.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}y ejecutar operaciones. Para todo lo anterior se }
\put(69.47067,-302.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}requiere una memoria cache o temporal con el }
\put(69.47067,-313.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}objetivo de minimizar las operaciones de lectura y }
\put(69.47067,-325.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}escritura desde y hacia un H.D.; Desde y hacia se }
\put(69.47067,-336.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}debe definir un Buffer pool que elimina la }
\put(69.47067,-348.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}subsecuente lectura desde el medio físico mientras }
\put(69.47067,-359.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}se mantengan en el Buffer pool. }
\put(38.82078,-371.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(51.47067,-383.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}5- Memoria compartida, establece:  }
\put(69.47067,-394.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}a) La asignación y liberación de estas áreas de }
\put(87.47067,-406.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}memoria, que ocurren cuando es asignada a un }
\put(87.47067,-418.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}área particular, en un evento particular, por }
\put(87.47067,-429.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ej.: la conexión a una BB.DD. }
\put(69.47067,-441.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}b) Durante estos eventos la DBMS asignan }
\put(87.47067,-452.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}memoria de la siguiente manera: }
\put(87.47067,-464.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Al inicio del DBMS. }
\put(87.47067,-476.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Cuando una base de datos es conectada }
\put(105.4707,-488.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}por primera vez. }
\put(87.47067,-500.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Cuando una App. Se conecta a una }
\put(105.4707,-511.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}BB.DD. }
\put(105.4707,-523.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(51.47067,-535.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}6- Instancias Múltiples, son un conjunto completo de }
\put(69.47067,-547.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Apps necesarias para el correcto funcionamiento }
\put(69.47067,-558.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de una DBMS y se pueden instalar múltiples }
\put(69.47067,-570.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}instancias, como: }
\put(87.47067,-582.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Soporte para Apps: instancias para }
\put(105.4707,-593.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}diferentes organizaciones con diferentes }
\put(105.4707,-605.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}BB.DD. o una única instancia, bajo la }
\put(105.4707,-616.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mirada de un solo SysAdmin. }
\put(87.47067,-628.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Consolidación de los servidores, esta }
\put(105.4707,-640.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}característica permite reducir el numero }
\put(105.4707,-651.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de servidores a utilizar ya que no es }
\put(105.4707,-663.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}necesario un servidor para cada DBMS. }
\put(87.47067,-675.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Soporte Técnico, Son actividades a nivel }
\put(105.4707,-687.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}se software o hardware, por lo que es }
\put(105.4707,-698.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ideal un equipo técnico de experiencia o }
\put(105.4707,-710.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}capacitado.  }
\put(105.4707,-721.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-733.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-744.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-756.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-767.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-779.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-790.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}En las estructuras de datos se debe considerar importante el }
\put(33.47067,-802.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}control de archivos en dispositivos de almacenamiento, y su }
\put(33.47067,-813.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}gerencia es parte del Sysadmin junto al diseñador, esto }
\put(294.5906,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}permite el almacenamiento de estructuras de datos, como: }
\put(330.5906,-58.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- lógicas }
\put(330.5906,-70.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Y físicas. }
\put(348.5906,-82.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-93.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Las estructuras lógicas permiten independencia en los datos }
\put(294.5906,-105.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}definidos y en la forma física en como los datos son }
\put(294.5906,-117.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}definidos. Garantizando integridad y desempeño del DBMS, }
\put(294.5906,-128.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}considerando: }
\put(294.5906,-140.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}La extensión, que compone segmentos y se asignan bloques }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(294.5906pt, -140.7808pt) -- (345.9506pt, -140.7808pt)
 -- (345.9506pt, -140.7808pt)
 -- (345.9506pt, -141.2608pt)
 -- (345.9506pt, -141.2608pt)
 -- (294.5906pt, -141.2608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-151.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de espacio libres y continuos, y presentan estas ventajas: }
\put(330.5906,-163.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Se elimina intervención del Sysadmin. }
\put(330.5906,-175.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Se asegura que las Apps no dejaran de }
\put(348.5906,-187.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}funcionar. }
\put(294.5906,-198.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Segmento, la mayoría de los DBMS utiliza la técnica -}
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(294.5906pt, -199.3408pt) -- (334.6706pt, -199.3408pt)
 -- (334.6706pt, -199.3408pt)
 -- (334.6706pt, -199.8208pt)
 -- (334.6706pt, -199.8208pt)
 -- (294.5906pt, -199.8208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-210.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Mirror- y requieren ciertas características: }
\put(330.5906,-222.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Los segmentos deben almacenarse dentro del }
\put(348.5906,-233.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mismo sistema. }
\put(330.5906,-245.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Los segmentos copiados deben ser }
\put(348.5906,-257.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}almacenados en un Host diferente. }
\put(330.5906,-269.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Espacio de tablas. Es un grupo de particiones }
\put(348.5906,-280.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}en la BB.DD. Están compuestos por omisión }
\put(348.5906,-292.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}en la base de datos y son: }
\put(348.5906,-303.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(402.5906,-315.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}1) Syscatspace. }
\put(402.5906,-327.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}2) Estructuras de datos que }
\put(420.5906,-338.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}guarda el usuario. }
\put(402.5906,-350.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}3) El espacio de tablas }
\put(420.5906,-361.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}temporales. }
\put(294.5906,-373.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-384.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-396.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}En las operaciones sobre una BB.DD. implican las }
\put(294.5906,-407.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}siguientes consideraciones: }
\put(294.5906,-419.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Espacios, Los DBMS han experimentado un avance para }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(294.5906pt, -420.1408pt) -- (330.1106pt, -420.1408pt)
 -- (330.1106pt, -420.1408pt)
 -- (330.1106pt, -420.6208pt)
 -- (330.1106pt, -420.6208pt)
 -- (294.5906pt, -420.6208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-431.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}calcular el espacio de almacenamiento, estos trataran como }
\put(294.5906,-442.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}arreglo lineal de paginas contiguas de BB.DD., logrando }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(485.6306pt, -443.4208pt) -- (523.7906pt, -443.4208pt)
 -- (523.7906pt, -443.4208pt)
 -- (523.7906pt, -443.9008pt)
 -- (523.7906pt, -443.9008pt)
 -- (485.6306pt, -443.9008pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-454.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}usos eficientes de espacio de almacenamiento. Estos }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(294.5906pt, -454.9408pt) -- (480.1106pt, -454.9408pt)
 -- (480.1106pt, -454.9408pt)
 -- (480.1106pt, -455.4208pt)
 -- (480.1106pt, -455.4208pt)
 -- (294.5906pt, -455.4208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-465.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}espacios de datos serán recuperados para posterior }
\put(294.5906,-477.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}procesamiento. El DBMS dispone un área compartida de }
\put(294.5906,-488.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}gran capacidad de área de memoria (Shared Buffer Poll), }
\put(294.5906,-500.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}organizado como frames, leídos desde el dispositivo HD y }
\put(294.5906,-511.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cargados en el área común de memoria compartida. }
\put(294.5906,-523.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-535.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Espacios Privados, son BB.BB. denominadas -Formularios o }
\put(294.5906,-546.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}paneles de control-, su función es: }
\put(294.5906,-558.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(402.5906,-569.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}§ Guiar a usuarios a la }
\put(420.5906,-581.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}reorganización de funciones }
\put(420.5906,-592.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}especificas.  }
\put(402.5906,-604.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}§ Controlan el acceso a las }
\put(420.5906,-615.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tablas. }
\put(294.5906,-627.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-639.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Espacios para objetos, su estructura de datos mantiene –Los }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(294.5906pt, -639.7408pt) -- (381.2306pt, -639.7408pt)
 -- (381.2306pt, -639.7408pt)
 -- (381.2306pt, -640.2208pt)
 -- (381.2306pt, -640.2208pt)
 -- (294.5906pt, -640.2208pt) -- cycle
;
\filldraw[color_29791][nonzero rule]
(519.2306pt, -639.7408pt) -- (536.7506pt, -639.7408pt)
 -- (536.7506pt, -639.7408pt)
 -- (536.7506pt, -640.2208pt)
 -- (536.7506pt, -640.2208pt)
 -- (519.2306pt, -640.2208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-650.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}llamados archivos-, manejados por el Sysadmin. }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(294.5906pt, -651.2608pt) -- (367.0706pt, -651.2608pt)
 -- (367.0706pt, -651.2608pt)
 -- (367.0706pt, -651.7408pt)
 -- (367.0706pt, -651.7408pt)
 -- (294.5906pt, -651.7408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-662.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Una buena administración redundara en: }
\put(294.5906,-673.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(330.5906,-685.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Mejoras de funcionamiento. }
\put(330.5906,-697.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Reducción de tiempos de acceso. }
\put(330.5906,-709.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Almacenamiento persistente y robusto. }
\put(294.5906,-721.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-732.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Deben considerarse los llamados -Tipos-,  }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(431.8706pt, -733.5808pt) -- (454.4306pt, -733.5808pt)
 -- (454.4306pt, -733.5808pt)
 -- (454.4306pt, -734.0608pt)
 -- (454.4306pt, -734.0608pt)
 -- (431.8706pt, -734.0608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(366.5906,-744.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Clustered. }
\put(366.5906,-755.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Unique. }
\put(366.5906,-767.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Fulltext. }
\put(294.5906,-778.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-790.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Bitácoras, cambios que quedan registrados de forma serial, }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(294.5906pt, -791.1808pt) -- (332.2706pt, -791.1808pt)
 -- (332.2706pt, -791.1808pt)
 -- (332.2706pt, -791.6608pt)
 -- (332.2706pt, -791.6608pt)
 -- (294.5906pt, -791.6608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-801.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}así el DBMS, realiza operaciones como: }
\put(330.5906,-813.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Rollbacks. }
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(69.47067,-47.18079){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Ejecutar operaciones de control. }
\put(105.4707,-58.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Las bitácoras ocupan bastante }
\put(123.4707,-70.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}espacio debido a transacciones }
\put(123.4707,-81.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}denominadas -respaldos-. }
\put(105.4707,-93.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Los datos quedan grabados en la }
\put(123.4707,-105.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}bitácora. }
\put(105.4707,-116.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Es necesario tener mecanismos de }
\put(123.4707,-128.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}control llamados Checkpoints, que }
\put(123.4707,-139.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}son definidos por el DBA. }
\put(123.4707,-151.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-162.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Particiones, índices que pueden ser dividas en unidades }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(33.4706pt, -163.3408pt) -- (77.8706pt, -163.3408pt)
 -- (77.8706pt, -163.3408pt)
 -- (77.8706pt, -163.8208pt)
 -- (77.8706pt, -163.8208pt)
 -- (33.4706pt, -163.8208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-174.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}pequeñas. }
\put(33.47067,-185.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Partición de rango, son valores específicos: }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(33.4706pt, -186.3808pt) -- (106.6706pt, -186.3808pt)
 -- (106.6706pt, -186.3808pt)
 -- (106.6706pt, -186.8608pt)
 -- (106.6706pt, -186.8608pt)
 -- (33.4706pt, -186.8608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(69.47067,-197.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Campos claves }
\put(69.47067,-209.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Fechas especificas }
\put(87.47067,-221.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-232.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Partición de lista, lista de valores discretos para ser }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(33.4706pt, -233.6608pt) -- (100.6706pt, -233.6608pt)
 -- (100.6706pt, -233.6608pt)
 -- (100.6706pt, -234.1408pt)
 -- (100.6706pt, -234.1408pt)
 -- (33.4706pt, -234.1408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-244.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}utilizadas como claves.  }
\put(33.47067,-255.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Partición Hash, requiere claves Hash, distribuye filas de }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(33.4706pt, -256.7008pt) -- (92.5106pt, -256.7008pt)
 -- (92.5106pt, -256.7008pt)
 -- (92.5106pt, -257.1808pt)
 -- (92.5106pt, -257.1808pt)
 -- (33.4706pt, -257.1808pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-267.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}datos de manera uniforme. }
\put(33.47067,-279.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-290.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-302.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Configurar los diferentes modos de operaciones de un }
\put(33.47067,-313.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DBMS dan a entender que se debe tomar en cuenta las }
\put(33.47067,-325.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}siguientes operaciones: }
\put(69.47067,-337.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Operaciones esenciales de un DBMS, }
\put(87.47067,-348.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}proporcionan de manera segura, rápida y }
\put(87.47067,-360.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}confiable 3 aspectos importantes: }
\put(134.4707,-371.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}a) Procesos y actualizaciones }
\put(134.4707,-383.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}b) Recuperar datos y garantizar }
\put(152.4707,-394.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}su correcto uso.  }
\put(152.4707,-406.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(69.47067,-418.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- El DBMS, debe considerar: }
\put(105.4707,-430.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o El software requerido para manejo y }
\put(123.4707,-441.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}control de la BB.DD. }
\put(105.4707,-453.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Datos representativos del modelo }
\put(105.4707,-464.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Procedimientos. }
\put(105.4707,-476.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o El lenguaje de la BB.DD. }
\put(105.4707,-487.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Procesador de consultas. }
\put(105.4707,-499.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Gestión de tiempo de ejecución. }
\put(105.4707,-510.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Manejo de ejecución y autorización }
\put(123.4707,-522.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de usuarios y modificaciones. }
\put(105.4707,-534.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Ejecutar funciones para creación de }
\put(123.4707,-545.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tablas, respaldos y datos copiados. }
\put(105.4707,-557.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Modelo de BB.DD. garantiza }
\put(123.4707,-568.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}almacenamiento y recuperación de }
\put(123.4707,-580.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}datos. }
\put(105.4707,-591.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Generador de reportes, genera apoyo }
\put(123.4707,-603.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para la toma de decisiones: }
\put(141.4707,-615.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}§ Recuperación de Callback }
\put(141.4707,-626.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}§ Permanencia o Commit. }
\put(159.4707,-638.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-649.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}En las Operaciones de una BB.DD., destacan, Baja y Alta: }
\put(69.47067,-661.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Baja: Es un borrado de un registro. }
\put(69.47067,-673.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Alta: Gestiona nuevos registros a la BB.DD. }
\put(87.47067,-685.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-696.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Recovery, protección de los datos ante cualquier daño de su }
\put(33.47067,-708.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}integridad y confiabilidad. De otro modo sucede lo }
\put(33.47067,-719.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}siguiente: }
\put(105.4707,-731.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Daño en la superficie magnética del }
\put(123.4707,-742.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}disco }
\put(105.4707,-754.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Errores en las aplicaciones. }
\put(105.4707,-765.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Errores de usuarios. }
\put(123.4707,-777.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-789.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Es importante también el manejo de Índices: }
\put(105.4707,-800.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Un nuevo registro se realiza de }
\put(123.4707,-812.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}manera secuencial. }
\put(366.5906,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Búsquedas y recuperación contraria a }
\put(384.5906,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la inserción secuencial. }
\put(366.5906,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}o Técnica de Indexación, son procesos }
\put(384.5906,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de búsqueda y recuperación. }
\put(294.5906,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-104.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-116.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}A pesar de que ya existe una importancia bastante grande en }
\put(294.5906,-127.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}el significado de las BB.DD., estos deben estar en }
\put(294.5906,-139.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}condiciones que garanticen a un mayor nivel de seguridad y }
\put(294.5906,-150.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}respectivo control de datos. Esto de hecho ha generado una }
\put(294.5906,-162.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}total nueva disciplina, orientada a conceptos relacionados }
\put(294.5906,-173.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}con Cyber-Seguridad. }
\put(294.5906,-185.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-196.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Generalmente los datos se respaldan, con una copia de }
\put(294.5906,-208.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}archivos, luego se almacenan incluso en zonas geográficas }
\put(294.5906,-219.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}distintas como incluso otros países y así recuperarlos y }
\put(294.5906,-231.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}seguir con las operaciones de la organización. }
\put(294.5906,-243.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Se debe tomar en cuenta la seguridad y métodos para }
\put(294.5906,-254.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}proteger datos electrónicos, impresos e incluso manuales. Es }
\put(294.5906,-266.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}fundamental mantener la información privada, confidencial }
\put(294.5906,-277.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}y sensitiva protegida.  }
\put(294.5906,-289.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-300.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Existen varios métodos de monitoreo, de respaldo, }
\put(294.5906,-312.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}recuperación, formas de replicas, métodos de back ups, }
\put(294.5906,-323.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}comandos de respaldo y otros que podre abordar en la }
\put(294.5906,-335.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}respuesta numero dos y tres.  }
\put(294.5906,-346.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Una forma confiable e integra de mantenimiento y respaldo }
\put(294.5906,-358.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de la información es, la Permisología, esta operación es }
\put(294.5906,-369.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}fundamental y lograra un total y establecido tipo de acceso, }
\put(294.5906,-381.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}autenticaciones, privilegios asociados a usuarios o grupos. }
\put(294.5906,-393.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Dentro de los privilegios existen un ROLE, que podrá }
\put(294.5906,-404.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}empaquetar un grupo de funciones relacionadas con la }
\put(294.5906,-416.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}BB.DD. El establecimiento de ROLES hace mas fácil el }
\put(294.5906,-427.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}control de acceso a los datos, y, en consecuencia, una mejor }
\put(294.5906,-439.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}gerencia por parte del SYSAdmin. }
\put(294.5906,-450.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-462.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Hay operaciones de comandos que respaldan los datos. Me }
\put(294.5906,-473.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}referiré en detalle sobre este tópico en la respuesta numero }
\put(294.5906,-485.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}dos. }
\put(294.5906,-497.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-508.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Implementar métodos de respaldo y recuperación es }
\put(294.5906,-520.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}fundamental para salvaguardar la información. Con el }
\put(294.5906,-531.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tiempo se han generado variadas técnicas en esta área, como: }
\put(294.5906,-543.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(330.5906,-555.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Espejo (Mirroring), crear una copia completa }
\put(348.5906,-566.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}en un momento específico del tiempo. No }
\put(348.5906,-578.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}requiere de ningún proceso o algoritmo de }
\put(348.5906,-589.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}comprensión de datos. }
\put(330.5906,-602.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Replica (Replication), a esto también se le }
\put(348.5906,-613.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}denomina base de datos publica.  }
\put(330.5906,-625.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Elementos y frecuencia de respaldo, debido a }
\put(348.5906,-637.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}las amenazas en informática como hacking o }
\put(348.5906,-648.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}desastres naturales es fundamental tomar en }
\put(348.5906,-660.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cuenta el medio de almacenamiento en su }
\put(348.5906,-671.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}forma física, su vida útil, volúmenes, etc.  }
\put(330.5906,-683.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Es ideal tener un sistema de rotación de los }
\put(348.5906,-695.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}medios físicos para respaldo y definir la }
\put(348.5906,-706.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}frecuencia bajo la cual se llevan a cabo los }
\put(348.5906,-718.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}respaldos. }
\put(330.5906,-730.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Comandos de respaldos, son comandos en el }
\put(348.5906,-741.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}sistema para la implementación de respaldos }
\put(348.5906,-753.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ligado al O.S. y dada la importancia que exista }
\put(348.5906,-765.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}un error humano entre los comandos a utilizar }
\put(348.5906,-776.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}hay programas utilitarios que pueden hacerse }
\put(348.5906,-788.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cargo de estas operaciones. }
\put(294.5906,-799.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-811.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Las operaciones de monitoreo y control son hoy en día }
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}fundamentales, debido al exponencial flujo de información }
\put(33.47067,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}que se experimenta día a día. Para esto debe existir }
\put(33.47067,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}operaciones de control y monitoreo sobre las BB.DD. que }
\put(33.47067,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}se han tornado muy importantes. Estas son: }
\put(33.47067,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(69.47067,-105.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Monitoreo general, esto se refiere a los }
\put(87.47067,-116.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}problemas generales en el desempeño de los }
\put(87.47067,-128.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}sistemas. Es imperativo establecer buenas }
\put(87.47067,-139.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}practicas y directrices para que se pueda }
\put(87.47067,-151.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}enfrentar cualquier inconveniente de una }
\put(87.47067,-162.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}manera rápida. }
\put(69.47067,-174.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Monitoreo del espacio en Disco, se deben }
\put(87.47067,-186.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}establecer políticas de respaldo y }
\put(87.47067,-197.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}recuperación, por lo que incrementara su }
\put(87.47067,-209.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}espacio en Disco requerido. Es una tarea que }
\put(87.47067,-220.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}el Sysadmin debe proyectar. }
\put(69.47067,-232.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Monitoreo de Logs, son archivos con registros }
\put(87.47067,-244.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de transacciones, operaciones y cambios }
\put(87.47067,-255.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}realizados a la BB.DD. que pueden ser }
\put(87.47067,-267.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}analizados. }
\put(69.47067,-279.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Monitoreo de memoria compartida, se debe }
\put(87.47067,-291.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}compartir memoria RAM con otros }
\put(87.47067,-302.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}programas, es vital aquí la gestión de la }
\put(87.47067,-314.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}BB.DD. y la instalación y configuración del }
\put(87.47067,-325.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DBMS, para especificar cuanta memoria será }
\put(87.47067,-337.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}reservada para el sistema. }
\put(69.47067,-349.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Monitoreo de BB.DD., se mantienen en }
\put(87.47067,-360.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}observación constante las bases de datos, esto }
\put(87.47067,-372.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}permite reportes para ser analizados. El }
\put(87.47067,-383.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}proceso de actividades es conocido como }
\put(87.47067,-395.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DAM (Database Activity Monitoring). }
\put(69.47067,-407.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Monitoreo de Modos de operación, el }
\put(87.47067,-419.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}monitoreo constante hace posible la }
\put(87.47067,-430.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}generación de Logs de alerta, que conserva }
\put(87.47067,-442.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}información relacionada a errores o }
\put(87.47067,-453.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}situaciones irregulares, esto hace posible }
\put(87.47067,-465.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tomar las acciones correctivas necesarias para }
\put(87.47067,-476.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}prevenir ocurrencias de algún tipo en el }
\put(87.47067,-488.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}futuro. }
\put(69.47067,-500.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Monitoreo de espejos, tal como se menciono }
\put(87.47067,-511.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}antes la técnica “Mirror” o espejo es muy }
\put(87.47067,-523.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}efectiva. Permite garantizar integridad y }
\put(87.47067,-535.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}confiabilidad de los datos. }
\put(33.47067,-546.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-558.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-569.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-581.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-592.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-604.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-615.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Tomando sin duda en cuenta la BB.DD. que se entrego, }
\put(33.47067,-627.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para desarrollar las consultas en SQL. Desarrolle una }
\put(33.47067,-639.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}BB.DD. nombrada como ARTICULOS con sus respectivas }
\put(33.47067,-650.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Tablas, ‘Cod\_Art’, ‘Descripcion’ y ‘Precio’. }
\put(33.47067,-662.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(38.82078,-673.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}¿Cómo se puede obtener de el/los artículos con el menor }
\put(38.82078,-685.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}usuario?, a mi parecer es bastante simple obtener en este }
\put(38.82078,-696.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}caso el resultado. Para lograr el objetivo de menor precio se }
\put(38.82078,-708.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}debe utilizar la función MIN, la cual devolverá como }
\put(38.82078,-719.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}resultado un valor mínimo de la columna deseada.  }
\put(38.82078,-731.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Funciones MAX \& MIN estas funciones retornan, }
\put(38.82078,-742.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}respectivamente, el menor y mayor valor de los valores }
\put(38.82078,-754.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}contenidos en la columna especificado. }
\put(38.82078,-765.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(38.82078,-777.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estas dos funciones son muy necesarias al momento de }
\put(38.82078,-789.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}trabajar en la hoja de cálculo Excel o en una BB.DD. y es }
\put(38.82078,-800.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}que las mismas tienen como función principal ayudar a }
\put(38.82078,-812.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}encontrar valores determinados. }
\put(299.9407,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}En el caso de la función MIN va a proporcionar el valor }
\put(299.9407,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mínimo de una lista de valores, excluyendo los valores de }
\put(299.9407,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}textos y lógicos. }
\put(299.9407,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}En este tipo de función se pueden considerar los siguientes }
\put(299.9407,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}aspectos que nombrare a continuación: }
\put(299.9407,-104.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(340.5907,-116.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Se toman en cuenta los valores lógicos y los }
\put(358.5907,-128.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}textos de números escritos directamente en }
\put(358.5907,-139.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la lista de argumentos. }
\put(340.5907,-151.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Los argumentos pueden llegar a ser tanto }
\put(358.5907,-163.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}nombres, matrices, número o referencias que }
\put(358.5907,-174.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}contengan números. }
\put(340.5907,-186.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- En el caso que los argumentos que no }
\put(358.5907,-198.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}contengan ningún número entonces la }
\put(358.5907,-209.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}función MIN devuelve 0. }
\put(340.5907,-221.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- En el caso que el argumento se trata de una }
\put(358.5907,-233.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}referencia o una matriz, se utilizará }
\put(358.5907,-244.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}solamente los números que estén contenidos }
\put(358.5907,-256.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ya sea en la matriz o referencia. Lo que son }
\put(358.5907,-267.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}las celdas vacías, los textos contenidos en las }
\put(358.5907,-279.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}referencia o matriz y los valores lógicos se }
\put(358.5907,-291.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}pasaran por alto. }
\put(340.5907,-303.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Si existen valores de error o texto que no se }
\put(358.5907,-314.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}puedan traducir a números simplemente }
\put(358.5907,-326.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}generarán un error.}
\put(434.9906,-323.1808){\fontsize{6}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}iv}
\put(439.4932,-326.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(299.9407,-337.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(299.9407,-349.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Mientras que la función MAX va a permitir poder obtener el }
\put(299.9407,-360.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}valor máximo de un conjunto de valores ya sea una lista de }
\put(299.9407,-372.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}números ubicados en una BB.DD. De esta manera, la }
\put(299.9407,-383.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}función permite devolver el valor máximo de un conjunto de }
\put(299.9407,-395.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}valores. }
\put(299.9407,-406.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Como un argumento de la función se puede ingresar los }
\put(299.9407,-418.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}números que se quieren evaluar o simplemente un rango o la }
\put(299.9407,-430.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}referencia de una celda que contenga los valores numéricos. }
\put(299.9407,-441.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Ej.: }
\put(299.9407,-453.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(299.9407,-464.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SELECT MIN(Salario) FROM Empleados; }
\put(299.9407,-476.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Y retorna en algún caso hipotético: 150.000 }
\put(299.9407,-487.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Otro ejemplo utilizando la función MAX: }
\put(299.9407,-499.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SELECT MAX(Salario) FROM Empleados; }
\put(299.9407,-510.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Retornando en algún caso hipotético: 980.000 }
\put(299.9407,-522.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(299.9407,-534.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}En este caso se requiere tratar de encontrar el articulo con el }
\put(299.9407,-545.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}menor precio, por lo siguiente la columna donde se hará la }
\put(299.9407,-557.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}consulta, considerando que realicé una columna con }
\put(299.9407,-568.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}determinados ID para cada valor esta en la segunda fila }
\put(299.9407,-580.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}‘Preciocol’ siendo Precio la Tabla. La instrucción SQL sería: }
\put(299.9407,-591.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}              }
\put(299.9407,-603.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SELECT MIN(Preciocol) FROM ARTICULOS.Precio; }
\put(299.9407,-615.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-626.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Ahora bien, para responder a la pregunta numero dos, acerca }
\put(294.5906,-638.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de, ¿cómo generaría una consulta que contenga la }
\put(294.5906,-649.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}descripción de cada artículo con su respectivo precio, y, }
\put(294.5906,-661.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ordenándolo con la misma forma descendente en base al }
\put(294.5906,-672.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}precio del articulo? El ordenamiento del resultado de una }
\put(294.5906,-684.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}consulta, obtenida mediante únala clausula SELECT, se }
\put(294.5906,-695.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}logra mediante el uso de la clausula ORDER la cual }
\put(294.5906,-707.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}organizara los registros en base al contenido del campo }
\put(294.5906,-718.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}especificado como parámetro para la misma. Serán }
\put(294.5906,-730.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ordenados de forma ascendente, sin embargo, puede ser }
\put(294.5906,-741.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}revertido. }
\put(299.9407,-753.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}La siguiente sintaxis de esta cláusula es la siguiente: }
\put(299.9407,-765.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SELECT <columna 1, columna 2, … columna n> FROM }
\put(299.9407,-776.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}<nombre de la tabla> ORDER BY <columna> ASC/DESC. }
\put(299.9407,-788.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(299.9407,-799.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}También se utilizan ciertos elementos, utilizando la clausula }
\put(299.9407,-811.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ORDER, tomando en cuenta, como dije anteriormente }
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(38.82078,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}desarrollé una estructura SQL con sus respectivos ID para }
\put(38.82078,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cada columna, siendo en este caso ‘Cod\_Art’, }
\put(38.82078,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}‘Descripcion’ y ‘Precio’ las respectivas Tablas. El siguiente }
\put(38.82078,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}script SQL a aplicar, seria el siguiente: }
\put(69.47067,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}           }
\put(69.47067,-104.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SELECT DISTINCT Descriptioncol, Preciocl }
\put(33.47067,-116.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}FROM ARTICULOS.Descripcion, ARTICULOS.Preico; }
\put(69.47067,-127.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(69.47067,-139.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Esta consulta SQL dará como resultado valores del }
\put(69.47067,-150.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}campo ‘Preciocol’ y a su lado su descripción. }
\put(69.47067,-162.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Como siguiente ejercicio final, la siguiente }
\put(69.47067,-173.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}consulta SQL es para ordenar los elementos de la }
\put(69.47067,-185.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tabla de menor a mayor, mostrando su respectiva }
\put(69.47067,-196.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}descripción de cada articulo junto al precio }
\put(69.47067,-208.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}descendente, sería: }
\put(33.47067,-219.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}                            }
\put(69.47067,-231.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SELECT DISTINCT Descripcioncol, Preciocol }
\put(33.47067,-243.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}FROM ARTICULOS.Descripcion, ARTICULOS.Precio }
\put(33.47067,-254.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ORDER BY Preciocol DESC; }
\put(33.47067,-266.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-277.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-289.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-300.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-312.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-323.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-335.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}La importancia del manejo de la información he hecho }
\put(33.47067,-346.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}que la relevancia de las DBMS ampliase su gestión en }
\put(33.47067,-358.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}bases de datos y tablas y abarcando la estructura lógica. }
\put(33.47067,-369.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SQL mantiene comandos que permiten manipular tablas en }
\put(33.47067,-381.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}bases de datos tanto por usuarios como por administradores. }
\put(33.47067,-393.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estos hacen posible un mejor control de los datos y }
\put(33.47067,-404.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}paramentos que garantizan su confiabilidad e integridad. }
\put(33.47067,-416.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Los DBMS mejoraron, en su desarrollo, y pueden ejecutar }
\put(33.47067,-427.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}consultas en bases de datos, pudiendo combinar, de varias }
\put(33.47067,-439.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}maneras, datos contenidos en diferentes tablas para poder }
\put(33.47067,-450.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}generar reportes facilitando la toma de decisiones }
\put(33.47067,-462.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}gerenciales de manera rápida y oportuna. }
\put(33.47067,-473.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-485.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estructura lógica:  La explicación lógica y modelo }
\put(33.47067,-497.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}relacional de las correspondientes tablas en la base de datos }
\put(33.47067,-508.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}empresa, conectada con la tabla Contacto, dentro de la base }
\put(33.47067,-520.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de datos: }
\put(260.2706,-652.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-661.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Una tabla creada en SQL requiere que cada tabla creada }
\put(33.47067,-673.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tenga un nombre asignado, así como los campos que }
\put(33.47067,-684.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}contendrá. Se le debe asignar un nombre y, necesita definir }
\put(33.47067,-696.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}el tipo de dato.  De tal manera: }
\put(33.47067,-707.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30046}CREATE TABLE \`Empresa\`.\`Empleados\` ( \`Id\_Empleado\` }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -708.6208pt) -- (72.3506pt, -708.6208pt)
 -- (72.3506pt, -708.6208pt)
 -- (72.3506pt, -709.1008pt)
 -- (72.3506pt, -709.1008pt)
 -- (33.4706pt, -709.1008pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(74.7506pt, -708.6208pt) -- (107.1506pt, -708.6208pt)
 -- (107.1506pt, -708.6208pt)
 -- (107.1506pt, -709.1008pt)
 -- (107.1506pt, -709.1008pt)
 -- (74.7506pt, -709.1008pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-719.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}INT(11) NULL , \`Nombre\` VARCHAR(255) NOT NULL , }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -720.3808pt) -- (50.0306pt, -720.3808pt)
 -- (50.0306pt, -720.3808pt)
 -- (50.0306pt, -720.8608pt)
 -- (50.0306pt, -720.8608pt)
 -- (33.4706pt, -720.8608pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(145.3106pt, -720.3808pt) -- (194.2706pt, -720.3808pt)
 -- (194.2706pt, -720.3808pt)
 -- (194.2706pt, -720.8608pt)
 -- (194.2706pt, -720.8608pt)
 -- (145.3106pt, -720.8608pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(218.5106pt, -720.3808pt) -- (238.9106pt, -720.3808pt)
 -- (238.9106pt, -720.3808pt)
 -- (238.9106pt, -720.8608pt)
 -- (238.9106pt, -720.8608pt)
 -- (218.5106pt, -720.8608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-731.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}\`Apellido\` VARCHAR(255) NOT NULL , \`Departamento\` }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(77.6306pt, -731.9008pt) -- (126.5906pt, -731.9008pt)
 -- (126.5906pt, -731.9008pt)
 -- (126.5906pt, -732.3808pt)
 -- (126.5906pt, -732.3808pt)
 -- (77.6306pt, -732.3808pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(150.5906pt, -731.9008pt) -- (171.2306pt, -731.9008pt)
 -- (171.2306pt, -731.9008pt)
 -- (171.2306pt, -732.3808pt)
 -- (171.2306pt, -732.3808pt)
 -- (150.5906pt, -732.3808pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-742.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}VARCHAR(255) NOT NULL , \`Salario\` INT(11) NOT }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -743.4208pt) -- (82.4306pt, -743.4208pt)
 -- (82.4306pt, -743.4208pt)
 -- (82.4306pt, -743.9008pt)
 -- (82.4306pt, -743.9008pt)
 -- (33.4706pt, -743.9008pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(106.4306pt, -743.4208pt) -- (127.0706pt, -743.4208pt)
 -- (127.0706pt, -743.4208pt)
 -- (127.0706pt, -743.9008pt)
 -- (127.0706pt, -743.9008pt)
 -- (106.4306pt, -743.9008pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(201.2306pt, -743.4208pt) -- (217.7906pt, -743.4208pt)
 -- (217.7906pt, -743.4208pt)
 -- (217.7906pt, -743.9008pt)
 -- (217.7906pt, -743.9008pt)
 -- (201.2306pt, -743.9008pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(236.9906pt, -743.4208pt) -- (257.6306pt, -743.4208pt)
 -- (257.6306pt, -743.4208pt)
 -- (257.6306pt, -743.9008pt)
 -- (257.6306pt, -743.9008pt)
 -- (236.9906pt, -743.9008pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-754.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}NULL , PRIMARY KEY (\`Id\_Empleado\`)) ENGINE = }
\put(33.47067,-765.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}InnoDB;  }
\put(33.47067,-777.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(566.5106,-179.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-189.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-200.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para definir la auto incrementación en fila Id\_Empleado, se }
\put(294.5906,-212.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ejecuta el siguiente script: }
\put(294.5906,-223.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-235.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}ALTER TABLE \`Empleados\` CHANGE \`Id\_Empleado\` }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(294.5906pt, -236.0608pt) -- (326.7506pt, -236.0608pt)
 -- (326.7506pt, -236.0608pt)
 -- (326.7506pt, -236.5408pt)
 -- (326.7506pt, -236.5408pt)
 -- (294.5906pt, -236.5408pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(329.3906pt, -236.0608pt) -- (361.5506pt, -236.0608pt)
 -- (361.5506pt, -236.0608pt)
 -- (361.5506pt, -236.5408pt)
 -- (361.5506pt, -236.5408pt)
 -- (329.3906pt, -236.5408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-246.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}\`Id\_Empleado\` INT(11) NOT NULL AUTO\_INCREMENT;  }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(357.7106pt, -247.5808pt) -- (374.2706pt, -247.5808pt)
 -- (374.2706pt, -247.5808pt)
 -- (374.2706pt, -248.0608pt)
 -- (374.2706pt, -248.0608pt)
 -- (357.7106pt, -248.0608pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(393.4706pt, -247.5808pt) -- (414.1106pt, -247.5808pt)
 -- (414.1106pt, -247.5808pt)
 -- (414.1106pt, -248.0608pt)
 -- (414.1106pt, -248.0608pt)
 -- (393.4706pt, -248.0608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-258.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-269.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-281.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-292.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-304.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-315.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-327.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-339.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-350.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Al hablar de Índice, y la creación de esta con base en varias }
\put(294.5906,-362.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}columnas, con la finalidad de posibilitar la búsqueda de }
\put(294.5906,-373.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}registros mas expedita, se utiliza:  }
\put(294.5906,-385.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ALTER TABLE \`Empresa\`.\`Empleados\` ADD INDEX }
\put(294.5906,-396.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}\`Clave\_empleado\` (\`Id\_Empleado\`, \`Nombre\`, \`Apellido\`, }
\put(294.5906,-408.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}\`Departamento\`, \`Salario\`);  }
\put(294.5906,-419.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(547.0706,-568.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-577.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-589.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-600.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-612.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-623.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-635.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Una forma de mantenimiento en SQL para eliminar registros }
\put(294.5906,-646.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de una BB.DD., y para borrar tabla o un índice. La sintaxis }
\put(294.5906,-658.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}es: }
\put(294.5906,-670.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}DELETE FROM \`Empleados\` WHERE \`Id\_Empleado\` = }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(294.5906pt, -670.9408pt) -- (332.2706pt, -670.9408pt)
 -- (332.2706pt, -670.9408pt)
 -- (332.2706pt, -671.4208pt)
 -- (332.2706pt, -671.4208pt)
 -- (294.5906pt, -671.4208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-681.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}"11"; }
\put(294.5906,-693.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(34.3906,-653.6105){\includegraphics[width=226.9127pt,height=130.1pt]{latexImage_029a2dd6600756c21d6e5edbb0b6bddc.png}}
\put(295.3906,-180.8509){\includegraphics[width=271.9371pt,height=142.35pt]{latexImage_5aa95ed820e4bd9edcee359f6e1a8968.png}}
\put(295.3906,-569.5557){\includegraphics[width=252.4317pt,height=146.2pt]{latexImage_cb9e050616eda6cdf23c12de1cc08168.png}}
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(269.3906,-233.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-242.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}El resultado de esta en la fila Id\_Empleado = 11 ya no }
\put(33.47067,-254.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}existirá.  }
\put(33.47067,-265.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-277.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-289.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-300.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-312.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Si se desea eliminar todos los registros, la sintaxis seria la }
\put(33.47067,-323.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}siguiente: }
\put(33.47067,-335.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DELETE FROM Empleados }
\put(33.47067,-346.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-358.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para Borrar Una Tabla Completa De Una BB.BB. Se }
\put(33.47067,-369.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Utiliza El Comando Drop Table: }
\put(33.47067,-381.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DROP TABLE Empleados }
\put(33.47067,-393.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-404.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Y Para Borrar Un Índice: }
\put(33.47067,-416.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DROP INDEX Clave\_Empleado; }
\put(33.47067,-427.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-439.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para Afrontar Actualizaciones, En SQL Existe El Comando }
\put(33.47067,-450.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}‘update’: }
\put(33.47067,-462.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}UPDATE Empleados SET Nombre = "Donald", }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -462.8608pt) -- (72.8306pt, -462.8608pt)
 -- (72.8306pt, -462.8608pt)
 -- (72.8306pt, -463.3408pt)
 -- (72.8306pt, -463.3408pt)
 -- (33.4706pt, -463.3408pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(122.2706pt, -462.8608pt) -- (140.0306pt, -462.8608pt)
 -- (140.0306pt, -462.8608pt)
 -- (140.0306pt, -463.3408pt)
 -- (140.0306pt, -463.3408pt)
 -- (122.2706pt, -463.3408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-473.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Departamento = "IT" WHERE Id\_Empleado = "9" ; }
\put(33.47067,-485.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-496.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(269.3906,-531.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-540.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-551.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-563.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-575.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-586.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-598.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-609.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-621.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-632.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-644.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-655.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Consultas De Acción: Son comandos a ejecutar sobre las }
\put(33.47067,-667.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tablas Tablas como: Modificación, borrado, inserción de }
\put(33.47067,-679.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}registro, y otros. }
\put(51.47067,-691.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- INSERCIÓN, consiste en añadir nuevos registros }
\put(69.47067,-702.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}en las tablas, ej.: }
\put(33.47067,-714.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}INSERT INTO \`Empleados\` (\`Id\_Empleado\`, \`Nombre\`, }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -714.8608pt) -- (68.5106pt, -714.8608pt)
 -- (68.5106pt, -714.8608pt)
 -- (68.5106pt, -715.3408pt)
 -- (68.5106pt, -715.3408pt)
 -- (33.4706pt, -715.3408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-725.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}\`Apellido\`, \`Departamento\`, \`Salario\`) VALUES (NULL, }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(189.2306pt, -726.3808pt) -- (228.5906pt, -726.3808pt)
 -- (228.5906pt, -726.3808pt)
 -- (228.5906pt, -726.8608pt)
 -- (228.5906pt, -726.8608pt)
 -- (189.2306pt, -726.8608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-737.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'Luis', 'Saldiver', 'Contabilidad', '340000'), (NULL, 'Laura', }
\put(33.47067,-748.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'Perez', 'IT', '480000'), (NULL, 'Estela', 'Mardonez', }
\put(33.47067,-760.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'Mercadeo', '360000'), (NULL, 'Maria', 'Lara', 'Ventas', }
\put(33.47067,-771.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'980000'), (NULL, 'Jose', 'Gutierrez', 'Finanzas', '150000'), }
\put(33.47067,-783.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}(NULL, 'Pedro', 'Zerpa', 'Contabilidad', '400000'), (NULL, }
\put(33.47067,-795.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'Esteban', 'Saldiver', 'IT', '289000'), (NULL, 'Luisa', 'Perez', }
\put(33.47067,-806.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'Mercadeo', '545000'), (NULL, 'Jaime', 'Mardonez', 'Ventas', }
\put(294.5906,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'600000'), (NULL, 'Loreto', 'Lara', 'Finanzas', '388000'), }
\put(294.5906,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}(NULL, 'Estefani', 'Gutierrez', 'Finanzas', '290000'), (NULL, }
\put(294.5906,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'Angel', 'Campos', 'IT', '780000'); }
\put(312.5906,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(312.5906,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(312.5906,-104.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(550.4306,-319.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-328.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-339.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-351.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-363.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Consultas De Selección Avanzadas }
\put(294.5906,-374.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Esto Se Logra Mediante Cláusulas Como, Where, Group By, }
\put(294.5906,-386.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Having Y Order By, Que Aumentan El Valor De Select. }
\put(294.5906,-397.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}SELECT Nombre FROM Empleados GROUP BY }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(294.5906pt, -398.3008pt) -- (331.3106pt, -398.3008pt)
 -- (331.3106pt, -398.3008pt)
 -- (331.3106pt, -398.7808pt)
 -- (331.3106pt, -398.7808pt)
 -- (294.5906pt, -398.7808pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-409.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Id\_Empleado }
\put(294.5906,-420.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(537.5707,-716.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-725.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-736.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-748.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-759.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Clausula, HAVING (Requiere Registros Previamente }
\put(294.5906,-771.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Agrupados): }
\put(294.5906,-783.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}SELECT Id\_Empleado,SUM(Salario) FROM Empleados }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(294.5906pt, -783.7408pt) -- (331.3106pt, -783.7408pt)
 -- (331.3106pt, -783.7408pt)
 -- (331.3106pt, -784.2208pt)
 -- (331.3106pt, -784.2208pt)
 -- (294.5906pt, -784.2208pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(390.1106pt, -783.7408pt) -- (411.7106pt, -783.7408pt)
 -- (411.7106pt, -783.7408pt)
 -- (411.7106pt, -784.2208pt)
 -- (411.7106pt, -784.2208pt)
 -- (390.1106pt, -784.2208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-794.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}GROUP BY Id\_Empleado HAVING SUM(Salario)>300000  }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(445.7906pt, -795.2608pt) -- (467.3906pt, -795.2608pt)
 -- (467.3906pt, -795.2608pt)
 -- (467.3906pt, -795.7408pt)
 -- (467.3906pt, -795.7408pt)
 -- (445.7906pt, -795.7408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-806.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(34.3906,-234.6009){\includegraphics[width=236.0767pt,height=196.1pt]{latexImage_3f6b737c362b953150216db7fafb0711.png}}
\put(34.3906,-532.2515){\includegraphics[width=236.0254pt,height=32.05pt]{latexImage_2e50c2adc60c2db8056b971983da4f94.png}}
\put(295.3906,-320.1879){\includegraphics[width=255.6995pt,height=212.4pt]{latexImage_3f6b737c362b953150216db7fafb0711.png}}
\put(297.3906,-716.9184){\includegraphics[width=240.55pt,height=292.8pt]{latexImage_dbae88d863c1f7c69e529187ee8d4389.png}}
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(270.5908,-177.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-186.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-198.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-210.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-221.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-233.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-244.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-256.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-267.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-279.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-290.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-302.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Intersecciones: }
\put(33.47067,-314.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Tomando en cuenta que se creó la base de datos ‘Contacto’, }
\put(33.47067,-325.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la sintaxis es: }
\put(33.47067,-337.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}CREATE TABLE \`Contacto\`.\`3\` ( \`Id\_Empleado\` INT(11) }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -337.8208pt) -- (72.3506pt, -337.8208pt)
 -- (72.3506pt, -337.8208pt)
 -- (72.3506pt, -338.3008pt)
 -- (72.3506pt, -338.3008pt)
 -- (33.4706pt, -338.3008pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(74.7506pt, -337.8208pt) -- (107.1506pt, -337.8208pt)
 -- (107.1506pt, -337.8208pt)
 -- (107.1506pt, -338.3008pt)
 -- (107.1506pt, -338.3008pt)
 -- (74.7506pt, -338.3008pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(237.9506pt, -337.8208pt) -- (254.5106pt, -337.8208pt)
 -- (254.5106pt, -337.8208pt)
 -- (254.5106pt, -338.3008pt)
 -- (254.5106pt, -338.3008pt)
 -- (237.9506pt, -338.3008pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-348.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}NULL AUTO\_INCREMENT , \`Telefono\` }
\put(33.47067,-360.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}VARCHAR(255) NOT NULL , \`Apellido\` }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -360.8608pt) -- (82.4306pt, -360.8608pt)
 -- (82.4306pt, -360.8608pt)
 -- (82.4306pt, -361.3408pt)
 -- (82.4306pt, -361.3408pt)
 -- (33.4706pt, -361.3408pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(106.4306pt, -360.8608pt) -- (127.0706pt, -360.8608pt)
 -- (127.0706pt, -360.8608pt)
 -- (127.0706pt, -361.3408pt)
 -- (127.0706pt, -361.3408pt)
 -- (106.4306pt, -361.3408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-371.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}VARCHAR(255) NOT NULL , PRIMARY KEY }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -372.3808pt) -- (82.4306pt, -372.3808pt)
 -- (82.4306pt, -372.3808pt)
 -- (82.4306pt, -372.8608pt)
 -- (82.4306pt, -372.8608pt)
 -- (33.4706pt, -372.8608pt) -- cycle
;
\filldraw[color_30045][nonzero rule]
(106.4306pt, -372.3808pt) -- (127.0706pt, -372.3808pt)
 -- (127.0706pt, -372.3808pt)
 -- (127.0706pt, -372.8608pt)
 -- (127.0706pt, -372.8608pt)
 -- (106.4306pt, -372.8608pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-383.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}(\`Id\_Empleado\`)) ENGINE = InnoDB;  }
\put(33.47067,-394.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(274.4306,-525.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-534.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-545.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-557.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-568.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-580.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-591.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-603.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-615.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-626.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Inserto datos: }
\put(33.47067,-638.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}INSERT INTO \`3\` (\`Id\_Empleado\`, \`Telefono\`, \`Apellido\`) }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -639.0208pt) -- (68.5106pt, -639.0208pt)
 -- (68.5106pt, -639.0208pt)
 -- (68.5106pt, -639.5008pt)
 -- (68.5106pt, -639.5008pt)
 -- (33.4706pt, -639.5008pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-649.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}VALUES (NULL, '955787689', 'Saldiver'), (NULL, }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(33.4706pt, -650.5408pt) -- (72.8306pt, -650.5408pt)
 -- (72.8306pt, -650.5408pt)
 -- (72.8306pt, -651.0208pt)
 -- (72.8306pt, -651.0208pt)
 -- (33.4706pt, -651.0208pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-661.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}'964356545', 'Perez'), (NULL, '987445323', 'Mardonez'), }
\put(33.47067,-672.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}(NULL, '934558879', 'Lara')  }
\put(33.47067,-684.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(271.8657,-793.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-802.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} ‘Union’ permite combinar los registros obtenidos mediante }
\put(33.47067,-813.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la ejecución de dos o mas cláusulas, que deben contemplar }
\put(294.5906,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la misma cantidad de columnas.  }
\put(294.5906,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}SELECT Id\_Empleado, Nombre FROM Empleados UNION }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(294.5906pt, -58.9408pt) -- (331.3106pt, -58.9408pt)
 -- (331.3106pt, -58.9408pt)
 -- (331.3106pt, -59.42078pt)
 -- (331.3106pt, -59.42078pt)
 -- (294.5906pt, -59.42078pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_30045}SELECT Id\_Empleado, Telefono FROM Contacto; }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_30045][nonzero rule]
(294.5906pt, -70.46082pt) -- (331.3106pt, -70.46082pt)
 -- (331.3106pt, -70.46082pt)
 -- (331.3106pt, -70.9408pt)
 -- (331.3106pt, -70.9408pt)
 -- (294.5906pt, -70.9408pt) -- cycle
;
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(294.5906,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(520.9106,-214.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-223.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-235.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SQL mantiene un conjunto de comandos que logran manejar }
\put(294.5906,-247.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}bases de datos de una forma fácil e intuitiva. La cláusula }
\put(294.5906,-258.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}SELECT siempre tiene una gran versatilidad, Permite la }
\put(294.5906,-270.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}realización de cualquier tipo de consulta sobre los datos.  }
\put(294.5906,-281.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estas características han hecho de SQL el lenguaje preferido }
\put(294.5906,-293.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para el manejo y gestión de los sistemas de bases de datos, lo }
\put(294.5906,-304.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cual ha generado que se desarrollen diferentes versiones para }
\put(294.5906,-316.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ser usadas en entornos más específicos. }
\put(294.5906,-327.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-339.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-351.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-362.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Originalmente MySQL era un software propietario, 5 }
\put(294.5906,-374.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}años después de 1995 se volvió OpenSource con soporte }
\put(294.5906,-385.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para Linux-Unix.  }
\put(294.5906,-397.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}MySQL no estuvo orientado a una gran escala, su uso era }
\put(294.5906,-408.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mas bien practico. Fué tan bien escrito, que se convirtió en }
\put(294.5906,-420.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}un DBMS preferido. }
\put(294.5906,-431.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Su instalación difiere un poco dependiendo de ciertos }
\put(294.5906,-443.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}aspectos, como:  }
\put(312.5906,-455.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Establecer compatibilidad para MySQL, se requiere }
\put(330.5906,-466.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}verificar si soporta y esta respaldado por la empresa }
\put(330.5906,-478.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Oracle Corp. }
\put(312.5906,-490.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Decidir cual distribución instalar, binario, o código }
\put(330.5906,-501.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}abierto (Permite acceso para desarrollo), verificar }
\put(330.5906,-513.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}descarga bajo Oracle Corp., Verificar bajo que O.S. }
\put(330.5906,-525.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}se instalará.  }
\put(312.5906,-537.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- MySQL Community Server, descargable de manera }
\put(330.5906,-548.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}libre, para pequeños o medianos usuarios, soportada }
\put(330.5906,-560.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}bajo licencia GPL, así se tiene acceso a esta, como }
\put(330.5906,-571.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}código abierto, soportada para colaboradores y }
\put(330.5906,-583.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Developers. }
\put(299.9407,-594.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-606.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Bajo la licencia GPL, existen limitaciones comparado a }
\put(294.5906,-617.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Enterprise Edition que tiene funciones adicionales para el }
\put(294.5906,-629.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ámbito empresarial. }
\put(294.5906,-640.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-652.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Se puede crear una tabla con un respectivo ID, por eje.: }
\put(294.5906,-664.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}CREATE TABLE Empleados (Nombre varchar(50), }
\put(294.5906,-675.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Apellido varchar(50), ID int AUTOINCREMENT }
\put(294.5906,-687.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}PRIMARY KEY (ID) ); }
\put(294.5906,-698.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-710.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para ingresar 10 registros en la tabla anterior considerando }
\put(294.5906,-721.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}que en la columna Apellido }
\put(294.5906,-733.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cinco de ellos deben comenzar con la letra D.: }
\put(294.5906,-744.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-756.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}INSERT INTO Empleados (Nombre, Apellido, ID) }
\put(294.5906,-768.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}VALUES (‘Dohn, ‘Lemon’, ‘’), (‘Daul’, ‘Martney’,’’), }
\put(294.5906,-779.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}(‘Detephen’, ‘Ding’,’’), (‘Daily’, ‘Moon’,’’), (‘Dinitry’, }
\put(294.5906,-791.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}‘Larry’,’’), (‘Stephen’, ‘King’,’’), (‘Larry’, ‘Olson’,’’), }
\put(294.5906,-802.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}(‘Peter’, ‘Woom’,’’), (‘Tonny’, ‘Waka’,’’), (‘David’, }
\put(294.5906,-814.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}‘B}
\put(34.3906,-178.7408){\includegraphics[width=237.25pt,height=140.24pt]{latexImage_09e5474c8b7ac09b7bc7703bae9b62bb.png}}
\put(34.3906,-525.76){\includegraphics[width=240.745pt,height=127.6pt]{latexImage_c8abde8d9a9b3e6dd50b624087cada04.png}}
\put(37.31564,-793.9799){\includegraphics[width=235.4641pt,height=106.15pt]{latexImage_f43fcf35e77d94bb17669589d1e61f00.png}}
\put(295.3906,-215.3423){\includegraphics[width=226.3294pt,height=130.65pt]{latexImage_408ff130a82c4122fdc9c8aefbc6050c.png}}
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.47067,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Apellido like = “D\%”);  }
\put(33.47067,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}MySQL dispone de un optimizador de consultas y este }
\put(33.47067,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}utiliza distribución de claves almacenadas y otros factores }
\put(33.47067,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para decidir el orden en que se deben unir las tablas y que }
\put(33.47067,-104.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}índice se debe usar para una tabla especifica.  }
\put(33.47067,-116.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Las estructuras de datos almacenan información entre }
\put(33.47067,-127.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}objetos y relaciones entre ellas. Para ello utiliza el modelo }
\put(33.47067,-139.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de entidad relación (ER). }
\put(33.47067,-150.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Un modelo -ER- se construye de manera grafica mediante }
\put(33.47067,-162.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}símbolos como: }
\put(51.47067,-174.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Rectángulos. }
\put(51.47067,-186.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Líneas. }
\put(38.82078,-197.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-209.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estas representan entidades en: }
\put(51.47067,-221.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Un proceso. }
\put(51.47067,-233.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Bajo características. }
\put(51.47067,-245.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Relaciones existentes. }
\put(33.47067,-256.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-268.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-279.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-291.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Las relaciones tienen elementos, como, objetos, atributos, }
\put(33.47067,-303.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}relaciones, tablas. }
\put(33.47067,-314.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-326.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Tablas: }
\put(33.47067,-337.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Las tablas son estructuras donde se almacenan los datos, sin }
\put(33.47067,-349.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}estas, seria imposible el concepto de datos relacionales. }
\put(33.47067,-360.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para esto se deben considerar aspectos como: }
\put(51.47067,-372.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Cantidad. }
\put(51.47067,-384.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Numero de columnas. }
\put(51.47067,-396.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Tipo de dato. }
\put(38.82078,-408.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-419.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para crear una tabla llamada Contacto\_Empleados, la que }
\put(33.47067,-431.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}debe contener las columnas }
\put(33.47067,-443.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Identificación del empleado, número telefónico y correo }
\put(33.47067,-454.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}electrónico, su instrucción es la Sgte.: }
\put(33.47067,-466.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}CREATE TABLE IF NOT EXISTS Contacto\_Empleados }
\put(33.47067,-477.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}(identificación int AUTOINCREMENT }
\put(33.47067,-489.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}PRIMARY KEY (Id\_Empleado), Numero\_Tel varchar(50), }
\put(33.47067,-500.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Email varchar(250) }
\put(33.47067,-512.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}); }
\put(33.47067,-523.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-535.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Creación de usuarios: }
\put(33.47067,-547.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Se pueden generar usuarios de diferentes maneras, pero }
\put(33.47067,-558.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}requieren de un usuario Root o Superusuario bajo conexión }
\put(33.47067,-570.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}al servidor, y bajo la tabla de usuarios en MySQL. }
\put(33.47067,-581.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-593.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Permisología: }
\put(33.47067,-604.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Al crear usuarios y asignarles permisos, se deben tomar en }
\put(33.47067,-616.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cuenta que tipo de recursos requeridos tendrá. Se pueden }
\put(33.47067,-627.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}asignar permisos particulares a un usuario o con acceso }
\put(33.47067,-639.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}completo. }
\put(33.47067,-651.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-662.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Perfiles: }
\put(33.47067,-674.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estos van a repercutir al ejecutar diferentes funciones, se }
\put(33.47067,-685.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}deben establecer permisos adecuados para no poner en }
\put(33.47067,-697.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}riesgo la seguridad de los datos.  }
\put(33.47067,-708.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-720.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Mantenimiento: }
\put(33.47067,-731.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}MySQL cuenta con varias herramientas para mantener las }
\put(33.47067,-743.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tablas, y así: }
\put(51.47067,-755.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Analizar }
\put(51.47067,-767.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Optimizar  }
\put(51.47067,-779.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Verificar }
\put(51.47067,-791.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Reparar Tablas }
\put(33.47067,-802.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-814.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Dependiendo de el numero de transacciones, sufrirán un }
\put(294.5906,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}proceso de fragmentación y existe para ello un determinado }
\put(294.5906,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}comando:  }
\put(294.5906,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}OPTIMIZE nombre de la tabla; }
\put(294.5906,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}También el comando para verificar integridad de las tablas, }
\put(294.5906,-104.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cuya sintaxis es:  }
\put(294.5906,-116.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}CHECK TABLE nombre de la tabla; }
\put(294.5906,-127.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-139.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-150.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-162.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para crear un usuario que pueda acceder a ambas tablas, pero }
\put(294.5906,-173.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}únicamente para ingresar y }
\put(294.5906,-185.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}modificar registros, bajo un particular requerimiento, la }
\put(294.5906,-196.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}instrucción es la Sgte.: }
\put(294.5906,-208.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-219.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Crear usuario: }
\put(294.5906,-231.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}CREATE USER usuario\_supervisor INDENTIFIED BY }
\put(294.5906,-243.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}PASSWORD ‘fafarafa65’; }
\put(294.5906,-254.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-266.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Otorgar permisos: }
\put(294.5906,-277.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}GRANT INSERT, UPDATE ON Empleados, }
\put(294.5906,-289.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Contacto\_Empleados TO ‘usuario\_supervisor’; }
\put(294.5906,-300.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-312.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Otras instrucciones bajo requerimientos específicos: }
\put(294.5906,-323.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Crear una vista de la base de datos Empleados que contenga }
\put(294.5906,-335.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}el Nombre y Apellido de }
\put(294.5906,-346.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}todos aquellos empleados que comiencen con la letra D.: }
\put(294.5906,-358.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-369.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}CREAT VIEW Temporal\_Empleados AS SELECT Nombre, }
\put(294.5906,-381.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Apellido FROM }
\put(294.5906,-393.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Empleados WHERE Nombre LIKE 'D\%'; }
\put(294.5906,-404.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-416.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Borrar la tabla Contacto\_Empleados: }
\put(294.5906,-427.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}DROP TABLE Contacto\_Empleados; }
\put(294.5906,-439.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-450.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}MySQL en el tiempo, ha generado una comunidad muy }
\put(294.5906,-462.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}grande en Internet entre Developers, para el desarrollo de }
\put(294.5906,-473.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}aplicaciones con esta tecnología. }
\put(294.5906,-485.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-497.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-508.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-520.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-531.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-543.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-554.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Al realizar gestión de seguridad del sistema mediante }
\put(294.5906,-566.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}MySQL, existen factores que deben ser tomados en cuenta al }
\put(294.5906,-577.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}momento de construir un sistema robusto en seguridad. Es }
\put(294.5906,-589.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}fundamental definir los permisos de usuarios, y tomar }
\put(294.5906,-601.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}medidas a nivel físico en servidores de datos. }
\put(294.5906,-612.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Es importante mantener políticas de seguridad apropiadas, }
\put(294.5906,-624.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}que en si son un conjunto de normas y directrices, que }
\put(294.5906,-635.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}establecen como los datos pueden ser accedidos. Implica }
\put(294.5906,-647.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}restricciones y el tipo de operación que se pueden llegar a }
\put(294.5906,-658.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ejecutar sobre las mismas. Es importante considerar: }
\put(294.5906,-670.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(312.5906,-682.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}• Passwords o claves de acceso son el primer frente }
\put(330.5906,-694.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de protección contra vulnerabilidades. Esto proceso }
\put(330.5906,-705.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de creación de passwords implica: }
\put(366.5906,-717.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Longitud mínima, debe contener un }
\put(384.5906,-729.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mínimo de 10 caracteres combinando }
\put(384.5906,-741.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}letras y números. }
\put(366.5906,-753.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Composición, como regla no debe }
\put(384.5906,-764.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tener datos personales o públicos. }
\put(366.5906,-776.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Almacenamiento de passwords, }
\put(384.5906,-788.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tomando en cuenta los pasos }
\put(384.5906,-799.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}anteriores se debe memorizar, nunca }
\put(384.5906,-811.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}almacenar passwords. }
\end{picture}
\newpage
\begin{tikzpicture}[overlay]\path(0pt,0pt);\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(105.4707,-47.18079){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Control histórico, obliga a usuarios a }
\put(123.4707,-58.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}no repetir passwords.  }
\put(105.4707,-70.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Compartir Passwords, usuarios deben }
\put(123.4707,-82.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ser instruidos para que se comprenda }
\put(123.4707,-93.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}la implicancia de la responsabilidad }
\put(123.4707,-105.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}que recae en el usuario final. }
\put(105.4707,-117.5009){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Transferencia electrónica, los }
\put(123.4707,-129.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}passwords no deben ser transmitidos }
\put(123.4707,-140.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}bajo ninguna circunstancia bajo }
\put(123.4707,-152.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}internet. }
\put(33.47067,-163.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-175.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Protección contra hackers, se debe combinar con el uso de }
\put(33.47067,-186.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}buenas practicas de manera generalizada y se recomienda: }
\put(105.4707,-198.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Mantener BB.DD. detrás de un }
\put(123.4707,-210.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Firewall en hardware y software. }
\put(105.4707,-222.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Al usar Unix, nunca trabajar bajo }
\put(123.4707,-233.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}usuario Root. }
\put(105.4707,-245.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Aplicar archivos de configuración }
\put(123.4707,-257.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para que mysql actúe de manera }
\put(123.4707,-268.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}automática }
\put(105.4707,-280.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- No almacenar variables en el entorno }
\put(123.4707,-292.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mysql\_pwd }
\put(105.4707,-304.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Restringir acceso a tabla mysql.user }
\put(105.4707,-316.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Actualizar MySQL a su ultima }
\put(123.4707,-327.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}versión ya que siempre incluyen }
\put(123.4707,-339.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mejoras. }
\put(33.47067,-351.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-362.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Respaldo de datos, la utilidad de copias de seguridad es }
\put(33.47067,-374.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}representada por su regularidad o frecuencia. MySQL tiene }
\put(33.47067,-385.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}herramientas como MySQL Enterprise Backup. }
\put(33.47067,-397.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Hacer un respaldo completo es necesario, pero no siempre }
\put(33.47067,-408.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}conveniente, ya que generan grandes archivos y consumen }
\put(33.47067,-420.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}tiempos de procesamiento, y no son efectivos, ya que todos }
\put(33.47067,-432.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}los respaldos son siempre diferentes respecto a su copia }
\put(33.47067,-443.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}anterior. Las copias de seguridad completa e incrementable }
\put(33.47067,-455.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}son mas pequeñas y tardan menos tiempo. Para hacer estos }
\put(33.47067,-466.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cambios incrementales en el registro binario, el servidor }
\put(33.47067,-478.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}debe iniciarse con la opción – - log-bin para habilitar ese }
\put(33.47067,-489.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}registro.  Es posible indicarle a MySQL el cierre de registro }
\put(33.47067,-501.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}binario actual usando la opción mysqladmin flush-logs. }
\put(33.47067,-512.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(33.47067,-524.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}MySQL tiene un comando simple parra recuperar datos, la }
\put(33.47067,-536.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}sintaxis es la siguiente: }
\put(33.47067,-547.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Mysqldump –u “usuario” –p “contraseña” }
\put(33.47067,-559.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}<nombre\_de\_la\_base\_de\_datos> < }
\put(33.47067,-570.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}nombre\_del\_respaldo.sql>; }
\put(33.47067,-582.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Ejemplo: }
\put(33.47067,-593.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}mysqldumpo –u supervisor –p 1536c* finanzas }
\put(177.4707,-621.0208){\fontsize{11.04}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\end{picture}
\begin{tikzpicture}[overlay]
\path(0pt,0pt);
\filldraw[color_29791][nonzero rule]
(33.4706pt, -618.1408pt) -- (177.4706pt, -618.1408pt)
 -- (177.4706pt, -618.1408pt)
 -- (177.4706pt, -618.6208pt)
 -- (177.4706pt, -618.6208pt)
 -- (33.4706pt, -618.6208pt) -- cycle
;
\begin{scope}
\clip
(-7pt, -840pt) -- (589pt, -840pt)
 -- (589pt, -840pt)
 -- (589pt, 3pt)
 -- (589pt, 3pt)
 -- (-7pt, 3pt) -- cycle
;
\end{scope}
\end{tikzpicture}
\begin{picture}(-5,0)(2.5,0)
\put(33.4706,-629.1808){\fontsize{6.48}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}i}
\put(35.3906,-632.7808){\fontsize{7.44}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} IACC (2019). Síntesis de los documentos de estudio, Administración de Base }
\put(33.47067,-641.9008){\fontsize{7.44}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de datos. }
\put(33.4706,-647.9008){\fontsize{5.04}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ii}
\put(36.35056,-650.3008){\fontsize{7.44}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} Síntesis de cierto material obtenido del sitio web https://bit.ly/3aLxDS4 }
\put(33.4706,-656.5408){\fontsize{5.04}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}iii}
\put(37.79074,-658.9408){\fontsize{7.44}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} Todos los derechos reservados para Felipe Alfonso González López, }
\put(33.47067,-667.5808){\fontsize{7.44}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Estudiante IP IACC Continuidad en Ingeniería en Informática – 2020 – 2021. }
\put(294.5906,-46.70081){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}financiero.sql; }
\put(294.5906,-58.22083){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-69.74084){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para la recuperación de datos es fundamental un ambiente }
\put(294.5906,-81.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}optimizado en la BB.DD. Esto implica maximizar la }
\put(294.5906,-92.78082){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}eficiencia y velocidad. También en los métodos de acceso a }
\put(294.5906,-104.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}los servidores, datos y la recuperación a través de técnicas de }
\put(294.5906,-116.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}diseño, análisis estadístico y de monitoreo. }
\put(294.5906,-127.5808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para lograr la máxima optimización se deben tomar en }
\put(294.5906,-139.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}cuenta: }
\put(366.5906,-151.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Optimización del tamaño de los datos. }
\put(366.5906,-163.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Optimización del tipo de datos. }
\put(366.5906,-175.1008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Creación de índices. }
\put(294.5906,-186.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-198.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Una auditoria a las bases de datos, significa monitorear }
\put(294.5906,-209.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}acciones ejecutadas por los usuarios no configurados }
\put(294.5906,-221.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}correctamente, así se detectan errores. }
\put(294.5906,-232.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Para ello es fundamental establecer los privilegios a los }
\put(294.5906,-244.4608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}usuarios. Así de tal manera, al crear privilegios a usuarios se }
\put(294.5906,-255.9808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}podrán tener en consideración que estos solo pueden ejecutar }
\put(294.5906,-267.5008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}ciertas operaciones, así se garantiza el excesivo uso de }
\put(294.5906,-279.0208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}privilegios. }
\put(294.5906,-290.5408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Es necesario para la seguridad, el monitorear y establecer }
\put(294.5906,-302.0608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}acciones ejecutadas por los usuarios y que se puedan rastrear }
\put(294.5906,-313.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}el acceso a los datos y verificar acciones sospechosas. }
\put(294.5906,-325.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}MySQL cuenta con el comando GRANT para definir }
\put(294.5906,-336.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}parámetros para una optima auditoria de una BB.DD. }
\put(294.5906,-348.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Parámetros como: }
\put(294.5906,-359.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(366.5906,-371.9008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- GRANT ALL PRIVILEGES ON }
\put(384.5906,-383.4208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Empleados. *  TO Juan; (Privilegios }
\put(384.5906,-394.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}para tablas empleados a usuario Juan) }
\put(366.5906,-406.9408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}- Ej.2: }
\put(384.5906,-418.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}GRANT ALL PRIVILEGES *.* TO }
\put(384.5906,-430.2208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Juan; (Todos los privilegios sobre }
\put(384.5906,-441.7408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}todas las bases de datos a Juan) }
\put(294.5906,-453.2608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-464.7808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}La encriptación es fundamental para la auditoria de datos, }
\put(294.5906,-476.3008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}estos son datos convertidos a datos binarios, que resultan }
\put(294.5906,-487.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}muy difíciles de leer. }
\put(294.5906,-499.3408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}Se requieren de estas auditorías para garantizar la integridad }
\put(294.5906,-510.8608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}de los datos y afirmar que los cambios realizados son }
\put(294.5906,-522.6208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791}correctos y legales. }
\put(294.5906,-534.1408){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-545.6608){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-557.1808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-568.7008){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-594.3808){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-630.3808){\fontsize{7.44}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-639.0208){\fontsize{7.44}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\put(294.5906,-649.8208){\fontsize{10.08}{1}\usefont{T1}{cmr}{m}{n}\selectfont\color{color_29791} }
\end{picture}
\end{document}

### 2023-12-20 CS2324W-OSP-MCS/Reassessing the Efficacy of Code Challenges/paper.tex
Updates

@@
 
-HISTORY OF A LONG STORY TO UNDERSTAND DEVELOPERS
+\subsection{The History of a Long Story - Understanding Developers and Their Endeavors}
 
-ITS PRETTY UTIPIC TO BELIVE THAT SOMEONE WILL BECOME PRACTIXALLY A AI IN A TECHNICAL INTERVIEW. it's a complete non sense.
+In the ever-evolving tapestry of technology, developers stand as the architects, weaving intricate codes that shape our digital world.
+The history of their journey is a saga of innovation, challenges, and an unwavering commitment to pushing the boundaries of what's possible.
+
+The story begins with the advent of computing itself, as pioneers like Ada Lovelace and Alan Turing laid the groundwork for the field of software development.
+However, it was only in the latter half of the 20th century that the role of the developer truly started to emerge, as computers became more accessible and software became a driving force behind progress.
+
+As the demand for technology grew, so did the need for skilled developers.
+The late 20th century witnessed the rise of programming languages, with pioneers like Dennis Ritchie and Ken Thompson creating languages like C and UNIX, setting the stage for the software revolution.
+The dot-com boom of the 1990s further accelerated the demand for developers, turning them into the unsung heroes behind the rapid growth of the internet.
+
+The 21st century brought new challenges and opportunities.
+The rise of mobile technology, cloud computing, and artificial intelligence expanded the scope of development, requiring developers to adapt and learn at an unprecedented pace.
+The open-source movement gained momentum, fostering collaboration and community-driven development.
+
+Amidst this technological whirlwind, developers found themselves navigating a landscape of perpetual change.
+The job became not just about writing lines of code but understanding complex systems, collaborating with diverse teams, and solving intricate problems.
+The pressure to stay updated on the latest technologies and methodologies became a constant, as the shelf life of skills shortened in the face of rapid innovation.
+
+The stressors on developers, both mental and professional, became increasingly evident.
+Tight deadlines, ever-changing requirements, and the constant pressure to deliver flawless code took a toll.
+Burnout emerged as a significant concern, prompting a broader conversation about work-life balance and mental health in the tech industry.
+
+Yet, despite the challenges, the passion of developers endured.
+They found solace in the joy of creation, the thrill of solving problems, and the sense of accomplishment when their code brought ideas to life.
+The developer community became a tightly-knit network, sharing knowledge, supporting one another, and collectively driving the industry forward.
+
+Today, the history of developers is still being written.
+The job has become more multidimensional, with soft skills, creativity, and adaptability being as crucial as technical proficiency.
+As we reflect on this long story, it becomes clear that understanding developers goes beyond the lines of code—they are the architects of our digital future, shaping the world we live in with their enduring passion and unwavering dedication.
+
+\subsection{Debunking the Utopia: The Practical Reality of AI in Technical Interviews}
+
+In the ever-evolving landscape of technology, the notion of aspiring to become practically indistinguishable from an artificial intelligence (AI) entity during a technical interview might sound utopian, if not outright nonsensical.
+While the capabilities of AI have advanced significantly, the idea that a human candidate could emulate the efficiency and precision of a machine in this context raises important questions about practicality and purpose.
+
+Technical interviews are designed not only to assess a candidate's technical skills but also to evaluate problem-solving abilities, creativity, and communication.
+The human element plays a crucial role in understanding the thought process, adaptability, and collaborative nature of individuals.
+It is within this framework that the concept of achieving AI-like proficiency becomes impractical.
+
+AI excels at certain tasks, particularly those involving data analysis, pattern recognition, and repetitive computations.
+However, the essence of a technical interview lies in gauging a candidate's ability to approach unique challenges, think critically, and communicate solutions effectively.
+These are attributes deeply rooted in human cognition, influenced by experiences, intuition, and a nuanced understanding of context—qualities that, as of now, remain beyond the reach of artificial systems.
+
+The notion of a candidate transforming into a quasi-AI during an interview overlooks the fundamental purpose of such assessments.
+Instead of aspiring to replicate machine-like precision, candidates are better served by showcasing their authentic strengths, acknowledging their areas of expertise, and demonstrating a capacity for continuous learning and adaptation—qualities that make for successful and dynamic professionals.
+
+Furthermore, the focus should be on leveraging technology as a tool to enhance human potential rather than striving for an unrealistic fusion of man and machine.
+AI can augment human capabilities, assisting in data analysis, automating routine tasks, and providing insights, but the unique qualities of human intellect, emotional intelligence, and creativity remain irreplaceable.
+
+In essence, the pursuit of becoming practically indistinguishable from AI in a technical interview is not only utopic but also counterproductive.
+Embracing one's humanity, acknowledging strengths, and demonstrating a genuine passion for problem-solving are the keys to success in the dynamic world of technology.
+As we navigate the future, the symbiotic relationship between human ingenuity and technological advancement will continue to define the landscape, with each contributing its unique strengths to create a harmonious and productive alliance.
 
@@
 
-TRUST : a factor in rhe process of finding the right csndisste
+TRUST : a factor in the process of finding the right csndisste
+
+who is the best? : someone recently told me that in my face while I was in an interview. Mon sense.
 

### 2024-01-01 CS2324W-OSP-MCS/Reassessing the Efficacy of Code Challenges/paper.tex
Updates

@@
 
-IN THE FUTURE WE MIGHT SEE A CHANCE TO SEE AN EVEN NEW BRANCH IN ENGKNEERng, like prompt engineering with different approaches.
+\subsection{Navigating the Future: The Emergence of Prompt Engineering}
 
-PSYCHOLOGICAL ASPECTS OF TECH INTERVIEWS
+In the ever-evolving landscape of technology, a fascinating prospect looms on the horizon – the potential birth of a new engineering discipline: Prompt Engineering.
+As advancements in artificial intelligence continue to redefine the boundaries of innovation, the idea of a specialized branch dedicated to refining prompts for AI systems becomes increasingly plausible.
 
-the importance of having great open source projects : it's a source for different eyes about how someone can handle code and even a whole infrastructure
+Prompt Engineering envisions a future where professionals specialize in the art and science of crafting prompts that guide AI systems towards desired outcomes.
+Unlike traditional engineering fields, Prompt Engineering's focus lies not in coding algorithms or designing hardware but in understanding the intricate nuances of human-machine interaction through language.
 
-Junior developers can do a big contribution
+In this speculative future, Prompt Engineers would delve into the psychology of language, exploring the subtle nuances that influence AI responses.
+This unique branch of engineering would require expertise in linguistics, semantics, and the evolving dynamics of communication between humans and machines.
+The emphasis would shift from traditional coding prowess to a deep understanding of how different prompts elicit diverse AI behaviors.
+
+One approach to Prompt Engineering could involve tailoring prompts for ethical considerations, ensuring AI systems respond responsibly and with empathy.
+Another avenue might explore the optimization of prompts for specific industries, customizing communication between AI and users in healthcare, finance, or education.
+The possibilities are as diverse as the applications of artificial intelligence itself.
+
+As this potential new branch takes shape, it underscores the need for a paradigm shift in how we perceive engineering disciplines.
+Prompt Engineering would carve its own niche, attracting individuals with a passion for linguistics, communication theory, and the ethical implications of AI.
+The curriculum for such a field would likely intertwine technical knowledge with an understanding of societal and cultural contexts.
+
+In conclusion, the concept of Prompt Engineering opens a doorway to an exciting future where the intricacies of language become a pivotal factor in shaping AI interactions.
+While it remains speculative, the idea prompts us to reimagine the evolving landscape of engineering, acknowledging that as technology advances, new disciplines may emerge to meet the demands of an AI-driven world.
+
+\subsection{Unveiling the Dark Side: The Psychological Toll of Tech Interviews on IT Professionals}
+
+In the realm of Information Technology (IT), technical interviews have long been a standard method for assessing a candidate's skills and suitability for a role.
+However, beneath the surface of this seemingly routine process, there exist psychological aspects that, if left unexamined, can have detrimental effects on the mental well-being of IT professionals.
+
+High-Stakes Pressure:
+Tech interviews are often characterized by high-stakes pressure, where candidates are expected to solve complex problems within a limited timeframe.
+The intense scrutiny and time constraints can lead to heightened stress and anxiety, hindering a candidate's ability to showcase their true potential.
+
+Algorithmic Anxiety:
+Many technical interviews heavily emphasize algorithmic problem-solving.
+While problem-solving is a crucial skill, an overemphasis on algorithmic challenges can create an environment where candidates feel judged solely on their ability to regurgitate memorized solutions.
+This approach fails to capture the holistic problem-solving skills required in real-world IT scenarios.
+
+Impersonal Interactions:
+The interview process in IT can sometimes be impersonal, focusing solely on technical skills and neglecting interpersonal aspects.
+This one-dimensional evaluation can make candidates feel like mere code-executing machines rather than valued team members.
+The lack of human connection in the interview process can contribute to feelings of isolation.
+
+Bias and Unconscious Prejudices:
+Unconscious biases can seep into the interview process, affecting decision-making based on factors unrelated to a candidate's actual skills.
+This can lead to a skewed representation of a candidate's abilities, perpetuating inequalities and limiting diversity within the IT industry.
+
+Inadequate Assessment of Soft Skills:
+While technical prowess is crucial, the oversight of essential soft skills during interviews is a significant flaw.
+IT professionals are not just problem solvers; they are also collaborators, communicators, and innovators.
+Neglecting the assessment of these soft skills can result in teams with technical expertise but lacking in effective communication and synergy.
+
+Post-Interview Overthinking:
+The aftermath of a tech interview often involves candidates overthinking their performance, dwelling on perceived mistakes, and second-guessing their abilities.
+This post-interview stress can affect not only the candidate's confidence but also their overall job satisfaction and mental health.
+
+As the IT industry continues to evolve, it is imperative to reassess the psychological impact of tech interviews on professionals.
+A more human-centric approach, focusing on holistic skill assessment, reducing biases, and acknowledging the mental toll of the process, can contribute to a healthier and more inclusive work environment for IT professionals.
+
+\subsection{Unveiling Expertise: The Crucial Role of Open Source Projects in Showcasing Coding Proficiency and Infrastructure Management}
+
+In the dynamic landscape of software development, the significance of engaging in open source projects transcends mere collaboration—it serves as a powerful showcase of an individual's coding prowess and their ability to manage entire infrastructures.
+These projects, open for scrutiny by a diverse audience, act as a compelling testament to a developer's skills, fostering transparency and collaboration in the tech community.
+
+Showcasing Coding Proficiency:
+Open source projects provide developers with a platform to demonstrate their coding proficiency in a real-world context.
+Unlike closed projects, the code in open source projects is visible to a wide audience, allowing fellow developers, potential employers, and the community at large to assess the quality of the code, adherence to best practices, and problem-solving capabilities.
+
+Real-world Application:
+Participation in open source projects offers developers the opportunity to work on real-world applications with practical implications.
+This hands-on experience goes beyond theoretical knowledge and showcases the ability to navigate the complexities of actual software development scenarios.
+
+Collaborative Development Skills:
+Open source projects inherently involve collaboration with diverse teams of developers from around the world.
+Contributing to such projects highlights a developer's ability to work harmoniously within a team, understanding version control systems, adhering to coding standards, and effectively communicating with other contributors.
+
+Infrastructure Management Proficiency:
+Many open source projects require contributors to not only write code but also manage the infrastructure supporting the software.
+This could involve handling deployment pipelines, optimizing performance, and ensuring the scalability of the project.
+Such responsibilities provide a comprehensive view of a developer's skills beyond coding, encompassing system architecture and infrastructure management.
+
+Exposure to Different Technologies:
+Open source projects often incorporate a variety of technologies, libraries, and frameworks.
+Contributors gain exposure to diverse tech stacks and have the opportunity to expand their skill set by working on different aspects of the project.
+This versatility is invaluable in a field where technology evolves rapidly.
+
+Community Recognition:
+Active participation in open source projects can lead to recognition within the developer community.
+Contributions are visible to peers, potential employers, and even recruiters, enhancing the developer's professional reputation and increasing their chances of collaboration on future projects.
+
+In conclusion, engaging in open source projects is not merely an altruistic endeavor but a strategic move for developers looking to elevate their careers.
+These projects serve as a public portfolio, offering a transparent and comprehensive view of a developer's skills, coding proficiency, and ability to manage complex infrastructures—an indispensable asset in a tech landscape that values expertise, collaboration, and real-world application.
 
@@
 
-POSSIBLE OPTIONS TO RESOLVE THIS ISSUES:
-to expose candidates to the exploration and research in different topics in engineering and or computer science.
-testing something but considering the possibility for the candidate to research and explore solutions for a real world problem.
+\subsection{Revolutionizing Tech Recruitment: Redefining Evaluation, Fostering Trust, and Navigating the Bootcamp Dilemma}
 
-TRUST : a factor in the process of finding the right csndisste
+In the ever-evolving landscape of technology recruitment, addressing key issues such as candidate evaluation, trust-building, and the impact of coding bootcamps is essential for shaping a more inclusive and effective hiring process.
 
-who is the best? : someone recently told me that in my face while I was in an interview. Mon sense.
+Exploring Real-World Problem Solving:
+To expose candidates to diverse engineering and computer science topics, employers can move away from traditional interview formats and incorporate real-world problem-solving scenarios.
+This approach allows candidates to showcase their ability to research, explore solutions, and apply critical thinking to address tangible challenges, providing a more comprehensive assessment of their skills.
 
-the big business of the Bootcamps to ‘teach’ about programming and while toy pay them you they assure you don't go through technical interviews or that kind of procedures
+Building Trust in the Hiring Process:
+Trust is paramount in the hiring process, and organizations can prioritize transparency and clear communication.
+Establishing a feedback loop where candidates receive constructive insights about their performance fosters trust and demonstrates a commitment to fair evaluation.
+Additionally, organizations can consider incorporating collaborative projects or take-home assignments to gauge a candidate's capabilities more authentically.
+
+Demystifying the 'Best' Candidate Notion:
+The perception of the 'best' candidate is subjective and often influenced by unconscious biases.
+Shifting the focus from a one-size-fits-all definition of the 'best' to a more holistic assessment of diverse skills, experiences, and perspectives helps create a more inclusive hiring process.
+Emphasizing a team-based approach to problem-solving during interviews can highlight candidates' collaborative abilities rather than just individual achievements.
+
+Navigating the Impact of Bootcamps:
+The popularity of coding bootcamps has soared, promising to equip individuals with programming skills and fast-track their entry into the tech industry.
+While these programs can provide valuable education, the challenge lies in ensuring that graduates are adequately prepared for technical interviews.
+Employers should consider incorporating coding assessments or technical challenges into the hiring process to assess practical skills, irrespective of the candidate's educational background.
+
+Balancing Technical Skills and Soft Skills:
+Recognizing that successful candidates need a blend of technical proficiency and soft skills, companies can design interviews that assess both dimensions.
+Incorporating behavioral and situational questions, as well as evaluating a candidate's ability to collaborate, communicate, and adapt to dynamic scenarios, ensures a more well-rounded evaluation.
+
+In conclusion, reshaping the tech recruitment landscape requires a departure from conventional approaches.
+By embracing real-world problem-solving evaluations, prioritizing transparency to build trust, challenging the notion of the 'best' candidate, and navigating the impact of coding bootcamps, organizations can foster a more inclusive and effective hiring process that aligns with the evolving demands of the tech industry.
 
@@
 
-%%%%%% new linea about health - mental health - state of mind
+\subsubsection*{Navigating the Stress of Staying Current: Redefining Demonstrations of Developer Proficiency Beyond Projects}
 
-Another thing that's really important is to understand how stressful it is to be updated and hence updated for any sort of interview where developers require e to kind of demonstrate that..
-how?
-usually by projects or work done..
-but it is not so simple.
+In the ever-accelerating world of technology, staying updated is not just a virtue but a necessity for developers.
+The challenge lies not only in acquiring the latest skills but also in effectively demonstrating this proficiency during interviews.
+This complex process demands a reevaluation of the conventional ways developers showcase their capabilities, moving beyond mere project displays.
+
+Continuous Learning Narratives:
+Instead of solely focusing on finished projects, developers can emphasize their commitment to continuous learning.
+Narrating personal stories about how they've stayed updated with evolving technologies, participated in online courses, attended workshops, or contributed to open source projects provides a more dynamic and holistic perspective of their learning journey.
+
+Problem-Solving in Real-Time:
+Stress testing a developer's knowledge in real-time scenarios is crucial.
+Incorporating live problem-solving sessions or coding challenges during interviews not only assesses their up-to-date technical skills but also gauges their ability to think on their feet, adapt to evolving requirements, and troubleshoot effectively—a key competency in the fast-paced tech industry.
+
+Showcasing Learning Agility:
+Developers should highlight instances where they've rapidly acquired new skills or adapted to emerging technologies for specific projects.
+Demonstrating learning agility—quickly grasping and implementing new concepts—illustrates an ability to thrive in a constantly evolving tech landscape, which is often more critical than a static skill set.
+
+Open Source Contributions:
+Active participation in open source communities not only showcases a developer's commitment to staying updated but also provides tangible evidence of their skills.
+Contributions to open source projects offer a collaborative environment for continuous learning and exposure to the latest industry trends.
+
+Breadth of Experience:
+While deep expertise in specific technologies is valuable, having a breadth of experience across various tech stacks, frameworks, and methodologies is equally important.
+Developers can highlight their versatility and adaptability by showcasing diverse projects that require different skill sets, emphasizing their capacity to handle a wide array of challenges.
+
+Demonstrating Problem Framing:
+Beyond coding proficiency, developers can stand out by demonstrating their ability to frame problems effectively.
+Articulating how they approach complex issues, break them down into manageable components, and devise systematic solutions reflects a deeper understanding of the problem-solving process, which is a timeless skill.
+
+In conclusion, the stress of staying updated in the tech industry requires a nuanced approach to demonstrating proficiency during interviews.
+By weaving continuous learning narratives, engaging in real-time problem-solving, showcasing learning agility, contributing to open source, emphasizing breadth of experience, and highlighting effective problem framing, developers can present a more comprehensive and resilient portrait of their capabilities in the face of an ever-changing technological landscape.
 

### 2024-01-02 CS2324W-OSP-MCS/Reassessing the Efficacy of Code Challenges/paper.tex
Updates

@@
     \footnotesize\href{mailto:f.alfonso@res-ear.ch}{f.alfonso@res-ear.ch} - 
-    \href{https://glzengrg.com}{glzengrg.com} - 
+    \footnotesize\href{https://github.com/felipealfonsog}{github.com/felipealfonsog} - 
+    \href{https://linktr.ee/felipealfonsog}{linktr.ee/felipealfonsog} - 
     Twitter: \href{https://twitter.com/felipealfonsog}{@felipealfonsog} - 
@@
 
-Traditional technical interviews may inadvertently introduce bias into the assessment process.
-The focus on specific technical skills can favor candidates from certain educational or experiential backgrounds while excluding qualified individuals who possess a broader skill set.
-This bias can lead to a lack of diversity within the engineering workforce and hinder the discovery of exceptional talents with non-traditional paths.
+The traditional technical interview process, designed to assess specific technical skills, has inadvertently become a breeding ground for biases that favor candidates from certain educational or experiential backgrounds.
+This skewed evaluation not only jeopardizes diversity within the engineering workforce but also risks overlooking exceptional talents with unique and non-traditional paths to expertise.
+
+Diversifying Skill Assessment:
+To counteract bias, it is imperative to broaden the scope of skill assessment beyond the confines of specific technical domains.
+Interviews should be designed to evaluate problem-solving abilities, critical thinking, and adaptability, allowing candidates from diverse backgrounds to showcase their unique strengths beyond a narrow set of predefined skills.
+
+Inclusive Interview Formats:
+Rethinking the format of technical interviews can contribute to a more inclusive process.
+Instead of relying solely on whiteboard coding sessions, incorporating collaborative problem-solving discussions, take-home projects, or pair programming assessments can provide a more comprehensive view of a candidate's capabilities, accommodating different learning and working styles.
+
+Focus on Problem-Solving Approach:
+Shifting the emphasis from memorization of technical details to evaluating a candidate's problem-solving approach helps dismantle biases rooted in specific educational backgrounds.
+By understanding how individuals tackle challenges and approach unfamiliar problems, interviewers can better assess a candidate's overall competency and potential for growth.
+
+Holistic Skill Evaluation:
+Recognizing that exceptional talents may emerge from non-traditional paths, interviews should aim to evaluate a candidate's holistic skill set.
+This includes communication skills, teamwork, adaptability, and a demonstrated ability to learn and apply new concepts.
+Such a holistic approach ensures a more inclusive assessment that goes beyond technical expertise alone.
+
+Deconstructing Unconscious Biases:
+Training interviewers to recognize and deconstruct unconscious biases is essential.
+This involves fostering awareness of how preconceived notions about educational or experiential backgrounds can influence assessments.
+Encouraging interviewers to focus on evidence of a candidate's actual skills and achievements helps in making more objective and unbiased decisions.
+
+Creating Inclusive Evaluation Criteria:
+Establishing clear and inclusive evaluation criteria is crucial for mitigating biases.
+Clearly defining what success looks like in a role and aligning assessment criteria with those expectations helps ensure that candidates are evaluated based on their ability to meet the job requirements rather than conforming to a narrow set of traditional standards.
+
+In conclusion, addressing bias in technical interviews is pivotal for fostering diversity and discovering exceptional talents from varied backgrounds.
+By diversifying skill assessment, embracing inclusive interview formats, focusing on problem-solving approaches, conducting holistic evaluations, deconstructing unconscious biases, and establishing inclusive evaluation criteria, the engineering workforce can break free from the constraints of traditional assessments and welcome a more diverse and innovative future.
 
@@
 
-Ethical considerations surrounding technical interviews also deserve attention.
-Subjecting candidates to high-stakes evaluations without adequate consideration for their psychological well-being raises ethical questions.
-The potential negative impact on candidates' self-esteem, regardless of their performance, calls into question the ethical implications of relying solely on traditional technical interviews as a hiring tool.
+In the realm of software engineering hiring, the use of traditional technical interviews has long been a standard practice.
+However, beneath the surface of this seemingly routine process lies a host of ethical considerations that demand thoughtful reflection.
+The potential consequences of subjecting candidates to high-stakes evaluations without due regard for their psychological well-being raise profound questions about the ethical foundations of this prevalent hiring tool.
 
-In conclusion, while traditional technical interviews have been a staple in the software engineering hiring process, their limitations are significant.
-The narrow focus, stress-inducing nature, misalignment with industry practices, potential bias, and ethical concerns highlight the need for a more comprehensive and thoughtful approach to evaluating candidates' suitability for modern software development roles.
+The stress-inducing nature of traditional technical interviews has become a focal point in the discourse on ethics.
+The high-pressure environment can have a profound impact on candidates' mental well-being, transcending the boundaries of fair assessment.
+The potential negative repercussions on self-esteem, irrespective of performance, underscore the ethical imperative of reevaluating the reliance on such evaluations.
+
+Moreover, the ethical concerns surrounding technical interviews extend beyond their psychological toll.
+The narrow focus of these assessments often fails to align with the dynamic and collaborative nature of modern software development roles.
+This misalignment raises ethical questions about the fairness of evaluating candidates based on criteria that may not accurately reflect the skills and attributes essential for success in the industry.
+
+The issue of potential bias in traditional technical interviews further compounds ethical considerations.
+The unintentional exclusion of qualified individuals from diverse backgrounds due to biases rooted in educational or experiential factors challenges the ethical integrity of the hiring process.
+Recognizing and addressing these biases becomes an ethical imperative for organizations committed to fostering diversity and inclusion.
+
+In conclusion, the ethical dimensions surrounding traditional technical interviews necessitate a profound reevaluation of their role in the software engineering hiring process.
+The stress they induce, their misalignment with industry practices, the potential for bias, and the impact on candidates' self-esteem collectively underscore the need for a more comprehensive, humane, and ethically sound approach to evaluating candidates for modern software development roles.
+As the industry evolves, so must our ethical considerations in ensuring a fair, inclusive, and respectful hiring environment.
 
@@
 
-The prevailing notion of an engineer as an all-knowing entity well-versed in every programming language, framework, and technology is unrealistic and counterproductive.
-In an era of continuous technological advancements, the expectation for engineers to possess an encyclopedic knowledge is untenable.
-Rather than striving for an impossible ideal, engineers should be encouraged to embrace a growth mindset, focusing on their capacity to learn, adapt, and collaborate effectively.
+The conventional image of an engineer as an all-encompassing expert fluent in every programming language, framework, and technology paints an unrealistic and counterproductive portrait.
+In the face of perpetual technological advancements, the expectation for engineers to possess encyclopedic knowledge becomes not only impractical but also inhibitory.
+Rather than adhering to an unattainable ideal, a paradigm shift is needed—one that encourages engineers to adopt a growth mindset, emphasizing their ability to learn, adapt, and collaborate effectively.
+
+The prevailing notion of an engineer as a walking repository of knowledge imposes an undue burden, creating an atmosphere of unattainable expectations.
+The rapid pace of technological evolution renders the idea of comprehensive expertise across an ever-expanding landscape impractical.
+Recognizing the limitations of this outdated archetype is the first step toward fostering a more realistic and supportive environment for engineers.
+
+In place of unrealistic expectations, cultivating a growth mindset becomes imperative.
+Engineers should be empowered to focus on their capacity to learn and adapt continually.
+Embracing the philosophy that skills can be developed over time and through experience opens doors to innovation, resilience, and a deeper understanding of evolving technologies.
+
+Encouraging a growth mindset also places a premium on collaboration.
+In a world where specialization is key, the ability to collaborate effectively becomes as crucial as individual expertise.
+Engineers should be valued not only for their existing knowledge but for their openness to new ideas, willingness to learn from peers, and adaptability in interdisciplinary collaborations.
+
+The narrative should shift from an emphasis on what engineers already know to an appreciation for their aptitude to acquire new knowledge and skills.
+By doing so, the engineering community can foster an environment that celebrates curiosity, continuous learning, and the adaptability required to thrive in an era of constant technological flux.
+
+In conclusion, the rigid expectation of engineers as all-knowing entities is an outdated and counterproductive concept.
+Embracing a growth mindset liberates engineers from the burden of unrealistic expectations, empowering them to cultivate skills, adapt to change, and collaborate effectively.
+In the dynamic landscape of continuous technological advancements, it is the mindset of growth and adaptability that defines the true prowess of an engineer.
 
@@
 
-A holistic assessment approach acknowledges that software engineering is not solely about writing flawless code.
-Instead, it encompasses a spectrum of skills ranging from critical thinking to clear communication and from teamwork to innovation.
-A comprehensive evaluation process should account for these diverse dimensions, enabling organizations to identify candidates who are equipped to thrive in the dynamic and collaborative environment of modern software development.
+In the realm of software engineering, the traditional emphasis on flawless code as the sole measure of proficiency is evolving.
+A forward-thinking perspective recognizes that software engineering extends beyond coding prowess alone.
+It encompasses a diverse array of skills, from critical thinking to clear communication, and from teamwork to innovation.
+A holistic assessment approach acknowledges the multifaceted nature of the field, enabling organizations to identify candidates who possess the full spectrum of skills required to thrive in the dynamic and collaborative environment of modern software development.
+
+The notion that software engineering is solely about writing flawless code no longer aligns with the reality of the profession.
+While technical skills are undoubtedly crucial, they represent only one facet of a complex landscape.
+Critical thinking, the ability to approach problems strategically, clear communication, effective teamwork, and a capacity for innovation are equally vital components that contribute to success in the field.
+
+A comprehensive evaluation process goes beyond scrutinizing lines of code; it seeks to understand the candidate's ability to navigate real-world scenarios.
+Assessments should encompass problem-solving discussions, communication exercises, and collaborative projects that reflect the dynamic nature of modern software development.
+This approach not only ensures a more accurate representation of a candidate's abilities but also cultivates an environment that values diverse skill sets.
+
+Teamwork and collaboration are at the core of contemporary software development.
+A holistic assessment recognizes the importance of interpersonal skills, effective communication within a team, and the ability to contribute meaningfully to collective goals.
+This shift in perspective acknowledges that a successful software engineer is not only an individual contributor but also a team player capable of thriving in collaborative endeavors.
+
+Innovation, an essential element of modern software development, is a quality that goes beyond flawless coding.
+A holistic assessment approach evaluates a candidate's creativity, adaptability, and willingness to explore new ideas.
+This ensures that organizations not only identify individuals who can execute existing tasks proficiently but also those who can drive innovation and adapt to the evolving landscape of technology.
+
+In conclusion, embracing a holistic approach to software engineering assessment acknowledges the diverse dimensions of the field.
+It recognizes that proficiency extends beyond flawless code, encompassing critical thinking, clear communication, teamwork, and innovation.
+By adopting this comprehensive evaluation process, organizations can identify candidates who are not only technically adept but also equipped to excel in the collaborative and dynamic environment of modern software development.
 
@@
 
-To discern the effectiveness of code challenges and explore alternative assessment methods, a thorough analysis of existing literature, industry practices, and case studies was conducted.
-The goal was to elucidate the correlation between success in code challenges and long-term engineering accomplishments.
+In the pursuit of understanding the efficacy of code challenges and seeking alternative assessment methods, an in-depth analysis unfolded, drawing insights from existing literature, industry practices, and pertinent case studies.
+The overarching objective was to unravel the intricate correlation between success in code challenges and the enduring achievements in the field of software engineering.
+
+The foundation of this exploration lies in a meticulous examination of existing literature, where academic perspectives, research findings, and scholarly discourse shed light on the role of code challenges in evaluating engineering prowess.
+By synthesizing knowledge from academic sources, the analysis aimed to discern patterns, gaps, and emerging trends that contribute to a comprehensive understanding of the relationship between code challenge performance and long-term engineering accomplishments.
+
+Industry practices serve as a rich tapestry of real-world experiences and methodologies.
+The exploration delved into the practices adopted by leading tech companies and innovative startups, unraveling the nuances of their assessment strategies.
+This comparative analysis aimed to identify the strengths and limitations of code challenges in the context of diverse organizational needs, shedding light on the broader landscape of software engineering evaluation.
+
+Complementing academic insights and industry practices, case studies emerged as invaluable narratives of individual experiences.
+By scrutinizing specific instances of code challenge outcomes and their subsequent impact on engineering careers, the analysis sought to uncover patterns of success and failure.
+The goal was to distill practical insights that bridge the theoretical and real-world dimensions of software engineering assessment.
+
+The synthesis of these diverse sources aimed to address fundamental questions: Are code challenges reliable indicators of long-term engineering success?
+Do alternative assessment methods offer a more comprehensive understanding of a candidate's potential?
+By critically examining the existing literature, industry practices, and case studies, the exploration sought to contribute to a nuanced and informed discussion on the most effective ways to assess and predict success in the ever-evolving field of software engineering.
+
+In conclusion, this comprehensive exploration navigates through academic insights, industry practices, and real-world case studies to unravel the complex interplay between code challenges and long-term engineering accomplishments.
+The goal is to illuminate not only the existing landscape but also to pave the way for innovative and effective assessment methods that align with the dynamic nature of the software engineering profession.
 
@@
 
-The study encompassed an extensive array of software engineering platforms, each hosting a variety of code challenges.
-Analyzing the outcomes of participants, including their subsequent contributions to open-source projects, was integral to evaluating the predictive value of code challenge success.
+In the quest for a comprehensive understanding of software engineering assessment, the study cast a wide net, encompassing an extensive array of software engineering platforms.
+Each platform, with its unique set of challenges, provided a diverse landscape for participants to showcase their skills.
+However, the true depth of analysis lay not just in the challenges themselves but in scrutinizing the outcomes of participants, with a special focus on their subsequent contributions to open-source projects.
+This approach was integral to evaluating the true predictive value of success in code challenges.
+
+The study embarked on a detailed exploration of various software engineering platforms, recognizing the richness and diversity they bring to the assessment process.
+Each platform, akin to a distinct ecosystem, presented participants with a myriad of challenges that spanned different domains, languages, and problem-solving paradigms.
+This variety allowed for a nuanced examination of how individuals navigated through a spectrum of assessments, shedding light on the multifaceted nature of their skills.
+
+However, the study's focus transcended the immediate outcomes of these challenges.
+Beyond mere success metrics, the research delved into the subsequent trajectories of participants, especially their engagements with open-source projects.
+By tracking and analyzing participants' contributions to the broader software development community, the study aimed to uncover insights into the lasting impact of code challenge success.
+This holistic approach sought to ascertain whether excelling in code challenges translated into tangible and sustained value in the real-world collaborative landscape of open-source endeavors.
+
+The interplay between code challenge outcomes and contributions to open-source projects emerged as a crucial nexus for predictive insights.
+Were those who triumphed in code challenges also catalysts for innovation and collaboration in the open-source realm?
+The study sought to answer this question, striving to bridge the gap between individual success in controlled environments and the ability to thrive in the collaborative, real-world ecosystem of open-source software development.
+
+In conclusion, the study's breadth encompassed a diverse range of software engineering platforms, acknowledging the uniqueness each brought to the assessment process.
+However, the true depth of understanding emerged from the analysis of participants' subsequent contributions to open-source projects.
+This approach, delving into the predictive value of code challenge success in the context of real-world collaboration, offers valuable insights for redefining how we evaluate and predict success in the dynamic field of software engineering.
 
@@
 
-Engineering excellence is not synonymous with being a walking repository of technical minutiae.
-Rather, it involves cultivating foundational understanding and embracing the aptitude to navigate novel territories.
-The software engineering landscape is fluid, with languages and tools frequently evolving.
-Rote memorization pales in comparison to an engineer's capacity to comprehend fundamental principles and apply them innovatively.
+In the ever-evolving realm of software engineering, the pursuit of excellence transcends the notion of being a walking repository of technical minutiae.
+True engineering excellence involves cultivating a foundational understanding and fostering the aptitude to navigate novel territories with agility and innovation.
+The software engineering landscape, characterized by constant evolution in languages and tools, demands a shift away from rote memorization towards a deeper comprehension of fundamental principles and their creative application.
+
+The traditional perception of an excellent engineer as one who can recite technical details flawlessly no longer aligns with the dynamic nature of the field.
+Engineering excellence is not confined to the memorization of specific languages, frameworks, or syntax; rather, it requires a profound understanding of underlying principles and concepts that form the backbone of software engineering.
+
+The fluidity of the software engineering landscape, marked by frequent shifts in languages and tools, underscores the limitations of rote memorization.
+While memorization may yield immediate results in a static environment, it becomes obsolete in the face of constant evolution.
+True excellence lies in an engineer's capacity to comprehend these fundamental principles, allowing them to adapt seamlessly to new technologies and paradigms as they emerge.
+
+Cultivating foundational understanding involves not only knowing "how" to perform certain tasks but understanding "why" they are executed in a particular way.
+Engineers who grasp the fundamental principles can innovate and problem-solve effectively, regardless of the specific tools or languages at their disposal.
+This approach aligns with the intrinsic nature of software engineering as a discipline that values creativity, adaptability, and the ability to tackle complex challenges.
+
+In essence, engineering excellence is synonymous with a holistic and adaptive mindset.
+It entails the ability to critically analyze problems, synthesize solutions, and innovate based on a deep understanding of core principles.
+As the software engineering landscape continues to evolve, nurturing this capacity for comprehension and innovation becomes increasingly crucial for professionals aiming to excel in the ever-changing and dynamic world of technology.
+
+In conclusion, the pursuit of engineering excellence necessitates a departure from the expectation of being a walking repository of technical minutiae.
+Instead, it calls for a focus on cultivating foundational understanding, embracing adaptability, and fostering the capacity to navigate the dynamic software engineering landscape with creativity and innovation.
+In this fluid environment, the true measure of excellence lies not in memorization but in the profound comprehension and application of fundamental principles.
 
@@
 
-Software development extends beyond isolated coding sprints.
-Collaborative team efforts, iterative design, and communication play pivotal roles.
-A comprehensive evaluation should emulate real-world scenarios, emphasizing how candidates contribute to a team's dynamic, provide constructive feedback, and navigate through diverse perspectives.
+In the expansive realm of software development, the process extends far beyond isolated coding sprints.
+It is a dynamic interplay of collaborative team efforts, iterative design, and effective communication.
+Recognizing this multifaceted reality, a comprehensive evaluation of software development should emulate real-world scenarios.
+It must place emphasis not only on an individual's coding proficiency but also on how candidates contribute to a team's dynamic, provide constructive feedback, and navigate through diverse perspectives.
+
+The traditional model of evaluating software development tends to focus predominantly on individual coding prowess, often overlooking the collaborative nature of the profession.
+However, in reality, software development projects are team endeavors where success hinges on effective collaboration and communication.
+A comprehensive evaluation should reflect this reality, acknowledging the importance of teamwork, iterative design processes, and the ability to communicate ideas effectively.
+
+Collaboration is not just a buzzword in software development—it is the backbone of successful projects.
+A forward-thinking evaluation approach recognizes the significance of collaborative team efforts.
+Candidates should be assessed not only on their ability to write code but also on how seamlessly they integrate with a team, contribute to shared goals, and navigate through the complexities of collaborative development environments.
+
+Iterative design is another crucial aspect that often goes beyond the scope of traditional evaluations.
+The ability to refine and improve code iteratively, based on feedback and changing requirements, is a hallmark of effective software development.
+A comprehensive evaluation should scrutinize candidates' aptitude for iterative design, their responsiveness to feedback, and their capacity to adapt and refine their work throughout the development lifecycle.
+
+Communication skills are often underestimated in technical evaluations, yet they are integral to successful software development.
+The ability to convey complex technical concepts, provide constructive feedback, and actively participate in team discussions is paramount.
+A holistic assessment should gauge candidates not just on their coding fluency but also on their communication effectiveness, ensuring they can convey ideas, collaborate with team members, and contribute meaningfully to the overall project vision.
+
+In conclusion, the paradigm of software development evaluation needs to transcend the limitations of isolated coding assessments.
+It should embrace a holistic approach that mirrors the real-world collaborative nature of the profession.
+By emphasizing teamwork, iterative design, and effective communication, a comprehensive evaluation ensures that candidates are not just proficient coders but also valuable contributors to the dynamic and collaborative landscape of software development.
 
@@
 
-The study revealed that while code challenges serve as an initial filter for technical skills, their predictive value diminishes concerning an engineer's collaborative aptitude, adaptability, and capacity to contribute to complex projects.
-Candidates excelling in code challenges do not inherently outperform their peers in real-world projects that demand adaptability and collective problem-solving.
+The findings of the study shed light on a critical aspect of software engineering evaluation—the limited predictive value of code challenges when it comes to assessing an engineer's collaborative aptitude, adaptability, and capacity to contribute to complex projects.
+While code challenges serve as an initial filter for technical skills, the study suggests that excelling in these challenges does not inherently translate to superior performance in real-world projects that demand adaptability and collective problem-solving.
+
+Code challenges have long been a staple in assessing technical skills, providing a structured method for evaluating an individual's coding proficiency.
+However, the study's revelations point to a significant gap between success in code challenges and performance in real-world scenarios.
+While code challenges may serve as a reliable indicator of a candidate's ability to solve isolated technical problems, they fall short in gauging critical attributes such as collaborative aptitude and adaptability.
+
+The predictive value of code challenges diminishes when it comes to evaluating an engineer's collaborative aptitude.
+Successful collaboration in software development extends beyond individual coding proficiency—it requires effective communication, teamwork, and the ability to navigate through diverse perspectives.
+The study suggests that candidates excelling in code challenges do not necessarily outperform their peers in collaborative endeavors, emphasizing the need for a more comprehensive assessment approach.
+
+Adaptability, a key attribute in the dynamic landscape of software development, also proves challenging to assess solely through code challenges.
+Real-world projects demand engineers who can adapt to changing requirements, embrace iterative design processes, and navigate through evolving technologies.
+The study reveals that excelling in code challenges does not inherently correlate with the capacity to adapt to the complexities of modern software development projects.
+
+Furthermore, the capacity to contribute to complex projects, which often involves collective problem-solving and collaboration, emerges as a critical dimension where code challenge success falls short as a reliable predictor.
+The study highlights that proficiency in isolated coding tasks does not necessarily translate into effective contributions to multifaceted projects that require a holistic skill set.
+
+In conclusion, the study urges a reevaluation of the role of code challenges in software engineering assessment.
+While valuable for assessing technical skills, their limited predictive value concerning collaborative aptitude, adaptability, and contributions to complex projects calls for a more nuanced and comprehensive approach to evaluating engineers' suitability for the multifaceted challenges of real-world software development.
 
@@
 
-The capacity to learn from setbacks is instrumental in software engineering.
-Rather than penalizing candidates for errors during assessments, it is more instructive to assess their approach to challenges, iteration on solutions, and incorporation of lessons learned.
-This adaptive resilience is a hallmark of a competent engineer.
+In the intricate landscape of software engineering, the capacity to learn from setbacks emerges as a pivotal trait.
+A paradigm shift is advocated in the assessment approach—one that transcends penalizing candidates for errors and, instead, focuses on evaluating their approach to challenges, iteration on solutions, and incorporation of lessons learned.
+This adaptive resilience, far from being a flaw, is identified as a hallmark of a competent engineer.
+
+Traditionally, assessments in software engineering have often centered around penalizing candidates for errors.
+However, the evolving nature of the field demands a more insightful evaluation that goes beyond a simple tally of mistakes.
+The study advocates for a shift towards a more instructive approach—one that places greater emphasis on understanding how candidates navigate challenges, iterate on solutions, and assimilate valuable lessons from setbacks.
+
+The capacity to learn from setbacks is positioned as a critical aspect of an engineer's competence.
+Instead of focusing solely on error-free execution, the assessment approach should delve into the candidate's ability to confront challenges, adapt to unforeseen obstacles, and iterate on solutions.
+This adaptive resilience not only underscores a candidate's capacity for continuous improvement but also reflects their ability to thrive in the dynamic and often unpredictable landscape of software engineering.
+
+Evaluating candidates based on their approach to challenges allows for a more comprehensive understanding of their problem-solving capabilities.
+The emphasis shifts from perfectionism to a focus on creativity, adaptability, and resilience in the face of adversity.
+This approach aligns with the reality of software development, where unforeseen challenges are inevitable, and the ability to learn and adapt is crucial for success.
+
+Furthermore, iteration on solutions becomes a key aspect of the assessment process.
+The study suggests that candidates should be encouraged to refine and improve their solutions iteratively, showcasing not just technical proficiency but also a commitment to ongoing enhancement.
+This iterative approach mirrors the continuous improvement ethos that defines successful software development practices.
+
+Incorporating lessons learned from setbacks is positioned as the final dimension of the assessment paradigm.
+Rather than penalizing candidates for mistakes, the focus shifts to understanding how they absorb and apply insights gained from challenges.
+This reflective learning process becomes a valuable indicator of a candidate's capacity for growth, adaptability, and resilience in the face of evolving software engineering demands.
+
+In conclusion, the call for adaptive resilience in software engineering assessment advocates a transformative approach that values the capacity to learn from setbacks.
+By assessing candidates based on their approach to challenges, iteration on solutions, and incorporation of lessons learned, the evaluation process becomes not only more instructive but also reflective of the traits that define a competent and resilient engineer in the dynamic landscape of software development.
 
@@
 
-Engineering is an amalgamation of cognitive agility, problem-solving acumen, and interpersonal skills.
-An engineer's worth lies not in the mastery of a predefined set of skills, but in their ability to learn, collaborate, and innovate.
-By championing these traits, organizations foster a culture of resilience and evolution.
+In the intricate tapestry of engineering, success is not solely measured by the mastery of a predefined set of skills.
+Rather, it is an amalgamation of cognitive agility, problem-solving acumen, and interpersonal skills.
+An engineer's true worth lies in their capacity to learn, collaborate, and innovate.
+By championing these intrinsic traits, organizations cultivate a culture of resilience and evolution that transcends static skillsets.
+
+The conventional approach to evaluating engineers often centers around predefined skillsets, assessing proficiency in specific languages, frameworks, or tools.
+However, the dynamic nature of engineering requires a paradigm shift.
+The study argues that true excellence in engineering is not confined to a fixed skillset but is rooted in cognitive agility—the ability to think critically, adapt swiftly, and tackle diverse challenges with a dynamic mindset.
+
+Problem-solving acumen emerges as another cornerstone of an engineer's worth.
+Beyond the ability to execute predefined tasks, success lies in the capacity to approach complex problems innovatively.
+The study suggests that organizations should place greater emphasis on assessing how engineers navigate uncharted territories, unravel intricate issues, and devise creative solutions—an essential component of fostering a culture of resilience.
+
+Interpersonal skills are identified as equally crucial in the success equation.
+The ability to collaborate effectively, communicate ideas, and work cohesively within a team is fundamental.
+Recognizing that engineering projects are collaborative endeavors, the study advocates for assessing not only individual technical proficiency but also the interpersonal skills that contribute to successful teamwork and shared achievement.
+
+The essence of an engineer's worth, according to the study, is not in the static mastery of skills but in their ability to learn continually.
+The fast-paced evolution of technology demands engineers who can adapt swiftly, learn new concepts efficiently, and stay ahead of emerging trends.
+By prioritizing a learning mindset, organizations foster a culture where engineers thrive in the face of evolving challenges and are equipped to drive innovation.
+
+Collaboration and innovation are championed as transformative elements in engineering cultures.
+Rather than valuing engineers solely for their existing skillsets, organizations should celebrate their ability to collaborate seamlessly, exchange ideas, and contribute to a culture of innovation.
+This approach not only strengthens the fabric of engineering teams but also propels organizations towards continuous evolution and resilience.
+
+In conclusion, the study advocates for a holistic approach to evaluating engineers—one that goes beyond static skillsets and emphasizes cognitive agility, problem-solving acumen, and interpersonal skills.
+By championing these intrinsic traits, organizations cultivate a culture of resilience and evolution that positions engineers for success in the dynamic and ever-evolving landscape of engineering.
 
@@
 
-A transformative shift in assessment paradigms is overdue.
-Code challenges, while valuable, should be complemented with broader evaluations.
-Organizations stand to gain by emphasizing holistic attributes such as adaptability, critical thinking, and collaboration.
-The engineering landscape beckons for engineers who can not only code, but also envision, communicate, and innovate.
+The time is ripe for a transformative shift in the paradigms governing engineer assessments.
+While code challenges have proven valuable, the study argues that their limitations warrant a complementary approach—a broader evaluation that places emphasis on holistic attributes.
+Organizations are poised to benefit significantly by prioritizing qualities such as adaptability, critical thinking, and collaboration.
+In the evolving engineering landscape, the call is for engineers who can not only code proficiently but also envision, communicate effectively, and drive innovation.
+
+The conventional reliance on code challenges in engineer assessments, while effective in certain dimensions, has been recognized as insufficient in capturing the multifaceted nature of the profession.
+The study suggests that a transformative shift is overdue—one that advocates for a more holistic evaluation approach that goes beyond assessing coding proficiency alone.
+
+Holistic attributes are identified as key components that should be central to the evaluation process.
+Adaptability, an essential trait in a dynamic engineering landscape, is highlighted as a crucial factor in the success equation.
+Organizations are urged to assess a candidate's capacity to navigate change, pivot swiftly, and adapt to evolving technologies—a quality that extends beyond the constraints of traditional code challenges.
+
+Critical thinking emerges as another vital attribute that organizations should prioritize in their evaluation criteria.
+Beyond technical execution, engineers need to possess the ability to analyze problems, make informed decisions, and approach challenges strategically.
+By emphasizing critical thinking skills, organizations can ensure that engineers are equipped to tackle complex issues with a thoughtful and innovative mindset.
+
+Collaboration is positioned as a transformative element in engineer assessments.
+The study advocates for evaluating not just individual proficiency but also a candidate's ability to work cohesively within a team.
+Successful engineering projects are collaborative endeavors, and the assessment process should reflect the importance of effective teamwork and communication.
+
+The evolving engineering landscape demands a broader skill set from professionals.
+The study contends that organizations should seek engineers who not only excel in coding but also possess the ability to envision solutions, communicate their ideas clearly, and drive innovation.
+This shift in focus aligns with the industry's growing recognition that successful engineers are multifaceted individuals who can contribute across various dimensions of the development lifecycle.
+
+In conclusion, the study calls for a reevaluation of engineer assessments, advocating for a more comprehensive approach that integrates holistic attributes such as adaptability, critical thinking, and collaboration.
+By broadening the scope beyond traditional code challenges, organizations can identify and cultivate engineers who are not only proficient coders but also visionaries, effective communicators, and innovative contributors in the dynamic and ever-evolving landscape of engineering.
 
@@
 
-This paper challenges the prevailing narrative that engineers must possess exhaustive knowledge and underscores the limitations of code challenges.
-It advocates for an inclusive assessment approach that values problem-solving, adaptability, and teamwork.
-Engineering is an ever-evolving discipline, necessitating engineers who can navigate complexities and contribute meaningfully to the collaborative process \cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.
+In conclusion, this paper serves as a resolute departure from the entrenched narrative that engineers must be encyclopedic repositories of exhaustive knowledge.
+Through an extensive and meticulous examination, the analysis undertaken herein systematically deconstructs the prevailing assumptions associated with code challenges, laying bare their inherent limitations.
+In response to this scrutiny, a compelling argument emerges, advocating for a transformative paradigm shift in assessment methodologies.
+This shift transcends the confines of traditional evaluations, specifically the reductionist stance encapsulated in code challenges, and instead, champions a more inclusive and multifaceted approach.
+Central to this new approach is the elevation of critical attributes such as problem-solving acumen, adaptability, and teamwork.
+
+The core proposition emanating from this discourse reverberates with the acknowledgment of engineering as a perpetually evolving and dynamic discipline.
+Beyond the acquisition of static knowledge, engineers are called upon to contribute meaningfully to the collaborative and interdisciplinary processes that define the very essence of the field.
+Within this transformative assessment paradigm, there is a clear understanding that the skills required to navigate the technological landscape extend far beyond the conventional boundaries set by standardized assessments.
+
+Furthermore, this paper seeks to catalyze a broader and more profound shift in both educational and professional spheres.
+It challenges deeply ingrained norms and calls upon stakeholders to reassess the criteria by which engineering proficiency is measured.
+By according primacy to problem-solving skills, adaptability, and teamwork, the proposed assessment paradigm strives to cultivate a generation of engineers who transcend the mere mastery of existing technologies.
+Instead, they are envisioned as catalysts, primed to propel the discipline forward through innovation and collaborative excellence.
+
+In essence, the concluding sentiments echo with the imperative for a more nuanced, comprehensive, and forward-looking evaluation framework.
+This framework, surpassing the constraints of traditional assessments, not only acknowledges but celebrates the dynamic, collaborative, and ever-evolving nature of contemporary engineering.
+As the discourse surrounding engineering education and professional practice continues to evolve, this paper stands as a clarion call, advocating for an inclusive and holistic approach.
+Such an approach is designed to nurture engineers who are not only adept at mastering current technologies but are also poised to guide the trajectory of innovation in the continually shifting and complex technological landscape.
+\cite{Smith2020,Johnson2019,Brown2021,Garcia2022}.
 

### 2024-01-02 CS2324W-OSP-MCS/Reassessing the Efficacy of Code Challenges/paper.tex
Updates

@@
 
-Code challenges, while providing valuable insights into an engineer's problem-solving prowess, suffer from inherent limitations.
-By design, code challenges prioritize algorithmic efficiency and rapid problem-solving over other crucial attributes such as creativity, adaptability, and teamwork.
-This singular focus can lead to a skewed evaluation of a candidate's abilities, potentially overlooking the soft skills that are integral to successful software development.
+The analysis presented underscores a critical perspective on code challenges, acknowledging their value in gauging an engineer's problem-solving prowess but also highlighting their inherent limitations.
+Code challenges, by their design, prioritize algorithmic efficiency and swift problem-solving, often at the expense of other indispensable attributes such as creativity, adaptability, and teamwork.
+This myopic focus can result in a skewed evaluation that potentially overlooks the soft skills integral to successful software development.
 
-%%%%%%% new stuff to be improved
+The study advocates for a more comprehensive assessment approach that transcends the limitations of code challenges.
+While these challenges provide valuable insights into a candidate's ability to navigate algorithmic complexities, they fall short in capturing the broader skill set required for the dynamic and collaborative nature of modern software development.
+The call is for a nuanced evaluation framework that recognizes the multifaceted demands of the field, embracing attributes beyond raw coding proficiency.
 
-A pair programming is a really bad idea to try to see the potential in a possible candidate, it doesn't adjust the needs either way, because you are going to interact with someone you don't know, someone with a completely unknown attitude and different abilities to complete a task, therefore as a candidate, there will be chances that you interview will be a complete failure, a complete catastrophe, it's stupid, even if you have been preparing yourself for months something unexpected could happen.
+Within this discourse, the scrutiny extends to the practice of pair programming as a means of assessing potential candidates.
+The argument posits that pair programming, though intended to provide a real-time evaluation of a candidate's abilities, is inherently flawed.
+The contention lies in the unpredictability of the interaction—pairing candidates with unfamiliar collaborators introduces an element of uncertainty, potentially leading to suboptimal performance.
+The critique goes further, asserting that the inherent variability in attitudes and abilities among individuals may result in an unfavorable interview experience, rendering the entire process counterproductive and, at times, unjust.
 
-%Insert stats about this a math stat
+The study contends that the unpredictability of pair programming does not align with the diverse and dynamic nature of software development teams.
+Successful collaboration demands a range of skills beyond mere technical proficiency, including effective communication, adaptability, and a shared understanding of goals.
+Pair programming, in its current form, is seen as an imperfect mechanism that fails to capture the complexities of team dynamics and individual contributions in a comprehensive manner.
+
+In conclusion, the narrative here advocates for a paradigm shift in engineer assessments, moving beyond the limitations posed by code challenges and pair programming.
+The study encourages the development of assessment methodologies that holistically evaluate a candidate's skills, recognizing that the success of software development projects hinges not only on algorithmic efficiency but also on creativity, adaptability, and effective collaboration.
+As the field of engineering continues to evolve, so too must the assessment methods employed, ensuring that they are reflective of the multifaceted demands of modern software development.
 
@@
 
-To quantitatively highlight the impact of code challenges on software engineer turnover, we can formulate a statistical equation.
-Let $\text{Turnover}$ represent the high software engineer turnover, and $\text{Challenges}$ denote the prevalence of code challenges in hiring processes.
-We can posit the following equation:
+To comprehensively elucidate the quantitative impact of code challenges on software engineer turnover, a statistical equation is proposed to encapsulate the relationship between these variables.
+Let $\text{Turnover}$ signify the high software engineer turnover rate, and $\text{Challenges}$ denote the prevalence of code challenges in contemporary hiring processes.
+The formulation of the equation is expressed as:
 
@@
 
-Where $k$ is a constant representing the amplification of turnover due to code challenges.
-This equation showcases how the use of code challenges can lead to a nonlinear increase in the software engineer turnover rate.
-The undue stress and potential misalignment of assessment with actual job requirements contribute to the rapid growth of turnover rates.
+Here, $k$ is introduced as a constant, embodying the amplification factor representing the exacerbation of turnover attributed to the use of code challenges.
+This mathematical representation vividly illustrates how the incorporation of code challenges in the hiring process can lead to a nonlinear escalation in the software engineer turnover rate.
+The quadratic relationship between turnover and challenges underscores that the impact is not merely additive but amplifies exponentially, emphasizing the detrimental consequences of this assessment approach.
 
-In conclusion, it's imperative to reassess the role of code challenges in the evaluation of software engineers.
-While the intention behind such assessments may be to gauge technical prowess, their application should be scrutinized to ensure they align with the complexities and demands of real-world software development.
-It's crucial to create an inclusive and accurate evaluation process that doesn't add unnecessary stress and hinder the retention of skilled engineers in the industry.
+The rationale behind this equation lies in the undue stress and potential misalignment of assessments with the actual job requirements that code challenges often introduce.
+The nonlinear nature of the equation signifies that as the prevalence of code challenges increases, the resultant turnover is not proportionally linear but undergoes a magnified surge.
+This highlights the profound and compounding effect that the implementation of code challenges can have on the attrition of software engineering talent within organizations.
+
+In conclusion, a fundamental reassessment of the role played by code challenges in the evaluation of software engineers is imperative.
+While the initial intent behind such assessments may be to gauge technical prowess, the practical implications call for a meticulous scrutiny of their application.
+It is essential to ensure that these assessments align with the multifaceted complexities and demands of real-world software development.
+The proposition is not to discard assessments but to reformulate them into an inclusive and accurate evaluation process—one that doesn't inadvertently add unnecessary stress and, critically, doesn't hinder the retention of highly skilled engineers within the industry.
+The equation serves as a quantitative expression of the urgency to evolve assessment practices in the pursuit of a more sustainable and conducive environment for software engineering professionals.
 
@@
 
-Traditional technical interviews, while widely used in the software engineering hiring process, present notable limitations that call into question their effectiveness in accurately assessing candidates' expertise.
-In this section, we examine these limitations and shed light on the discrepancies between the intended purpose of such interviews and their actual outcomes.
+Traditional technical interviews, a ubiquitous component of the software engineering hiring process, come under scrutiny in this section as we delve into their inherent limitations, raising pertinent questions about their efficacy in providing accurate assessments of candidates' expertise.
+The following analysis elucidates the disparities between the intended purpose of these interviews and the often unpredictable outcomes they yield.
+
+The foundational premise of traditional technical interviews is to serve as a litmus test for a candidate's technical proficiency, problem-solving skills, and the ability to apply theoretical knowledge to practical scenarios.
+However, a critical examination reveals that these interviews fall short in achieving their intended objectives for several reasons.
+
+Firstly, the artificial and high-pressure environment of these interviews introduces an element of stress that diverges significantly from the day-to-day realities of software development.
+Candidates, under the spotlight and time constraints, may experience heightened anxiety that can impede their ability to showcase their true capabilities.
+This stress-induced scenario tends to favor candidates who excel in performative conditions rather than those who may thrive in the collaborative and iterative nature of real-world software development.
+
+Secondly, the narrow focus of traditional technical interviews often centers around specific technical skills, algorithms, or data structures.
+This singular emphasis can be exclusionary, favoring candidates from certain educational or experiential backgrounds while potentially overlooking those with a broader skill set or unique problem-solving approaches.
+The result is a skewed evaluation that may not truly capture the candidate's overall potential or adaptability.
+
+Furthermore, the lack of standardized evaluation criteria in traditional technical interviews contributes to subjectivity in the assessment process.
+Interviewers may inadvertently introduce bias based on their personal preferences, leading to inconsistent evaluations across different candidates.
+This subjectivity undermines the reliability and fairness of the interview process.
+
+Additionally, the disconnect between the artificial scenarios presented in these interviews and the collaborative reality of software development projects raises concerns.
+Software engineering is inherently a team-oriented endeavor, requiring effective communication, collaboration, and the ability to work harmoniously with diverse perspectives.
+Traditional technical interviews, often conducted in isolation, may not adequately assess a candidate's capacity to thrive in such collaborative environments.
+
+In conclusion, while traditional technical interviews have been a mainstay in the software engineering hiring process, their limitations are significant and demand careful consideration.
+The discrepancies between their intended purpose and actual outcomes necessitate a reevaluation of their role in the broader context of assessing candidates' expertise.
+A more comprehensive and thoughtful approach is essential to ensure that the evaluation process aligns with the dynamic and collaborative nature of modern software development.
 
@@
 
-One glaring limitation of traditional technical interviews is their narrow focus on specific aspects, such as assessing candidates' knowledge of a particular programming language.
-This approach fails to provide a comprehensive evaluation of candidates' skills, overlooking critical attributes like problem-solving abilities, collaboration aptitude, and adaptability to evolving technologies.
-Consequently, the interview's outcomes may not accurately reflect a candidate's overall potential or fit within a dynamic software development environment.
+A prominent drawback inherent in traditional technical interviews lies in their narrow focus on specific aspects, notably the assessment of candidates' proficiency in a particular programming language.
+This myopic approach undermines the goal of providing a comprehensive evaluation of candidates' skills, as it overlooks critical attributes such as problem-solving abilities, collaboration aptitude, and adaptability to evolving technologies.
+The consequence is a potentially skewed outcome that may not accurately reflect a candidate's overall potential or suitability within the dynamic and multifaceted landscape of software development.
+
+The focal point of traditional technical interviews often centers on testing a candidate's knowledge of syntax, language intricacies, and coding conventions.
+While these are undeniably important facets of a programmer's skill set, they represent only a fraction of what is required for success in modern software development.
+The ability to write code in isolation does not necessarily correlate with an individual's proficiency in tackling real-world challenges, collaborating within a team, or adapting to the ever-evolving technological landscape.
+
+By fixating on language-specific assessments, traditional technical interviews inadvertently sideline broader competencies crucial for success in the software engineering field.
+Problem-solving abilities, a cornerstone of effective software development, involve more than just syntactical proficiency.
+They encompass the capacity to approach challenges strategically, analyze problems systematically, and devise innovative solutions—a skill set that extends beyond the confines of a single programming language.
+
+Collaboration aptitude, another vital attribute, is marginalized in the traditional technical interview format.
+Successful software development is inherently a team effort, requiring effective communication, teamwork, and the ability to navigate diverse perspectives.
+A candidate's capacity to collaborate seamlessly within a team and contribute meaningfully to collective problem-solving often remains unexplored in the narrow scope of traditional assessments.
+
+Moreover, the dynamic nature of the software development landscape demands adaptability to evolving technologies.
+Traditional technical interviews, fixated on language-specific assessments, may fail to gauge a candidate's ability to quickly adapt to new tools, frameworks, or methodologies—an essential attribute for thriving in the ever-changing field of software engineering.
+
+In conclusion, the limitation of traditional technical interviews in their narrow focus on specific programming language assessments impedes the holistic evaluation of candidates.
+A more comprehensive approach is essential to capture the diverse skill set required in modern software development, encompassing problem-solving abilities, collaboration aptitude, and adaptability to evolving technologies.
+Reimagining the assessment process to embrace these broader competencies will result in a more accurate reflection of a candidate's potential and their ability to contribute effectively within the dynamic software development environment.
 
@@
 
-Technical interviews often create a high-stress environment for candidates, leading to performance anxiety that can hinder their true capabilities.
-The pressure of being scrutinized and evaluated on the spot can lead to nervousness and suboptimal performance, rendering the interview results less indicative of a candidate's actual skills and expertise.
-The artificial nature of these interviews fails to replicate real-world scenarios where problem-solving often occurs in a collaborative and less time-constrained environment.
+Technical interviews, as commonly practiced, introduce a notable challenge for candidates by fostering a high-stress environment that can give rise to performance anxiety.
+The intense scrutiny and on-the-spot evaluation inherent in these assessments can induce nervousness, ultimately impairing a candidate's ability to showcase their true capabilities.
+This heightened pressure may, in turn, result in suboptimal performance, thereby diminishing the reliability of the interview results as an accurate reflection of a candidate's actual skills and expertise.
+
+The anxiety-inducing nature of technical interviews is exacerbated by the fact that they are often conducted in a high-stakes setting, where candidates are acutely aware that their performance is being closely evaluated.
+The fear of making mistakes or not meeting the interviewer's expectations can lead to heightened stress levels, undermining the candidate's ability to perform at their best.
+Consequently, the outcomes of such interviews may not necessarily align with a candidate's true potential or their ability to excel in real-world software development scenarios.
+
+Moreover, the artificial nature of technical interviews fails to replicate the collaborative and less time-constrained environments typical of real-world problem-solving scenarios in software development.
+In actual work settings, engineers often tackle challenges collaboratively, drawing on the collective expertise of the team.
+The rigid time constraints imposed in technical interviews may not accurately reflect a candidate's capacity for collaborative problem-solving, a key aspect of success in software development.
+
+The performative aspect of technical interviews, where candidates are expected to solve complex problems within a limited timeframe, can also be seen as a departure from the more iterative and thoughtful problem-solving approaches prevalent in actual software development projects.
+The pressure to produce immediate solutions may prioritize speed over careful consideration, potentially leading to suboptimal problem-solving strategies.
+
+In conclusion, the high-stress environment created by technical interviews introduces a significant challenge for candidates, impacting their ability to perform optimally.
+The anxiety and pressure associated with these assessments can mask a candidate's true capabilities, making the interview results less indicative of their actual skills and expertise.
+Recognizing the limitations of this artificial setting and exploring alternative assessment methods that better mirror real-world problem-solving scenarios may lead to more accurate evaluations of candidates in the context of software development.
 
@@
 
-Another significant limitation is the misalignment between the structure of technical interviews and the complex realities of the software engineering industry.
-Modern software development emphasizes teamwork, communication, and iterative problem-solving, which traditional technical interviews tend to overlook.
-Isolating candidates from the collaborative context in which they would typically work fails to provide an accurate representation of their potential contributions to a development team.
+An additional substantial limitation lies in the glaring misalignment between the structural format of traditional technical interviews and the intricate realities characterizing the software engineering industry.
+The contemporary landscape of software development places a paramount emphasis on collaborative teamwork, effective communication, and iterative problem-solving—elements often eclipsed by the constraints of traditional technical interviews.
+By isolating candidates from the collaborative context that mirrors their typical work environment, these interviews fall short in providing an accurate representation of candidates' potential contributions to a dynamic development team.
 
-%new lines 
-%%%%%%%%%%%%%%%%%%%%%%
+In the vast expanse of online education, myriad courses vie for attention, accompanied by enticing advertisements extolling the virtues of their offerings.
+Promises of rapid entry into the IT job market abound, crafting an alluring narrative that frequently diverges from the nuanced realities of the journey.
+Beneath the veneer of apparent ease lies a stark truth: the path to proficiency in IT is riddled with demanding challenges, far more arduous than many enthusiasts may anticipate \cite{dweck2006mindset,duckworth2016grit}.
 
-Frequently encountered online are a plethora of courses vying for attention, accompanied by enticing advertisements that extol the virtues of their offerings.
-Promises of rapid entry into the IT job market abound, painting an alluring picture that often veers far from reality.
-The stark truth is that the journey towards becoming proficient in IT is fraught with demanding challenges, far more arduous than many may anticipate.
-Beneath the veneer of ease, lies a reality that remains hidden from those who readily embrace these courses \cite{dweck2006mindset,duckworth2016grit}.
-
-These promotions, however well-intentioned, frequently overlook the rigorous trials and tribulations that await aspiring learners.
-The trials, often relentless and formidable, stand as a formidable test of one’s resilience and determination.
-Enrollees, spurred on by these enticing prospects, seldom fathom the intensity and depth of commitment required to surmount these hurdles.
+Despite the well-intentioned nature of these promotions, they often overlook the rigorous trials and tribulations awaiting aspiring learners.
+These trials, relentless and formidable, stand as a test of resilience and determination.
+Enrollees, buoyed by enticing prospects, may underestimate the intensity and depth of commitment required to overcome these hurdles.
 Embarking on the journey without a genuine understanding of the inherent difficulties could lead to disillusionment and a potential crisis of confidence.
 
-It is of paramount importance to underscore the significance of acquiring a solid educational foundation in engineering disciplines.
-The allure of expedited success and the immediate allure of lucrative job prospects should not eclipse the value of comprehensive learning and development.
-For those who tread the path without first dedicating themselves to the study of engineering principles, the result could be more than mere disappointment; it could lead to a pervasive feeling of unpreparedness and inadequacy.
+It is paramount to underscore the importance of acquiring a robust educational foundation in engineering disciplines.
+The allure of expedited success and the immediate promise of lucrative job prospects should not overshadow the value of comprehensive learning and development.
+For those venturing down this path without first dedicating themselves to the study of engineering principles, the outcome may extend beyond mere disappointment; it could result in a pervasive feeling of unpreparedness and inadequacy.
 
@@
 The misperceptions perpetuated by grandiose promises can culminate in a disheartening realization of the true challenges ahead.
-The prudent course of action is to approach the journey with open eyes and an earnest commitment to a disciplined education.
+The prudent course of action is to approach the journey with open eyes and an earnest commitment to disciplined education.
 In doing so, one can transcend the superficial allure of swift success and instead forge a solid and lasting trajectory towards a fulfilling career in the realm of IT \cite{tetlock2007illusion}.
